The Reock score was introduced by \citet{reock1961note} as a measure of how
much of an enclosing circle a district fills.
Its key conceptual insight is that the minimum bounding circle (MBC) is determined
entirely by the most extreme points of a district---its ``tips''---rather than
by the full boundary perimeter.
This makes Reock insensitive to boundary digitisation artifacts, coastline
irregularities, and river boundaries that strongly affect perimeter-based
metrics such as Polsby-Popper.

\subsection{Reock in Court}

Reock has been cited in redistricting litigation in several states:
\begin{itemize}
  \item \textbf{North Carolina}: Expert reports in \textit{Harper v.\ Hall} (2022)
    cited Reock scores alongside PP values.
    The NC state supreme court's compactness analysis referenced the general
    principle that compact districts should ``fill'' their bounding geometry~\citep{harper2022}.
  \item \textbf{Wisconsin}: Expert reports in \textit{Gill v.\ Whitford} (2018)
    included Reock computations for Assembly districts under the challenged
    legislative plan~\citep{gill2018}.
  \item \textbf{Georgia}: Redistricting litigation in Georgia has produced
    expert reports using Reock for House and Senate districts.
  \item \textbf{Florida}: The Florida Supreme Court-ordered redistricting (2015--2016)
    included Reock-based analysis by court-appointed experts.
\end{itemize}

\subsection{Why Reock for Coastal States}

For states with Atlantic, Gulf, or Pacific coastlines, PP is unreliable because
coastal boundaries contribute zigzag perimeter without corresponding district
``mass.''
Reock avoids this problem: the MBC is determined by the extreme points
(typically inland or cross-bay points), and the jagged coastline between
those extreme points does not affect the MBC.
This paper compares the current \textsc{bisect} Reock proxy's boundary
sensitivity against PP boundary sensitivity under controlled computational
conditions.  It keeps the canonical MBC definition visible while separating it
from the deployed centroid approximation.

\subsection{Moving-Knife Algorithm}

The moving-knife algorithm (MKA) is a \textsc{bisect} structure algorithm that
uses projected tract centroids to search for directional bisection cuts with
strong worst-side compactness.  It is not a global Reock optimizer and it does
not use polygon exact-MBC Reock inside the splitter.  The implementation projects
centroids, sorts units by candidate direction, cuts a population prefix, scores
both sides with a point-cloud minimum-enclosing-circle proxy, and keeps the best
worst-side candidate before local rebalance.
