Computing the exact minimum bounding circle (MBC) of a polygon is equivalent to
computing the smallest enclosing disk (SED) of its vertex set, solvable in
$O(n)$ expected time by Welzl's randomised algorithm~\citep{welzl1991smallest}.
The current \textsc{bisect} implementation does not yet call Welzl directly;
it uses the centroid-plus-max-radius approximation described below.

\subsection{Implementation in bisect}

The implementation in \texttt{crates/bisect-analysis/src/compactness.rs}
uses a \textbf{centroid-plus-max-radius approximation}: the bounding circle
centre is taken as the polygon centroid, and the radius is the maximum distance
from the centroid to any vertex on the polygon boundary.
This approximation matches the convention used in the Python redistricting
pipeline that bisect supersedes, ensuring comparability of Reock scores across
pipeline versions.

\begin{remark}[Future work]
Exact Welzl MBC remains future work.
The centroid-plus-max-radius approximation yields a circle that contains the polygon
but may be larger than the true MBC (the approximation errs toward over-coverage,
producing Reock values that are conservative---i.e., slightly lower than true-MBC Reock).
Practitioners requiring exact MBC-based Reock (e.g., for reproduction against
other redistricting software) should note this approximation.
\end{remark}

Test invariants (L0):
\begin{itemize}
  \item Reock of a circle is $\leq 1.0$ (the approximation may give Reock $< 1$
    for a circle if the centroid is not the true MBC centre; for a centred circle
    the approximation is exact).
  \item Approximation circle contains all input boundary points: for all $p$ on the
    boundary, $\|p - \text{centroid}\| \leq r + 10^{-9}$.
  \item Approximation circle area $\geq$ polygon area for any convex polygon.
\end{itemize}
