Reock has been used in redistricting litigation and expert testimony across
multiple states and jurisdictions.

\subsection{North Carolina: Harper v.\ Hall}

\textit{Harper v.\ Hall}, 384 N.C.\ 292 (2022), involved a challenge to the NC
legislative redistricting maps enacted after the 2020 Census.
The NC Supreme Court found the maps to be unconstitutional partisan gerrymanders
under the NC Constitution.
Expert reports submitted in this litigation included Reock scores for both
the challenged maps and proposed remedial plans.
The court's opinion referenced the principle that compact districts should
not have ``unnecessarily irregular boundaries,'' which expert witnesses operationalised
using Reock alongside PP~\citep{harper2022}.

\subsection{Wisconsin: Gill v.\ Whitford}

\textit{Gill v.\ Whitford}, 585 U.S.\ 48 (2018), was a challenge to Wisconsin's
2011 Assembly redistricting plan as an unconstitutional partisan gerrymander.
Expert reports filed in the district court included Reock computations for
Assembly districts under the 2011 plan and under simulated neutral alternatives~\citep{gill2018}.
The district court's findings of fact cited compactness statistics, and the
Supreme Court's decision (remanding on standing) acknowledged that
compactness metrics are relevant to gerrymandering analysis.

\subsection{Advantages of Reock for Coastal States}

Reock is preferred over PP in expert testimony for states with coastlines
because it avoids the boundary-sensitivity problem documented in Section~4.
When opposing counsel challenges PP values as inflated by coastline artifacts,
a Reock-based analysis provides a boundary-insensitive alternative that is
difficult to dismiss.
The MBC overlay is also visually intuitive for courtroom presentation:
the expert draws the minimum enclosing circle on the district map and the
court can see immediately what fraction of that circle the district occupies.

\subsection{Recommended Disclosure}

For court filings using Reock:
\begin{quote}
  \emph{All Reock scores were computed using a centroid-plus-max-radius approximation
  to the minimum bounding circle (MBC) of district polygons derived from 2020 Census
  TIGER/Line tract geometries projected to Albers Equal Area Conic (EPSG:5070).
  The approximation circle is centred at the polygon centroid with radius equal to the
  maximum distance from the centroid to any boundary vertex; it contains the polygon
  and may be slightly larger than the true MBC, producing conservative (slightly lower)
  Reock values. Reock $= A(\text{district})/A(\text{approx.\ MBC})$.
  Exact Welzl MBC computation remains future work.}
\end{quote}
