\section{Correction Events}

The \texttt{canvass-correction} fixture models a small but important case. The
unofficial snapshot reports one fewer vote for Candidate B in precinct P-001
than the canvassed snapshot. The package includes both sets of summaries and a
public correction event from unofficial to canvassed status.

\begin{center}
\small
\begin{tabularx}{\textwidth}{lrrrrrrr}
\toprule
Snapshot & A & B & WI & Under & Over & Blank & Counted \\
\midrule
P-001 unofficial & 40 & 34 & 1 & 3 & 1 & 0 & 79 \\
P-001 canvassed & 40 & 35 & 1 & 3 & 1 & 0 & 80 \\
Jurisdiction unofficial & 65 & 64 & 1 & 7 & 1 & 1 & 139 \\
Jurisdiction canvassed & 65 & 65 & 1 & 7 & 1 & 1 & 140 \\
\bottomrule
\end{tabularx}
\end{center}

The local equation for P-001 therefore changes by one declared vote, not by a
hidden overwrite:
\[
  40 + 34 + 1 + 3 + 1 + 0 = 79
  \quad\rightarrow\quad
  40 + 35 + 1 + 3 + 1 + 0 = 80.
\]

The verifier checks three things. First, each summary satisfies local contest
arithmetic: candidate votes plus residual counts equal counted ballots. Second,
each status-specific jurisdiction total equals the sum of its reporting units.
Third, a correction event exists with authority, source references, explanation,
and a transition from unofficial to canvassed.

The minimal normalized event contract is deliberately small:
\begin{center}
\small
\begin{tabularx}{\textwidth}{lX}
\toprule
Field & Meaning \\
\midrule
\texttt{event\_id} & Stable identifier within the package. \\
\texttt{event\_type} & Typed reason, such as correction, late-mail-batch-added, recount-update, court-order, or certification. \\
\texttt{status\_before}, \texttt{status\_after} & The declared transition. Initial unofficial reports may have the same before and after status. \\
\texttt{effective\_at} & Public event date or timestamp asserted by the source adapter. \\
\texttt{authority} & Office, board, court, or source authority making the public assertion. \\
\texttt{source\_refs} & Package source identifiers whose raw hashes bind the event to declared evidence. \\
\texttt{explanation} & Human-readable reason for the transition. \\
\bottomrule
\end{tabularx}
\end{center}

A compact record for this fixture is:
\begin{quote}
\footnotesize
\begin{verbatim}
{
  "event_id": "event-0002",
  "event_type": "correction",
  "status_before": "unofficial",
  "status_after": "canvassed",
  "effective_at": "2024-11-12T18:22:00Z",
  "authority": "SYN County Canvassing Board",
  "source_refs": ["source:canvass-minutes-2024-11-12"],
  "explanation": "Canvass correction added one Candidate B vote in P-001..."
}
\end{verbatim}
\end{quote}

This does not prove that the correction was lawful. It proves that the package
does not hide the delta. A public reader can see the before snapshot, the after
snapshot, the event metadata, and the source hashes that bind the package to its
declared evidence. A passing RCOUNT event check means "this transition is
declared and linked," not "this board action satisfies every statute."
