\subsection{Why Bisect Is MAUP-Robust}

The primary finding---that PP varies by less than 2\% across a 130$\times$
unit-count range for 48 of 50 states---calls for a mechanistic explanation.
We identify three structural reasons why bisect exhibits this robustness.

\paragraph{Graph-distance objective.}
The METIS partitioner minimises the weighted edge cut, which is an
\emph{abstract graph quantity} with no direct reference to node area.
Switching from tracts to block groups changes the graph (more nodes, more
edges, different edge weights) but preserves the topological skeleton: the
same geographic boundaries remain, and the same natural features that guide
cuts at tract resolution also guide cuts at block-group resolution.  The
algorithm converges to the same geographic cuts because those cuts minimise
edge weight regardless of how finely the intervening geography is
triangulated.

\paragraph{Polsby-Popper insensitivity to small boundary fluctuations.}
The PP metric scales as $A / P^2$.  A district with area 8{,}000 km$^2$
and perimeter 800 km has $PP \approx 0.157$.  Adding or removing a strip of
territory 100 m wide and 50 km long (approximately one block-group increment
of boundary resolution) changes perimeter by $\approx 0.1$ km and area by
$\approx 5$ km$^2$---a change of $<0.02\%$ in PP.  This calculation shows
that fine-grained boundary placement, while meaningful for VRA purposes
(capturing individual block groups with specific demographics), is essentially
invisible to PP.

\paragraph{Population-balance dominance.}
The binding constraint in bisect is population balance ($\pm 0.5\%$).  Once
population balance is enforced, the algorithm has limited freedom to choose
district shapes: both the county-level and block-group-level algorithms
must find a cut that satisfies the same population constraint, and the
resulting shapes are therefore approximately the same.  The resolution only
affects how precisely the boundary can ``track'' geography near the cut
points.

\subsection{The Small-$k$ Exception}

Idaho and Wyoming are the exception that proves the rule.  At $k=2$, the
single cut divides the state into two halves.  The constraint is
satisfied by many possible cuts, giving the algorithm real freedom to choose
among alternatives.  At county resolution (44 nodes for Idaho), the algorithm
chooses cut $C_1$.  At block-group resolution (770 nodes), the algorithm
finds cut $C_2$ that has slightly lower edge weight.  The two cuts produce
different PP values even though both satisfy the balance constraint.

The lesson is that MAUP-robustness is strongest when $k$ is large relative to
the number of available cuts.  For large states with many districts ($k \geq
8$), the population-balance constraint is tight enough to force all
resolutions to the same cut.  For $k=2$, there is genuinely more freedom and
the resolution can matter.  Practitioners should be aware of this for at-large
or two-district states.

\subsection{Policy Implications}

\paragraph{Research validity.}
Portfolio papers in the bisect series have reported compactness scores,
D-seat counts, and minority representation figures using census tracts.
This paper provides empirical validation that these findings are robust: if
the same experiments were run at block-group resolution (closer to operational
practice), the conclusions would be essentially identical.  This strengthens
the external validity of the tract-level literature.

\paragraph{Practitioner guidance.}
State redistricting commissions that prefer to draw districts from census
blocks or block groups can use the bisect algorithm directly at those
resolutions without any algorithm changes.  The MAUP robustness demonstrated
here means that the algorithmic behaviour characterised at tract level in
research papers generalises to finer-resolution operational use.

\paragraph{Legal significance.}
Courts evaluating algorithmic redistricting plans sometimes question whether
reported compactness metrics would change if a different spatial unit were
used.  This paper provides the first systematic empirical answer: for bisect,
compactness varies by less than 2\% across the resolution range tested, well
within legal thresholds for compactness comparison.

\subsection{Limitations}

Several limitations bound this study.  First, all results are single-run with
a fixed seed; ensemble variation at each resolution level is not characterised.
Second, we test three resolution levels; the very finest (census blocks,
$\sim$11 million nationally) is not tested due to computational cost.
Third, the election data used to compute D-seats is disaggregated from
precinct to block group by population-proportional allocation, which may
introduce measurement error in block-group partisan lean.  Fourth, the
geographic edge-weighting mode is tested; results for VRA-weighted or
county-weighted modes may differ slightly.
