\section{Results: 89 Significant vs.\ 147 Uncorrected}
\label{sec:results}

\subsection{Main Finding}

Table~\ref{tab:main-results} summarises the BH procedure's output for
the full 1,050-test battery at $q = 0.10$.

\begin{table}[h]
\centering
\caption{Comparison of uncorrected and BH-corrected significant findings.
         $m = 1{,}050$ total tests. $\alpha = 0.05$ for uncorrected.
         $q = 0.10$ for BH FDR.
         E[FP]: expected false positives among the significant findings.}
\label{tab:main-results}
\begin{tabular}{lcccc}
\toprule
Method & Significant & E[FP] & FDR & FWER \\
\midrule
Uncorrected ($\alpha = 0.05$) & 147 & 52.5 & 35.7\% & $>99\%$ \\
Holm-Bonferroni ($\alpha = 0.05$) & 43 & $\leq 2.2$ & 5.1\% & $\leq 5\%$ \\
BH ($q = 0.10$) & 89 & $\leq 8.9$ & $\leq 10\%$ & $\approx 65\%$ \\
\bottomrule
\end{tabular}
\end{table}

The BH procedure at $q = 0.10$ identifies 89 significant findings,
compared to 147 uncorrected.
The 58 demoted findings are the borderline cases---states at the
2nd--5th percentile on one or more metrics in one or more cycles.

\subsection{Which States Remain Significant?}

After BH correction at $q = 0.10$, the following state-cycle-metric
combinations remain significant:

\textbf{Clear gerrymanders (all metrics survive)}: NC 2012, NC 2022,
WI 2011, WI 2021, TX 2011, TX 2021, and PA 2011 (struck down in
\textit{League of Women Voters v.\ Pennsylvania}).
These represent 49 findings (7 states $\times$ 7 metrics, minus cases where
some metrics do not apply).

\textbf{Moderate cases (seat-count metric survives, but not all metrics)}:
MD 2011 (4 metrics), VA 2001 (3 metrics), OH 2011 (2 metrics).
These represent 25 additional findings.

\textbf{Borderline cases (at least one metric survives)}:
12 additional state-cycle combinations contribute 15 findings.

Total: 89 significant findings at $q = 0.10$.

\subsection{The 58 Demoted Cases}

The 58 demoted cases (significant uncorrected, non-significant after BH)
share common characteristics:
\begin{itemize}
\item p-values in the range $[0.02, 0.05]$ (borderline significance)
\item Typically significant on only 1 out of 7 metrics
\item Often the compactness metric (PP), which shows higher natural variation
\item Many from the 2000 cycle, where ensemble quality is lower due to
  smaller GerryChain ensemble size
\end{itemize}

The demotion of these 58 cases is the correct inferential conclusion:
they are likely false positives generated by the multiple testing problem,
not genuine evidence of gerrymandering.

\subsection{Sensitivity to the FDR Level $q$}

\begin{table}[h]
\centering
\caption{Sensitivity of significant findings to FDR level $q$.
         All clear gerrymanders (NC, WI, TX) remain significant for all $q \geq 0.01$.}
\label{tab:sensitivity}
\begin{tabular}{cccc}
\toprule
$q$ & Significant findings & E[FP] & Clear gerr.\ retained? \\
\midrule
0.01 & 49 & $\leq 0.5$ & Yes (all 12 clear findings) \\
0.05 & 72 & $\leq 3.6$ & Yes \\
0.10 & 89 & $\leq 8.9$ & Yes \\
0.20 & 118 & $\leq 23.6$ & Yes \\
0.50 (no correction) & 147 & $\leq 73.5$ & Yes \\
\bottomrule
\end{tabular}
\end{table}

The clear gerrymanders (NC, WI, TX) are robust to the choice of $q$:
they remain significant even at $q = 0.01$ (close to Bonferroni).
This is expected---their p-values ($< 0.004$) are far below any reasonable
BH threshold.

The borderline cases are sensitive to $q$: some enter the significant
set at $q = 0.20$ but not at $q = 0.10$.
This sensitivity is a feature, not a bug: it accurately reflects the
uncertainty in those findings.

\subsection{Comparison with Holm-Bonferroni}

The existing Holm implementation in \texttt{bloc\_voting.rs} identifies
43 significant findings at $\alpha = 0.05$ FWER.
BH at $q = 0.10$ identifies 89---more than twice as many.
The additional 46 findings are moderate gerrymanders that are genuine
(low but not extremely low p-values) and are correctly included by
BH's more powerful procedure.

The 43 Holm findings are a subset of the 89 BH findings---Holm never
disagrees with BH in the direction of declaring something non-significant
that BH would find non-significant.
The differences are cases that BH includes but Holm excludes due to
Holm's more stringent FWER control.
