\section{Conclusion}
\label{sec:conclusion}

\shb{} is a practical search mode for minimum-edge-cut compactness
optimisation.
By restarting short \recom{} chains from each burst's endpoint, it
concentrates exploration near good plans without sacrificing the
Markov property.
The checked-in mechanics package verifies the deterministic seed and endpoint
retention rules for a future \texttt{--search short-burst} surface.  The
15--22\% Polsby-Popper improvement over single-seed METIS remains a
state-run empirical claim that needs archived assignments, metrics, parameters,
and lineage before it is cited as production evidence.

The key correctness requirement is \emph{endpoint retention} rather
than within-burst minimization: retaining the burst endpoint
preserves the Markov restart property and avoids systematic bias
that would make the selected-plan sequence uninterpretable as a
well-defined stochastic process.

For the statutory baseline, \shb{} is most effective combined with
\ar{} (\texttt{--structure prime-factor}): \ar{} provides a good
initialisation, and \shb{} refines via concentrated neighbourhood
exploration.

\paragraph{Legal and practitioner use.}
\shb{} is appropriate for litigation support only when the submitted plan is
paired with its run package.  The 15--22\% PP improvement over single-seed
METIS, if reproduced in such a package, enables a straightforward argument:
the submitted plan was not merely drawn to pass a minimum compactness threshold
but was refined by neighbourhood exploration.
Because \shb{} operates on the initial METIS structure (preserving its
prime-factor bisection tree), the plan carries the same political-neutrality
properties as T.4's \ar{} plan while achieving measurably higher
Polsby-Popper compactness.
The efficiency-gap neutrality claim ($|\mathrm{EG}| < 0.05$ across NC, WI, TX)
is likewise a package-required empirical claim, not a consequence of the smoke
fixture alone.

\paragraph{Open directions.}

\begin{enumerate}
\item \textbf{Multi-run statistics.}
      This paper presents single-run results.
      Phase~2 will provide multi-run statistics (mean, standard deviation,
      and confidence intervals over 50 independent base seeds) to confirm
      that the reported improvements are consistent rather than
      seed-dependent.

\item \textbf{Combination with BisectionEnsemble.}
      The natural next step is to apply \shb{} as a post-processing
      step on the output of \be{}: \be{} optimizes each bisection node,
      and \shb{} then refines the final assembled plan.

\item \textbf{Burst length theory.}
      The empirical rule $\ell^* \approx k$ (burst length equal to
      the number of districts) warrants a formal analysis.
      The \recom{} step modifies $\approx 1/k$ of the plan per step;
      a burst of $k$ steps covers the plan once on average.
      Whether this is the optimal coverage rate for neighbourhood
      concentration is an open question.

\item \textbf{Non-uniform stationary distributions.}
      The stochastic dominance result (Proposition~\ref{thm:dominance})
      assumes a uniform stationary distribution.
      For compactness-weighted chains (where plans with higher $\pp$
      are preferred), the analysis changes: the chain naturally
      concentrates near good plans, and the Short-Burst advantage
      may be smaller or larger depending on the weight distribution.
\end{enumerate}
