\section{Proof Model}

The current RCOUNT proof sketch has a deliberately small public surface:

\begin{center}
\small
\begin{tabularx}{\textwidth}{lX}
\toprule
Field & Public meaning \\
\midrule
\texttt{proof\_id} & Stable package-local proof identifier. \\
\texttt{kind} & Accepted-token or counted-token proof kind. \\
\texttt{token\_hash} & Anonymized token digest, not a candidate-choice digest. \\
\texttt{reporting\_unit\_id} & Public reporting unit associated with the proof. \\
\texttt{candidate\_selections} & Must be empty in the baseline public proof. \\
\texttt{voter\_id}, \texttt{ballot\_style}, \texttt{issued\_at} & Must be absent in public baseline proofs. \\
\bottomrule
\end{tabularx}
\end{center}

This is a public proof contract, not a complete token-issuance protocol:

\begin{center}
\small
\begin{tabularx}{\textwidth}{lX}
\toprule
Contract point & Requirement \\
\midrule
Identifier scope & \texttt{proof\_id} is package-local and should not be a durable voter identifier. \\
Allowed evidence & The public record may expose proof kind, anonymized token hash, and coarse reporting-unit reference. \\
Forbidden evidence & Candidate selections, voter identity, ballot style, precise timestamps, device identifiers, and equivalent linkers are outside the baseline public proof. \\
Token caveat & A hash is not privacy by itself; token issuance, salts, nonces, key custody, and disclosure policy remain protocol work. \\
Gate meaning & Passing \texttt{proof\_privacy\_gate} is a necessary baseline check, not a formal coercion-resistance proof. \\
\bottomrule
\end{tabularx}
\end{center}

The distinction between accepted and counted matters. An accepted-token proof
follows ballot acceptance decisions: the voter returned material, the office
resolved eligibility and cure questions under local law, and the package
represents an anonymized accepted-record token. A counted-token proof follows
counting and any adjudication path represented in the package. Neither form says
which candidates were selected, whether a recount will later change the public
record, or whether a canvassing board has certified final totals.

The token is also not the same thing as a voter, ballot card, ballot, CVR row,
contest row, or vote. Those accounting relationships belong to the larger
RCOUNT evidence model. V.5 handles ballot manifests and V.6 handles
CVR-to-summary reconciliation; V.4 only defines the receipt-safe public edge of
an inclusion-style proof.
