\section{Privacy Gate}

The \texttt{proof\_privacy\_gate} is intentionally conservative. It rejects a
proof if it exposes candidate selections, combines voter identity with ballot
style and timestamps, or uses an invalid token hash. These are mechanical gates,
not a full privacy proof.

\begin{center}
\small
\begin{tabularx}{\textwidth}{lX}
\toprule
Actor demand & Why the baseline proof should not satisfy it \\
\midrule
Vote buyer asks for proof of a selected candidate & Candidate selections are absent and any choice-bearing proof fails. \\
Employer asks a worker to reveal a public proof id & The identifier is package-local and should not be a durable identity token. \\
Campaign combines ballot style and issue time & Ballot style and precise timestamps are forbidden in the public proof. \\
Observer targets a tiny reporting unit or rare contest & Small-cell and rare-contest metadata require disclosure review before publication. \\
Insider tries deterministic token linkage across exports & Token issuance, salts, nonces, and custody are protocol controls outside this public gate. \\
\bottomrule
\end{tabularx}
\end{center}

\begin{center}
\small
\begin{tabularx}{\textwidth}{lX}
\toprule
Rejected public field pattern & Risk \\
\midrule
Candidate selections in proof & Creates a direct vote-choice receipt. \\
Voter identity in proof & Links public proof material to a person. \\
Ballot style plus timestamp & May reidentify a voter in small populations. \\
Small reporting unit plus rare contest facts & Can create small-cell disclosure risk. \\
\bottomrule
\end{tabularx}
\end{center}

Passing the gate means only that the baseline public record avoids these obvious
receipt channels. It does not prove end-to-end coercion resistance.

The proof surface can be visualized as a narrow public window:

\begin{center}
\small
\begin{tabular}{ccc}
\fbox{\begin{tabular}{c}private election\\status records\end{tabular}} &
$\longrightarrow$ &
\fbox{\begin{tabular}{c}public proof\\kind + token hash\\+ coarse unit\end{tabular}} \\
 & & \\
\multicolumn{3}{c}{\textit{choices, voter identity, ballot style, precise time, and device facts stay outside}} \\
\end{tabular}
\end{center}
