\section{Boundaries}

RCOUNT does not decide voter eligibility. Eligibility, cure, provisional
adjudication, and ballot acceptance are legal and administrative decisions. A
proof can refer to an accepted or counted token only after those decisions have
been represented in the package.

The ordinary lifecycle is returned, accepted or rejected, counted if eligible
and countable, canvassed, and certified. Jurisdictions vary in cure windows,
provisional-ballot handling, recount procedures, public-record rules, and
voter-verification rights. V.4 therefore defines a minimum receipt-safety rule
for public artifacts, not a national canvass workflow.

RCOUNT also does not give voters a candidate-choice receipt. A public artifact
that proves a voter selected a candidate is a failure, not a feature. The
baseline proof sketch avoids candidate selections by construction.

Reporting-unit references need care. Vote centers, central-count mail batches,
split precincts, and batch boundaries may be less intuitive than a voter-facing
precinct label. In small units, even coarse metadata can become sensitive when
combined with rare contests, write-ins, ballot styles, or timing facts.

Finally, V.4 is not a complete end-to-end voting protocol. It does not specify
ballot casting, voter authentication, mixnets, cryptographic tallying, or a
formal coercion-resistance proof. It also does not specify token salts, nonces,
key custody, detached signatures, or issuance ceremonies. Those would require a
separate protocol paper. V.4 only defines the minimum privacy gate RCOUNT public
proof records must pass.
