\label{sec:legal}

\subsection{State-by-State Adjustment Requirements}

The threshold legal question for any redistricting plan in a state
with a prison population adjustment law is straightforward: does state
law require the adjusted or unadjusted Census data? The answer is
not always obvious, because some state adjustment laws apply only to
state legislative redistricting (not congressional redistricting),
and some apply only to the decennial census year, not to mid-decade
special redistricting cycles.

Table~\ref{tab:state-requirements} summarizes the scope of each state's
adjustment law based on statutory text as of 2026.

\begin{table}[ht]
\centering
\caption{Scope of State Prison Population Adjustment Laws}
\label{tab:state-requirements}
\begin{tabular}{llcc}
\toprule
State & Law & Congressional & State Legislative \\
\midrule
California & Cal.\ Elec.\ Code \S\ 21003 & No & Yes \\
Colorado & Colo.\ Rev.\ Stat.\ \S\ 2-1-901 & Yes & Yes \\
Delaware & Del.\ Code Ann.\ tit.\ 29, \S\ 804 & No & Yes \\
Illinois & 10 ILCS 5/3A-1 et seq. & No & Yes \\
Maryland & Md.\ State Gov't Code \S\ 2-2A-01 & No & Yes \\
Michigan & Mich.\ Comp.\ Laws \S\ 168.931 & No & Yes \\
Montana & Mont.\ Code Ann.\ \S\ 13-2-109 & No & Yes \\
Nevada & Nev.\ Rev.\ Stat.\ \S\ 218C.130 & No & Yes \\
New Jersey & N.J. Stat.\ Ann.\ \S\ 52:14-1.4 & No & Yes \\
New York & N.Y.\ Legis.\ Law \S\ 83-m & No & Yes \\
Virginia & Va.\ Code Ann.\ \S\ 24.2-304.03 & No & Yes \\
Washington & Wash.\ Rev.\ Code \S\ 44.05.160 & No & Yes \\
\bottomrule
\end{tabular}
\begin{tablenotes}
\small
\item Colorado is the only state whose adjustment law applies explicitly
to congressional redistricting. All others apply to state legislative
redistricting only. Expert witnesses must verify current law; statutes
may have been amended after this paper's writing.
\end{tablenotes}
\end{table}

Congressional redistricting in most adjustment states thus proceeds on
unadjusted Census data, while state legislative maps use adjusted data.
This asymmetry means that a state commission drawing both congressional
and state legislative maps must maintain two separate population datasets.
The \textsc{bisect} pipeline handles this through the proposed
\texttt{--adjust-prison-population} preprocessing mode: omitting the mode
for the congressional run and including it for the state legislative run.

\subsection{Algorithmic Neutrality and Court Filings}

A key advantage of the \textsc{bisect} approach to prison population
adjustment is its neutrality: the adjustment is applied uniformly
to all census tracts in all states covered by state law, without any
discretion about which facilities to adjust or how to weight the
home-county reallocation. The \textsc{bisect} SHA audit chain records
both the unadjusted and adjusted population inputs, enabling any
party to verify that the adjustment was applied consistently.

For expert witnesses, this neutrality is essential. Opposing counsel
cannot argue that the adjustment was selectively applied to benefit
one party, because the \textsc{bisect} implementation applies the same
DHC-A data and NPS home-county reallocation to every correctional
facility in the state. The adjustment methodology is documented in the
\textsc{bisect-data} source code and in the NPS data provenance chain.

\subsection{VRA Implications for Court Filings}

When prison population adjustment affects a minority-opportunity district's
BVAP or HVAP, the expert witness must explain the adjustment to the court.
The key points are:

\begin{enumerate}
\item \textbf{Population adjustment is required by state law}: In
adjustment states, using unadjusted Census data would produce a plan
that does not comply with the applicable statutory requirement for
the population base.

\item \textbf{The adjustment increases accuracy of minority representation}:
Because incarcerated people cannot vote, counting them in the prison
district rather than their home district overstates the political representation
of the prison district's residents and understates the representation of
the minority community's home district. Adjustment corrects this distortion.

\item \textbf{The VRA analysis uses the adjusted population}: In adjustment
states, the \textit{Gingles} threshold analysis (whether a majority-minority
district is feasible) should use the adjusted MVAP, not the unadjusted
Census MVAP. Using unadjusted data in an adjustment state would systematically
undercount the minority population of urban districts where majority-minority
districts are most likely to be required.
\end{enumerate}

\subsection{The Non-Adjustment Case: Federal and Non-Adjustment States}

In states without adjustment laws, and for congressional redistricting in
most adjustment states, \textsc{bisect} uses unadjusted Census data.
This is the legally required approach for those contexts: the Supreme Court
has not required states to adjust prison populations, and absent a state
law requiring adjustment, using the Census Bureau's official counts is the
legally safe choice.

Expert witnesses in non-adjustment states should nonetheless be prepared to
discuss prison gerrymandering if opposing parties raise it, particularly in
states with large prison populations (Texas, Florida, Pennsylvania, Ohio).
The appropriate response is that (a) the applicable law does not require
adjustment, (b) the Census Bureau's official methodology governs in the
absence of a contrary state requirement, and (c) \textsc{bisect} can produce
both adjusted and unadjusted plans for comparison if the court requests it.
