\label{sec:consequences}

\subsection{The 1941 Congressional Decision}

The Alabama Paradox was politically toxic.
It meant that:
\begin{enumerate}
  \item A state's senators could vote to expand the House,
    believing this would help their state, only to discover that
    their state lost a seat.
  \item The ``natural'' solution of increasing $H$ until all paradoxes
    disappeared was infeasible: the paradox can recur at any $H$.
  \item States near the remainder-seat boundary were perpetually
    uncertain about their seat count, making redistricting planning
    difficult.
\end{enumerate}

The National Academy of Sciences panel, reporting in 1929 and again
in 1941, recommended Huntington-Hill specifically because it is
a divisor method and therefore paradox-immune.
The 1941 Act adopted Huntington-Hill, ending the use of Hamilton
for federal apportionment.
The \textit{Montana} case (1992) confirmed that Congress's rejection
of quota compliance in favour of paradox immunity was constitutional.

\subsection{Implications for State Legislative Apportionment}

Some states use Hamilton for their own legislative apportionments,
accepting the quota-compliance guarantee at the cost of paradox
susceptibility.
This is legally permissible: the choice of apportionment method for
state legislatures is governed by state law and subject to the
one-person-one-vote standard of \textit{Wesberry v.\ Sanders} (1964)
and \textit{Reynolds v.\ Sims}, 377 U.S.\ 533 (1964),
but not to any specific formula requirement.

Redistricting practitioners should be aware that when advising on
state legislative apportionment, the choice of method carries
the Balinski-Young tradeoff: Hamilton offers quota compliance
at the cost of the Alabama Paradox; divisor methods offer
paradox immunity at the cost of occasional quota violations.

\subsection{Implications for \textsc{bisect-apportion}}

The \textsc{bisect-apportion} crate implements the
\texttt{check\_alabama\_paradox} function, which tests whether
a given population vector and House size produce a paradox
under Hamilton:
\begin{verbatim}
fn check_alabama_paradox(
    populations: &[(String, u64)],
    house_size: usize,
) -> Option<(String, usize, usize)>
\end{verbatim}
Returns \texttt{Some((state\_name, seats\_at\_H, seats\_at\_H+1))}
if a paradox is found, \texttt{None} otherwise.
For Huntington-Hill, this function always returns \texttt{None}
(formal guarantee from Theorem~\ref{thm:divisor-immune}).
For Hamilton, it may return a paradox for some House sizes.

The function enables practitioners to:
\begin{itemize}
  \item Verify that their apportionment method is paradox-free
    for the current census data.
  \item Demonstrate to a court that the chosen method is
    house monotone, supporting the argument that the method
    does not produce arbitrary results.
  \item Analyse whether alternative methods (e.g., Hamilton for a
    state legislative chamber) would produce paradoxes at relevant
    House sizes.
\end{itemize}

\subsection{International Context: Electoral Systems}

The Balinski-Young framework applies equally to PR electoral systems.
The equivalences in J.4 (Jefferson = d'Hondt) and J.2 (Webster =
Sainte-Lagu\"e) establish that the U.S.\ apportionment debate
is formally equivalent to the PR system debate.

Countries using Hamilton (Hare-Niemeyer method, as it is called
in Germany's historical context) are subject to the same paradoxes
for party-list seat allocation.
Germany switched from Hamilton to Webster (Sainte-Lagu\"e) in 2008
partly because of documented paradox instances in German Bundestag
seat allocation under Hamilton.
This international parallel confirms that the Balinski-Young theorem
is not a U.S.-specific curiosity but a universal mathematical
constraint on any proportional allocation system.
