Open seats---congressional districts in which no sitting incumbent's home census tract
has been assigned---are the most consequential redistricting outcome for electoral
competition. Open-seat races attract more candidates, receive more campaign investment,
produce stronger challengers, and are more likely to change partisan hands than
incumbent-protected seats. Enacted redistricting maps minimise open-seat creation by
engineering home districts for all sitting incumbents; algorithmic maps produce open
seats at the rate that geographic chance implies. This paper derives the analytical
expectation for open-seat creation under random redistricting for North Carolina
($k=14$), Wisconsin ($k=8$), and Texas ($k=38$), and compares \textsc{bisect}
algorithmic map outcomes to enacted 2022 congressional maps and to available
commission-drawn and court-drawn maps. We find that \textsc{bisect} produces
$4^\dagger$ open seats in NC, $2^\dagger$ in WI, and $7^\dagger$ in TX.
The NC bisect count of 4 is slightly above the random baseline 95\% CI of $[0,3]$---
a known consequence of the pairing effect documented in I.1. WI and TX fall within
their respective baseline ranges. All counts are approximately twice the open-seat
count of the corresponding enacted maps (NC: 2, WI: 1, TX: 3). Retirement incentive
analysis confirms that incumbents in \textsc{bisect} plans face structurally stronger
retirement pressure: the combination of higher pairing rates (I.1) and more competitive
districts (I.2) increases the fraction of incumbents who face either a primary opponent
or a genuinely competitive general election. We conclude that algorithmic redistricting
produces open-seat dynamics comparable to neutral human redistricting, not more
disruptive, but substantially more accountable than enacted maps designed to protect
all sitting members.
