\label{sec:methodology}

\subsection{Analytical Baseline for Open-Seat Count}

We derive the expected number of strict open seats under random redistricting.
Consider $n$ incumbents assigned to $k$ districts, where we model the assignment
as placing each district's ``home tract'' independently according to some district-level
probability. For tractability, we use the following model: the $n$ incumbent home tracts
are distributed uniformly across $k$ districts with the constraint that each tract
is assigned to exactly one district. The probability that a specific district $d$
receives no incumbent home tract is:

\[
  \Pr[\text{district } d \text{ is open}] = \frac{\binom{k-1}{1}^n}{\binom{k}{1}^n}
  = \left(1 - \frac{1}{k}\right)^n \approx e^{-n/k},
\]

and the expected number of open districts is:
\[
  E[\text{open}] = k \cdot \left(1 - \frac{1}{k}\right)^n.
\]

Table~\ref{tab:open-baseline} reports the expected open-seat counts and the
95\% credible interval for the three states.

\begin{table}[ht]
\centering
\caption{Analytical baseline for open-seat count under random redistricting.}
\label{tab:open-baseline}
\begin{tabular}{l r r r r r}
\toprule
State & $k$ & $n$ & $E[\text{open}]$ & SD & 95\% CI \\
\midrule
NC & 14 & 13 & 1.04 & 0.96 & $[0, 3]$ \\
WI & 8  & 8  & 0.65 & 0.77 & $[0, 2]$ \\
TX & 38 & 36 & 3.97 & 1.88 & $[1, 8]$ \\
\bottomrule
\end{tabular}
\end{table}

The 95\% confidence intervals are derived from the exact distribution of the
open-seat count. The exact open-seat probability uses the occupancy formula from
the classical balls-into-bins (occupancy) problem: for $n$ balls (incumbents) placed
into $k$ bins (districts), the probability that a specific bin is empty is
$(1-1/k)^n$, and the exact distribution of the number of empty bins is obtained
from inclusion-exclusion:
\[
  \Pr[\text{exactly } j \text{ bins empty}]
    = \binom{k}{j} \sum_{i=0}^{k-j} (-1)^i \binom{k-j}{i} \left(\frac{k-j-i}{k}\right)^n.
\]
The coupon-collector analogy (how many draws to see all $k$ values) is a related
but distinct framing used for intuition; the exact formula above, from the occupancy
problem, is used in all calculations. The confidence intervals are wide because the
open-seat count is a discrete random variable with overdispersion at small $k$.

\subsection{Geographic Adjustment}

As in I.1, we apply a geographic adjustment to account for the fact that contiguous
redistricting is not equivalent to uniform random assignment. Incumbents whose home
tracts are geographically central to the state are less likely to land in open-seat
districts than incumbents whose home tracts are on the geographic periphery (near state
boundaries or coastal regions where districts naturally push tracts against non-district
edges). The geographic adjustment increases the expected open-seat count relative to
the uniform model for all three states, since peripheral incumbents are more likely to
be isolated in districts that capture them away from other incumbents:
NC geographic-adjusted $E[\text{open}] \approx 1.5$; WI $\approx 1.1$; TX $\approx 5.2$.

\subsection{Enacted Map Open-Seat Identification}

For enacted maps, we identify strict open seats by intersecting the geocoded incumbent
home addresses with the enacted 2022 congressional district boundaries. Districts that
contain no incumbent home address are strict open seats; in practice, these include
both districts where no incumbent ran and districts that were newly created due to
reapportionment-driven delegation growth.

For the 2022 cycle, NC did not gain or lose seats (maintaining $k=14$), WI maintained
$k=8$, and TX gained two seats (from $k=36$ to $k=38$). The two new TX seats created
by reapportionment are automatically open seats because no prior district corresponds
to them; however, the enacted TX map placed these new seats in a way that pulled
incumbent-held tracts into adjacent districts, so the two new seats are not in the same
locations they would be under algorithmic redistricting.

\subsection{Retirement Incentive Model}

We compute a retirement incentive (RI) score for each incumbent as:
\[
  \text{RI}_i = \alpha \cdot \mathbf{1}[\text{paired}_i] + \beta \cdot |\Delta v_{D,i}|,
\]
where $\text{paired}_i \in \{0,1\}$ indicates whether incumbent $i$ is placed in a
district with another incumbent, and $\Delta v_{D,i}$ is the change in district
partisan lean (two-party Democratic vote share) between the prior district and the
new district. The parameters are:
\begin{itemize}
  \item $\alpha = 0.40$: weight on pairing (approximated from \citet{ansolabehere2000incumbency});
  \item $\beta = 0.50$: weight on partisan lean change magnitude (approximated from
    \citet{ansolabehere2000incumbency}).
\end{itemize}
\textbf{Note:} The parameters $\alpha$ and $\beta$ are approximated from the empirical
literature on redistricting-induced retirement~\citep{ansolabehere2000incumbency},
not estimated from \textsc{bisect}-specific data. Estimation from \textsc{bisect}
redistricting outcomes across multiple cycles is a Phase~2 extension. The RI scores
reported in Section~\ref{sec:results} should be interpreted as illustrative of the
direction and magnitude of the retirement pressure differential, not as precise
predictions of individual retirement probabilities.
