\label{sec:introduction}

When a state legislature or redistricting commission proposes an algorithmic map for
legislative approval or court review, incumbents and their allies reliably raise a
specific objection: the algorithm does not consider incumbency. This objection takes
several forms. In legislative proceedings, it appears as a policy argument---``We have
a responsibility to maintain continuity in representation''---invoking the constituents'
interests in an uninterrupted relationship with their existing representative. In
litigation, it appears as a legal argument---``The enacted map appropriately considered
the legitimate redistricting criterion of preserving the cores of prior districts and
avoiding contests between incumbents''---citing the Supreme Court's language in
\textit{Karcher v.\ Daggett}\footnote{\textit{Karcher v.\ Daggett}, 462 U.S.\ 725, 740
(1983).} as authority for the proposition that the algorithm's failure to consider
incumbency is a legal defect.

Both forms of the objection deserve a thorough response. This paper provides that
response, grounded in the federal constitutional doctrine, VRA requirements, and state
constitutional law that govern redistricting in the three states studied in I.1--I.3.

\subsection{The Objection in Three Forms}

The incumbency objection to algorithmic redistricting presents itself in three distinct
legal framings:

\paragraph{The \textit{Karcher} objection.}
The Supreme Court in \textit{Karcher v.\ Daggett} identified ``avoiding contests between
incumbents'' as a ``legitimate objective'' that may justify minor population deviations
from mathematical equality. Opponents of algorithmic maps argue from this language that
incumbency consideration is a constitutionally endorsed redistricting criterion, and that
a redistricting methodology that ignores it is legally incomplete or defective.

\paragraph{The VRA objection.}
In states with majority-minority districts, protecting minority incumbents is sometimes
argued to be required by VRA \S{}2 as part of the remedy for vote dilution.
If \textsc{bisect} maps pair a minority incumbent with an Anglo incumbent, opponents
argue that the algorithmic indifference to incumbency has produced a map that dilutes
minority voting strength by forcing a primary contest that the minority incumbent may
lose.

\paragraph{The state law objection.}
Several states have redistricting criteria statutes or commission guidelines that
explicitly list ``preserving communities of interest'' or ``maintaining incumbency''
as permissible considerations. Opponents of algorithmic maps in these states argue
that the algorithm's failure to consider these factors violates state statutory law.

\subsection{The Response in Three Parts}

This paper develops a three-part response:

\begin{enumerate}
  \item Incumbency protection is permissible under federal constitutional doctrine, but
    not required. \textit{Karcher}'s identification of it as a ``legitimate objective''
    does not mandate its consideration; it merely confirms that a mapmaker who considers
    it is not constitutionally defective. An algorithm that does not consider it is
    equally not constitutionally defective.
  \item The VRA objection conflates the VRA's requirements for district drawing with
    requirements for \textit{how} districts are drawn. The VRA requires that minority
    voters have an equal opportunity to elect candidates of their choice; it does not
    require that the redistricting process protect any specific incumbent's electoral
    prospects. When \textsc{bisect} maps produce VRA-compliant majority-minority
    districts, the VRA is satisfied regardless of where specific incumbents' homes happen
    to fall.
  \item State redistricting criteria that permit incumbency consideration impose a duty
    to consider, not a mandate to protect. An algorithmic process that systematically
    excludes incumbency data cannot ``consider'' incumbency in the way state statutes
    contemplate; but the proper remedy---if one is needed---is to run the algorithm
    with an awareness of its incumbency outputs and to make post hoc adjustments where
    VRA compliance requires them, not to inject incumbency data into the algorithmic
    optimisation.
\end{enumerate}

\subsection{Paper Organisation}

Section~\ref{sec:background} reviews the federal constitutional doctrine on
incumbency, covering the full lineage from \textit{Karcher} through \textit{Cox v.\
Larios}. Section~\ref{sec:methodology} develops the \textit{Karcher} doctrine in
detail, clarifying what the ``legitimate objective'' language does and does not
authorise. Section~\ref{sec:results} analyses the VRA interaction, drawing on
\textit{Thornburg v.\ Gingles} and its progeny. Section~\ref{sec:discussion}
surveys state constitutional frameworks and develops the algorithmic indifference
standard. Section~\ref{sec:conclusion} concludes with practical guidance for
practitioners using \textsc{bisect} in litigation.
