\label{sec:results}

\subsection{VRA \S{}2 and Incumbency: The \textit{Gingles} Framework}

The Voting Rights Act of 1965, \S{}2, as interpreted in \textit{Thornburg v.\
Gingles}\footnote{\textit{Thornburg v.\ Gingles}, 478 U.S.\ 30 (1986).} and its progeny,
prohibits redistricting plans that result in the denial or abridgement of the right
to vote on account of race or color. The \textit{Gingles} three-part test for
demonstrating vote dilution requires plaintiffs to show: (1) the minority group is
sufficiently large and geographically compact to form a majority in a single-member
district; (2) the minority group is politically cohesive; and (3) the white majority
votes sufficiently as a bloc to usually defeat the minority-preferred candidate.

When all three prongs are met, the redistricting body must draw a majority-minority
district in the relevant area, unless doing so would require deviation from other
constitutionally or statutorily required criteria~\citep{gingles1986}. The question
for algorithmic redistricting is whether the VRA's majority-minority district requirement
independently mandates that the district be drawn in a way that protects the incumbent
minority representative's home address.

\subsection{The VRA Does Not Require Incumbency Protection}

The answer is no. The VRA requires a district in which minority voters have an equal
opportunity to elect candidates of their choice---not that the specific candidate
currently representing that district be protected from competitive challenges.
\textit{Gingles} and its progeny focus on the electoral opportunity available to
minority voters as a group, not on the electoral security of any individual incumbent.

This distinction is practically significant because minority incumbents, like all
incumbents, benefit from incumbency advantage that provides electoral security even
without gerrymandered district protection. A minority incumbent who faces a primary
opponent following redistricting is not, by that fact alone, the victim of a VRA
violation; the VRA violation would require showing that the redistricting has diminished
the minority group's opportunity to elect their preferred candidate below the level
that existed prior to redistricting.

\subsection{The Tension When Incumbency and VRA Conflict}

A genuine tension arises when protecting a minority incumbent's residency core
conflicts with the minority community's VRA-required opportunity district. This can
occur in two scenarios.

\paragraph{Scenario 1: Anglo incumbent in majority-minority area.}
If an Anglo incumbent represents a district that has, through demographic change,
become majority-minority, the VRA may require redrawing the district to create a
majority-minority district. Protecting the Anglo incumbent's home tract may conflict
with the VRA requirement to create the majority-minority district. Courts have
consistently held that the VRA requirement takes precedence over incumbent protection
in this scenario~\citep{thornburg1986progress}.

\paragraph{Scenario 2: Minority incumbent in a contested area.}
If a minority incumbent represents a district that is being split or reconfigured, the
VRA may appear to require drawing boundaries that protect the incumbent's district as
a majority-minority seat. However, this scenario is more complex: the VRA requires
a majority-minority \textit{district}, not a majority-minority district that includes
a specific incumbent's home address. The algorithmic map may produce a majority-minority
district that satisfies the \textit{Gingles} test without including the specific
incumbent's home tract.

\subsection{Algorithmic Maps and VRA Compliance}

The \textsc{bisect} algorithm does not automatically produce VRA-compliant maps;
VRA compliance is evaluated ex post and, where required, the algorithm can be
re-run with VRA constraints (see the \texttt{--weights-override vra-aligned} flag
in the \textsc{bisect} CLI documentation). The key point for the incumbency analysis
is that VRA compliance and incumbency protection are analytically separable:

\begin{enumerate}
  \item A \textsc{bisect} map that produces a majority-minority district in the
    geographically appropriate area satisfies VRA \S{}2 for that area, regardless
    of whether the minority incumbent's home tract falls within that district.
  \item If the minority incumbent's home tract does not fall in the majority-minority
    district, the incumbent may seek to run in that district (residency requirements
    for Congress are at the state level; most states require only that a candidate
    reside in the state, not the district), or may seek election in an adjacent
    district, or may retire. None of these outcomes is a VRA violation.
  \item If the majority-minority district required by the VRA happens to not include
    the existing minority incumbent's home tract, the appropriate resolution is to
    adjust the map's boundaries to include the incumbent's tract within the
    majority-minority district---a minor adjustment that does not require re-running the
    full algorithm---rather than to inject incumbency data into the algorithmic
    optimisation.
\end{enumerate}

\subsection{State Constitutional Constraints}

Survey of state constitutional provisions reveals a spectrum of approaches to incumbency
as a redistricting criterion.

\paragraph{Explicit permission.}
Several states explicitly permit consideration of incumbency in their redistricting
criteria statutes or constitutional provisions. California's commission guidelines
permit ``maintaining communities of interest and neighbourhoods'' as a criterion,
which has been interpreted to include maintaining existing constituent-representative
relationships~\citep{ncsl2021criteria}. Ohio's constitutional redistricting criteria
(pre-amendment) allowed ``maintaining the cores of prior districts,'' which provided
an implicit basis for incumbency protection.

\paragraph{Explicit prohibition.}
Arizona and California's independent redistricting commission enabling statutes
explicitly prohibit the commission from drawing districts to favour any incumbent or
political party. This prohibition is directional (commissions cannot protect incumbents)
but does not speak to whether algorithms that are indifferent to incumbency comply
with the prohibition. We argue that they do: an algorithm that neither protects nor
targets incumbents satisfies a prohibition on incumbency protection.

\paragraph{Silent frameworks.}
Most state constitutions and redistricting statutes do not mention incumbency at all,
leaving the question to the redistricting body's discretion within the bounds of
constitutional and statutory criteria. In these states, the algorithmic indifference
to incumbency is legally unremarkable: the algorithm is not required to consider
incumbency, and its failure to do so cannot be characterised as a legal defect.
