\section{Current Data and Validation Boundary}

This paper is an external-comparison protocol unless the GerryChain traces are
archived with state, year, graph identity, population tolerance, proposal
kernel, seeds, step count, burn-in/thinning rule, metric version, and diagnostic
summary.  Percentile statements such as ``0.2nd percentile'' should be cited
only from those artifacts.

The BISECT-side plan also needs its RPLAN/RCTX package and verifier status.  A
comparison between a deterministic plan and an ensemble is meaningful only when
both sides are hash-bound and the sample scope is clear; it does not prove legal
sufficiency or ensemble representativeness by itself.

The current repository records this as a concrete package gap rather than a
completed empirical result.  The manifest at
\texttt{docs/examples/g-ensemble-evidence-packages/G.1-G.3+missing-evidence/}
is a hash-bound missing-evidence package: it validates the absence of archived
external traces, metric outputs, diagnostics, and RPLAN/RCTX baseline packages
for the headline percentile claims.  Those claims should not be upgraded from
protocol statements to findings until that package is replaced or supplemented
by active evidence packages.

The repository now also includes an active synthetic package at
\texttt{docs/examples/g-ensemble-evidence-packages/G.1-G.3+active-synthetic/}.
It supplies every required artifact role---external trace, metric output,
diagnostics, BISECT plan, and RCTX context---and is hash-validated by
\texttt{bisect-ensemble::evidence\_manifest}.  This proves the package contract
is executable, but the files are synthetic smoke data; real GerryChain/ReCom
traces remain required before citing the paper's state percentiles as empirical
findings.
