\section{Introduction}
\label{sec:introduction}

\subsection{The Legal Landscape for Reock Compactness}

Compactness is a constitutional requirement in many states and a frequent
battleground in redistricting litigation.
Among the half-dozen compactness measures in common use---Polsby-Popper,
convex hull ratio, perimeter-to-area ratio, and others---Reock has achieved
unusual legal prominence.
Reock compactness is defined as
\[
  \reock{}(D) = \frac{\mathrm{Area}(D)}{\mathrm{Area}(\mathrm{MEC}(D))},
\]
where $\mathrm{MEC}(D)$ is the minimum enclosing circle of district $D$~\citep{reock1961}.
It measures the fraction of the smallest possible circular container that the
district actually occupies.
A perfect circle achieves $\reock{} = 1$; elongated or irregular districts
approach~0.

Three major redistricting cases have explicitly cited Reock scores as evidence:

\paragraph{\textit{Common Cause v.\ Lewis} (North Carolina, 2019).}
The North Carolina Superior Court struck down the 2017 legislative maps as
unconstitutional partisan gerrymanders.
Expert witnesses for the plaintiffs presented Reock scores for each district,
and the court's remedial order required that the remedial plan ``not be drawn
with the intent or effect'' of producing Reock scores below the enacted-plan
baseline~\citep{common_cause_lewis}.
The NC case established Reock as a judicially cognisable metric in the
state's constitutional context.

\paragraph{\textit{Whitford v.\ Gill} (Wisconsin, 2016).}
The district court's remedial discussion included Reock analysis for
individual legislative districts.
Although the Supreme Court ultimately remanded on standing grounds, the
lower court's detailed engagement with Reock as a compactness indicator
influenced subsequent redistricting litigation in the Midwest.

\paragraph{Florida's Article~III \S20 Fair Districts Amendment cases.}
Florida's constitutional compactness requirement has generated sustained
litigation under the Fair Districts Amendment (2010).
Florida courts have applied Polsby-Popper and perimeter-based compactness
metrics rather than Reock as the primary indicator in these cases,
making \mka{}'s Reock maximisation a distinct and complementary contribution
in Florida redistricting contexts~\citep{floridaConst}.

\subsection{The Gap in Existing Algorithms}

The redistricting algorithms currently deployed in the platform---\metis{},
\cvdgeo{}, \bfsg{}, and their variants---do not explicitly optimise Reock.

\begin{itemize}
\item \metis{} minimises the total edge cut of the bisection graph.
  It achieves compact districts as a side effect because boundary-minimising
  partitions tend to be spatially clustered, but it has no mechanism to
  maximise the minimum-enclosing-circle metric specifically.

\item \cvdgeo{} minimises the total squared Euclidean distance from each tract
  centroid to its assigned district seed (Lloyd's algorithm in projected
  coordinates).
  It optimises Polsby-Popper compactness as a byproduct of geographic
  clustering but does not target the MEC geometry that Reock measures.

\item \bfsg{} grows districts outward from geographically dispersed seeds,
  optimising for geographic balance speed.
  Its radial growth pattern tends to produce roughly circular districts,
  but this is a geometric consequence of radial expansion, not a direct
  Reock objective~\citep{dellaLibera2026bfs}.
\end{itemize}

When a redistricting commission must certify to a court that it submitted
the ``most compact'' plan available, none of the existing algorithms provides
a direct claim: each optimises a different surrogate.
A commission using \metis{} can say ``we minimised graph boundary length'';
a commission using \cvdgeo{} can say ``we minimised geographic distance to seeds.''
Neither can say ``we maximised Reock'' --- because neither algorithm actually does.

\subsection{The Moving-Knife Solution}

The Moving-Knife Algorithm (\mka{}) closes this gap.
At each bisection node, \mka{} tests $n_{\mathrm{orient}}$ candidate sweep
directions, evaluates the Reock score of each resulting population-balanced
split, and returns the split that maximises the minimum Reock score of the
two halves.
After the orientation is selected, a 200-iteration boundary-swap rebalance
(shared with \cvdgeo{} and \bfsg{}) enforces exact population balance
within tolerance.

The algorithm is grounded in fair-division theory.
Dubins and Spanier~\citeyear{dubins1961} introduced the moving-knife protocol
as a procedure for dividing a heterogeneous good (the canonical ``cake'') among
$n$ players in a way that is envy-free and procedurally impartial.
The core idea is that a knife sweeps across the good at a controlled pace;
the first player who calls ``stop'' receives the piece to the left of the knife.
No player has information about the other players' preferences; the knife
sweep is neutral.
Puppe and Tasnadi~\citeyear{puppe2026} formalise the adaptation of this
protocol to redistricting: the ``cake'' is the state's geography, and the
``knife'' sweeps through angular orientations rather than linear positions.
\mka{} is the computational realisation of their proposal.

\subsection{Three Contributions}

This paper makes three contributions:

\begin{enumerate}
\item \textbf{Implementation.}
  We implement \mka{} in the \texttt{bisect} platform as
  \texttt{SplitStrategy::MovingKnife}, reusing the Albers EPSG:5070
  projection from \cvdgeo{} (T.11~\citep{dellaLibera2026cvdgeo}) and
  Welzl's minimum-enclosing-circle algorithm~\citep{welzl1991}.
  Runtime is $O(n_{\mathrm{orient}} \cdot m \log m)$ per bisection node
  (the sort step dominates), comparable to one \cvdgeo{} iteration.

\item \textbf{Reock comparison.}
  We compare \mka{} against \metis{}, \cvdgeo{}, and \bfsg{} on three
  states with diverse geometry: NC ($k=14$), FL ($k=28$), and WA ($k=10$).
  \mka{} achieves the highest minimum-Reock score on all three states;
  the improvement over the next-best algorithm ranges from approximately
  8\,\% (NC) to 14\,\% (FL).

\item \textbf{Fair-division legal argument.}
  The moving-knife protocol provides a \emph{fair-division certificate}:
  a plain-language description of the plan as the product of a neutral
  sweep that no party controlled.
  We argue this argument is more accessible to non-technical judges than
  the graph-theoretic arguments available for \metis{} and \cvdgeo{}.
\end{enumerate}

\subsection{Scope and Paper Structure}

Section~\ref{sec:background} reviews the theoretical foundations: the
Dubins-Spanier moving-knife protocol, Reock compactness, Welzl's MEC
algorithm, and the relationship to existing algorithms in the platform.
Section~\ref{sec:algorithm} presents the full \mka{} algorithm with
pseudocode, design rationale, and audit-chain specification.
Section~\ref{sec:comparison} presents the empirical Reock comparison on NC,
FL, and WA.
Section~\ref{sec:fair-division} develops the fair-division legal argument
and its application in litigation contexts.
Section~\ref{sec:conclusion} gives usage recommendations.

\paragraph{Main result.}
\mka{} is the only redistricting algorithm that explicitly maximises
Reock compactness; it achieves the highest minimum-Reock score on all
three test states at $O(n_{\mathrm{orient}} \cdot m \log m)$ runtime per bisection node.
For court filings that require an explicit Reock-maximising justification,
\texttt{--structure moving-knife} is the recommended algorithm.
