\section{Empirical Comparison}
\label{sec:comparison}

\subsection{Experimental Setup}

We compare four structure algorithms on three states representing the spectrum
from regular geometry to geometrically challenging:

\begin{itemize}
\item \textbf{NC} (North Carolina, $k=14$, $n \approx 2{,}660$ tracts):
  compact rectangular geometry, competitive partisan state.
  Reock scores are already moderate for METIS due to the regular shape.
  The FL panhandle geometry is absent, so the orientation sweep is expected
  to provide a modest but measurable gain.

\item \textbf{FL} (Florida, $k=28$, $n \approx 4{,}200$ tracts):
  the 200-mile panhandle extends westward from the main peninsula;
  coastal geometry around Tampa Bay, Miami, and the Keys creates elongated
  and concave subgraph regions.
  Gains from angle-optimised splits are expected to be largest here,
  because naive horizontal or vertical sweeps miss the locally optimal
  orientation for panhandle subgraphs.

\item \textbf{WA} (Washington, $k=10$, $n \approx 1{,}400$ tracts):
  Puget Sound creates a geographic discontinuity between the westside
  (Kitsap Peninsula, Olympic Peninsula) and the eastside (Cascade foothills,
  eastern plains).
  Intermediate geometry between NC and FL.
\end{itemize}

All runs use 2020 Census data, geographic boundary-length edge weights
(B.2 default), standard bisection tree, and base seed $s = 42$.
Single-run results are marked with $\dagger$ (not ensemble-averaged); the
purpose of this table is to illustrate \emph{relative ordering} of Reock
performance, not to report production-quality compactness figures.
Production figures from \textit{G.15 Comprehensive Comparison}~\citep{dellaLibera2026comparison}
should be used for plan certification.

\paragraph{Algorithms.}
\begin{itemize}
\item \textbf{\metis{}}: standard METIS graph partitioning,
  \texttt{--structure standard-bisect} (the platform default).
\item \textbf{\cvdgeo{}}: Geographic CVD Phase~2,
  \texttt{--structure centroidal-voronoi --cvd-metric geographic},
  20 iterations~\citep{dellaLibera2026cvdgeo}.
\item \textbf{\bfsg{}}: BFS Growth,
  \texttt{--structure bfs-growth}~\citep{dellaLibera2026bfs}.
\item \textbf{\mka{}}: Moving-Knife Algorithm,
  \texttt{--structure moving-knife --mka-orientations 180 --mka-metric reock}.
\end{itemize}

\paragraph{Metrics.}
\begin{itemize}
\item \textbf{Min-Reock}: minimum Reock score across all $k$ districts;
  this is the quantity \mka{} directly maximises.
  Higher is better.
\item \textbf{Mean-Reock}: mean Reock score across all $k$ districts.
  Higher is better.
\item \textbf{EC}: total weighted edge cut of the final plan.
  Lower is better.
\item \textbf{Time}: wall-clock time in seconds on a reference workstation
  (8 cores, \texttt{--workers 8}).
\end{itemize}

\subsection{Results}

Table~\ref{tab:reock-comparison} presents the Reock comparison.

\begin{table}[t]
\centering
\caption{%
  Reock compactness and edge-cut comparison on NC ($k=14$), FL ($k=28$),
  and WA ($k=10$), 2020 census, base seed $s=42$.
  $\dagger$: single run; not ensemble-averaged.
  Min-Reock: minimum Reock across all $k$ districts (higher is better).
  EC: total edge cut (lower is better).
  Bold: best Min-Reock per state.
}
\label{tab:reock-comparison}
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{@{}llrrrr@{}}
\toprule
State & Algorithm & Min-Reock$^\dagger$ & Mean-Reock$^\dagger$ & EC$^\dagger$ & Time (s) \\
\midrule
\multirow{4}{*}{NC ($k=14$)}
 & \metis{}   & 0.187 & 0.241 & 4{,}821 &  4.2 \\
 & \cvdgeo{}  & 0.219 & 0.268 & 5{,}512 &  5.9 \\
 & \bfsg{}    & 0.224 & 0.271 & 6{,}104 &  3.8 \\
 & \textbf{\mka{}} & \textbf{0.243} & \textbf{0.279} & 6{,}318 & 6.1 \\
\midrule
\multirow{4}{*}{FL ($k=28$)}
 & \metis{}   & 0.152 & 0.201 & 9{,}247 &  9.1 \\
 & \cvdgeo{}  & 0.174 & 0.228 & 10{,}219 & 10.7 \\
 & \bfsg{}    & 0.181 & 0.233 & 11{,}037 &  7.4 \\
 & \textbf{\mka{}} & \textbf{0.206} & \textbf{0.251} & 11{,}482 & 11.3 \\
\midrule
\multirow{4}{*}{WA ($k=10$)}
 & \metis{}   & 0.171 & 0.219 & 3{,}614 &  3.3 \\
 & \cvdgeo{}  & 0.198 & 0.247 & 3{,}971 &  4.0 \\
 & \bfsg{}    & 0.203 & 0.251 & 4{,}318 &  2.9 \\
 & \textbf{\mka{}} & \textbf{0.227} & \textbf{0.263} & 4{,}591 &  4.3 \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Analysis}

\paragraph{Min-Reock improvement.}
\mka{} achieves the highest minimum Reock score on all three states.
Over the next-best algorithm (\bfsg{} in every case), the improvement is:
NC: $+0.019$ ($+8.5\,\%$);
FL: $+0.025$ ($+13.8\,\%$);
WA: $+0.024$ ($+11.8\,\%$).
The improvement is largest on FL, consistent with the hypothesis that the
panhandle geometry provides large gains when the optimal orientation is found
rather than defaulting to a fixed radial growth or graph-topology partition.
NC shows the smallest gain, because the regular geometry of the Piedmont plateau
means nearly all orientations produce similar splits --- the sweep finds a
modest improvement but the feasible range is narrower.

\paragraph{Mean-Reock.}
Mean Reock follows the same ordering.
The mean--min gap is approximately 0.03--0.05 for all algorithms, indicating
that Reock is reasonably uniform across districts for all methods; no algorithm
produces one extraordinarily compact district at the expense of others.

\paragraph{Edge cut.}
\mka{} has the highest edge cut on all three states, as expected: it does
not minimise edge cuts (that is \metis{}'s objective) and its angular sweep
does not align with the graph-topology boundaries that METIS exploits.
The excess EC over \metis{} is 31\,\% on NC, 24\,\% on FL, and 27\,\% on WA.
The excess over \cvdgeo{} is 15\,\% on NC, 12\,\% on FL, and 16\,\% on WA.
These are acceptable for most redistricting contexts: EC is a secondary
criterion when the primary mandate is Reock-based compactness; and the 10--15\,\%
EC premium over \cvdgeo{} translates to modestly less boundary-following districts,
not to grossly non-contiguous ones.

\paragraph{Comparison with prior geometric algorithms.}
\cvdgeo{} improves Polsby-Popper over \metis{} by approximately 8--13\,\%
(T.11~\citep{dellaLibera2026cvdgeo}).
\mka{} improves Min-Reock over \metis{} by approximately 30\,\% on NC,
36\,\% on FL, and 33\,\% on WA --- substantially larger gains, because
the sweep directly targets the Reock geometry.
Each algorithm is best-in-class for its own objective.

\subsection{Orientation Sensitivity}

To characterise how strongly performance depends on orientation granularity,
Table~\ref{tab:orientation-sensitivity} compares $n_{\mathrm{orient}} \in \{36, 90, 180\}$
on NC ($k=14$).

\begin{table}[h]
\centering
\caption{%
  \mka{} orientation granularity sensitivity: NC ($k=14$), 2020, seed $s=42$.
  $\dagger$: single run.
  Diminishing returns beyond 90 orientations on NC.
}
\label{tab:orientation-sensitivity}
\renewcommand{\arraystretch}{1.25}
\begin{tabular}{@{}rrrr@{}}
\toprule
$n_{\mathrm{orient}}$ & Min-Reock$^\dagger$ & EC$^\dagger$ & Time (s) \\
\midrule
 36 & 0.237 & 6{,}291 & 1.4 \\
 90 & 0.241 & 6{,}309 & 3.2 \\
180 & 0.243 & 6{,}318 & 6.1 \\
\bottomrule
\end{tabular}
\end{table}

The 36-orientation sweep captures approximately 97.5\,\% of the 180-orientation
Min-Reock at 23\,\% of the runtime.
The default $n_{\mathrm{orient}} = 180$ is recommended for production use to ensure the
globally best direction is not missed by more than $0.5^\circ$.
For rapid exploratory runs, \texttt{--mka-orientations 36} reduces runtime
to approximately 1.4\,s on NC.
Phase~2 will benchmark 36 vs.\ 180 on FL and WA to confirm whether the
5$\times$ speedup justifies the small quality loss in panhandle geometries.

\subsection{Partisan Neutrality Check}

Table~\ref{tab:partisan} verifies that the Reock improvement does not
introduce systematic partisan bias.

\begin{table}[h]
\centering
\begin{tabular}{@{}lrrrr@{}}
\toprule
State & $k$ & \metis{} D-seats & \cvdgeo{} D-seats & \mka{} D-seats \\
\midrule
NC & 14 & 7 & 7 & 7 \\
FL & 28 & 12 & 12 & 12 \\
WA & 10 & 6 & 6 & 6 \\
\bottomrule
\end{tabular}
\caption{%
  2020 presidential two-party Democratic seat share under \metis{}, \cvdgeo{},
  and \mka{} (seed $s=42$).
  $\dagger$: single run.
  The angular sweep selects the most compact orientation; it does not
  select for or against either party's seat share.
}
\label{tab:partisan}
\end{table}

Partisan outcomes are identical across all three algorithms on all three
states at seed $s=42$.
The \mka{} angular sweep is parametrised by geographic coordinates and
population balance; it contains no electoral data and has no mechanism
to advantage either party.
The partisan-neutrality result is consistent with the fair-division
theoretical justification (Section~\ref{sec:fair-division}).
Ensemble verification across 5 independent seeds confirms stable D-seat outcomes; we defer to the G.15 comprehensive comparison for multi-seed partisan analysis.
