\section{Reconciliation Contract}

The full CVR-to-summary contract has three layers:

\begin{center}
\small
\begin{tabularx}{\textwidth}{lX}
\toprule
Layer & Required reconciliation \\
\midrule
Manifest to CVR rows & Ballot-card or CVR-row counts match ballot manifest accounting for the represented batch or reporting unit. \\
CVR rows to contest aggregates & Candidate selections, write-in buckets, undervotes, overvotes, and blank contests aggregate by contest. \\
Contest aggregates to summaries & Aggregated CVR interpretations match published summaries by contest, reporting unit, status, and optional batch. \\
\bottomrule
\end{tabularx}
\end{center}

\begin{align*}
\texttt{manifest card count}
  &= \texttt{CVR row/card count} \\
  &= \texttt{summary counted basis}
\end{align*}

\[
\sum \texttt{CVR contest interpretations}
  = \texttt{summary selections and residuals}
\]

The word CVR must stay precise. A CVR row is not necessarily a voter or a whole
ballot; some systems export one row per ballot card, one row per sheet side, or
one row per contest grouping. A CVR can record an undervote, overvote, blank
contest, adjudicated write-in bucket, or contest not on that ballot style. V.6
therefore treats the adapter's row semantics as part of the evidence contract,
not as an implementation detail.

The first normalized RCOUNT row uses this package-local contract:

\begin{center}
\small
\begin{tabularx}{\textwidth}{lX}
\toprule
Field & Meaning \\
\midrule
\texttt{cvr\_id} & Package-local CVR/card identifier; not a voter identifier. \\
\texttt{contest\_id} & Contest interpreted by this row. \\
\texttt{reporting\_unit\_id} & Public reporting unit for aggregation. \\
\texttt{batch\_id} & Optional batch link when manifest evidence is batch-level. \\
\texttt{status} & Count status represented by the row. \\
\texttt{selection\_ids} & Candidate or write-in bucket selections interpreted for the contest. \\
\texttt{undervote}, \texttt{overvote}, \texttt{blank\_contest} & Residual contest interpretation flags. \\
\texttt{source\_refs} & Source evidence rows supporting the normalized CVR row. \\
\bottomrule
\end{tabularx}
\end{center}

The implemented verifier treats the row as one contest interpretation. Each
CVR id and contest id pair must be unique. A row fails if it combines a
candidate selection with a residual flag, sets multiple residual flags,
references an unknown selection id, or contains more selections than the
contest's \texttt{vote\_for} rule permits.

\begin{center}
\small
\begin{tabularx}{\textwidth}{lX}
\toprule
Check & Failure caught \\
\midrule
\texttt{cvr\_summary\_total} & CVR aggregates disagree with public summary selections or residuals. \\
\texttt{contest\_selection\_sum} & Summary selections plus residuals do not equal counted ballots. \\
\texttt{jurisdiction\_contest\_total} & Lower-level summaries do not roll up to the jurisdiction total. \\
\texttt{source\_hash\_match} & Published source evidence bytes no longer match declared hashes. \\
\bottomrule
\end{tabularx}
\end{center}
