\section{Fixtures}

The positive fixture is \texttt{cvr-summary}. It contains 140 normalized CVR
contest rows: 80 rows for P-001 and 60 rows for P-002. For precinct P-001, the
public summary declares 80 counted ballots:

\begin{center}
\small
\begin{tabularx}{\textwidth}{lrrrrrr}
\toprule
Unit & A & B & Write-in & Under & Over & Blank \\
\midrule
P-001 & 40 & 35 & 1 & 3 & 1 & 0 \\
P-002 & 25 & 30 & 0 & 4 & 0 & 1 \\
\bottomrule
\end{tabularx}
\end{center}

For P-001, the CVR aggregate is \(40 + 35 + 1 + 3 + 1 + 0 = 80\), so
\texttt{cvr\_summary\_total} and \texttt{contest\_selection\_sum} both pass.

The negative fixture is \texttt{bad-cvr-summary}. It changes one P-001 CVR row
from \texttt{cand-a} to \texttt{cand-b} while leaving the public summaries
unchanged. The summary arithmetic still passes, but the CVR aggregate no longer
matches the summary: Candidate A computes to 39 instead of 40.

\begin{center}
\small
\begin{tabularx}{\textwidth}{lX}
\toprule
Path & Role \\
\midrule
\path{normalized/cvr.ndjson} & Normalized CVR contest interpretation rows. \\
\path{normalized/summaries.ndjson} & Public summary targets for CVR aggregation. \\
\path{normalized/contests.ndjson} & Selection ids and \texttt{vote\_for} constraints. \\
\path{sources/source-index.json} & Raw source hash evidence. \\
\path{transcripts/verify-transcript.json} & Verifier pass/fail transcript. \\
\bottomrule
\end{tabularx}
\end{center}

The regression caught by \texttt{bad-cvr-summary} is precise: a verifier that
checks only summary arithmetic, or silently ignores \path{normalized/cvr.ndjson},
would miss a CVR-side selection drift.
