\section{Manifest Model}

The current RCOUNT batch manifest row is intentionally small:

\begin{center}
\small
\begin{tabularx}{\textwidth}{lX}
\toprule
Field & Meaning \\
\midrule
\texttt{batch\_id} & Stable package identifier for the batch. \\
\texttt{reporting\_unit\_id} & Public unit associated with the batch, such as a precinct or central-count unit. \\
\texttt{kind} & Operational class, such as election-day, mail, provisional, vote-center, or central-count. \\
\texttt{status} & Publication status represented in the package. \\
\texttt{accepted\_ballots} & Ballots accepted into the batch's accounting path. \\
\texttt{counted\_ballots} & Ballots counted from that batch for the represented contest summaries. \\
\texttt{rejected\_ballots} & Accepted-accounting records later rejected or removed from counting. \\
\texttt{source\_refs} & Source evidence rows that support the manifest row. \\
\bottomrule
\end{tabularx}
\end{center}

\begin{center}
\small
\begin{tabularx}{\textwidth}{lX}
\toprule
Object & Boundary \\
\midrule
Voter & Eligibility and participation records are not public batch manifest rows. \\
Ballot or ballot card & Physical or electronic voting artifacts may be represented by manifests, but are not identical to batch rows. \\
CVR row & A cast-vote record records contest interpretation; V.6 reconciles CVRs to summaries. \\
Contest summary & A contest summary can reference a batch, but it is not the manifest itself. \\
Selection or vote & Candidate choices, write-ins, overvotes, undervotes, and blanks are contest semantics, not batch identity. \\
\bottomrule
\end{tabularx}
\end{center}

The package contract is:
\begin{itemize}
\item normalized batch manifest rows live at \path{normalized/batches.ndjson};
\item summaries that claim batch-level support carry a \texttt{batch\_id};
\item \texttt{batch\_id} is stable inside the package and may be mapped from
  vendor ids, scanner labels, or election-office labels by an adapter;
\item \texttt{source\_refs} bind the normalized row back to hashed source
  evidence in \path{sources/source-index.json}.
\end{itemize}

Vendor export adapters and external schema mapping belong to V.9. V.5 only
defines the normalized RCOUNT contract those adapters must satisfy.
