\section{Boundaries}

Ballot manifest verification does not certify an election. It verifies a public
accounting relationship inside the package. Officials, canvassing boards,
recount procedures, and courts still decide eligibility, acceptance, cure,
duplication, adjudication, recount changes, and certification.

\begin{center}
\small
\begin{tabularx}{\textwidth}{lX}
\toprule
Workflow event & Manifest implication \\
\midrule
Late mail accepted & Add or update a mail batch with accepted and counted ballots. \\
Provisional review & Add, reject, or reclassify provisional batch evidence after eligibility decisions. \\
Ballot cure & Move records into accepted accounting only after the cure decision. \\
Duplicated ballot resolution & Link counted batch evidence to duplication/adjudication source records. \\
Recount or court order & Publish amended batch and summary records with event lineage. \\
\bottomrule
\end{tabularx}
\end{center}

A missing batch is therefore a verification failure, not a standalone fraud
finding. Investigation determines whether the cause is an export defect,
publication lag, data-entry omission, lawful amendment not yet represented, or
misconduct.

The manifest layer also does not prove voter intent or tabulator correctness. A
batch may reconcile numerically while a later CVR or audit layer still finds a
parser, contest-definition, or hand-audit problem. RCOUNT should report those
as separate failures instead of collapsing them into one generic claim.

Finally, batch granularity has privacy implications. Very small batches,
single-voter provisional batches, rare write-in batches, or precise batch times
can create small-cell disclosure risk. Publication policy must be reviewed with
V.4's receipt-safety rule in mind.

\begin{center}
\small
\begin{tabularx}{\textwidth}{lX}
\toprule
Publication pattern & Privacy risk \\
\midrule
Single-voter provisional batch & Can reveal whether one person's ballot was accepted or rejected. \\
Tiny cure batch & Can expose timing and status of a small group of voters. \\
Rare write-in or contest batch & Can narrow candidate-choice inference in a small cell. \\
Precise scanner time plus small unit & Can combine with observation or check-in records. \\
\bottomrule
\end{tabularx}
\end{center}

Jurisdictions may need public/restricted manifest views, aggregation, redaction,
or delayed publication. Hashes and source references provide tamper evidence;
they do not by themselves provide privacy.
