\section{Background and the Symmetry Collapse}
\label{sec:background}

\subsection{Standard ReCom}

Let $G = (V, E)$ be the census-tract adjacency graph for a state, and let
$\pi : V \to [k]$ be a valid redistricting plan (each district connected
and population-balanced within tolerance $\epsilon$).
A single \recom{} step proceeds as follows~\citep{deford2021}:

\begin{enumerate}
  \item Sample an adjacent pair of districts $(d_i, d_j)$ uniformly at
        random from all pairs sharing at least one boundary edge.
  \item Form the merged region $R = \pi^{-1}(d_i) \cup \pi^{-1}(d_j)$,
        with the induced subgraph $G[R]$.
  \item Sample a uniformly random spanning tree $T$ of $G[R]$ via
        Wilson's algorithm~\citep{wilson1996}.
  \item Enumerate the balanced cuts of $T$ --- edges whose removal partitions
        $T$ into two components, each satisfying the population balance
        constraint.
  \item If at least one balanced cut exists, sample one uniformly and
        accept the resulting plan $\pi'$.
        Otherwise, reject and repeat from step~1 (pair reselection).
\end{enumerate}

Step~3 is the dominant cost: Wilson's algorithm runs in $O(|R| \log |R|)$
expected time for a graph with $|R|$ vertices.

\subsection{Why the Acceptance Ratio Is Always 1}

The Metropolis-Hastings acceptance ratio for the move $\pi \to \pi'$ is:
\[
  \alpha(\pi, \pi') = \min\!\left(1,\;
    \frac{\pi^*(\pi')\, Q(\pi', \pi)}{\pi^*(\pi)\, Q(\pi, \pi')}
  \right),
\]
where $\pi^*$ is the target distribution and $Q(\pi, \pi')$ is the proposal
probability.
If we target the uniform distribution $\pi^*(\pi) = \pi^*(\pi')$,
the ratio simplifies to $Q(\pi', \pi) / Q(\pi, \pi')$.

The proposal probability for standard \recom{} is:
\begin{align}
  Q(\pi, \pi') &= \frac{1}{|\mathcal{P}(\pi)|}
    \cdot \frac{1}{|\mathcal{T}(R)|}
    \cdot \frac{1}{|\mathcal{C}(T, R)|},
\end{align}
where $\mathcal{P}(\pi)$ is the set of adjacent pairs of districts in $\pi$,
$\mathcal{T}(R)$ is the set of spanning trees of $G[R]$, and
$\mathcal{C}(T, R)$ is the set of balanced cuts of the sampled tree $T$.
The reverse proposal $Q(\pi', \pi)$ has the same structure, with the
same merged region $R$ (since $\pi$ and $\pi'$ differ only in how
$R$ is partitioned).

In standard \recom{}, both the forward and reverse proposals sample a spanning
tree from the same graph $G[R]$ with the same uniform measure.
Consequently:
\[
  \frac{Q(\pi', \pi)}{Q(\pi, \pi')}
  = \frac{|\mathcal{P}(\pi)| \cdot |\mathcal{C}(T', R)|}{|\mathcal{P}(\pi')|
    \cdot |\mathcal{C}(T, R)|}.
\]
Here $T$ is the spanning tree used to propose $\pi'$, and $T'$ is the
(hypothetical) spanning tree that would propose $\pi$ from $\pi'$.
Standard \recom{} does \emph{not} sample $T'$---it only ever samples $T$.
Without an independent sample of $T'$, the ratio $|\mathcal{C}(T', R)|$ is
unknown, and the ratio $Q(\pi', \pi) / Q(\pi, \pi')$ is not computed.
The implementation effectively sets $\alpha = 1$ (always accept if a balanced
cut exists), which is equivalent to using an unknown acceptance ratio.

The stationary distribution of the resulting chain is not the uniform
distribution; it is a distribution that weights plans by the implicit bias
of the spanning tree measure and the number of balanced cuts of the typical
tree in the region.

\subsection{Breaking the Symmetry via Independent Reverse Tree}

Merge-Split~\citep{janson2022} restores computable Metropolis-Hastings
ratios by sampling $T'$---the reverse spanning tree---independently:

\begin{enumerate}
  \item After sampling $T_{\mathrm{fwd}}$ and selecting the proposed plan
        $\pi'$, sample a second uniformly random spanning tree
        $T_{\mathrm{rev}}$ of the same merged region $G[R]$, using an
        independent random source.
  \item Count $\mathrm{cuts\_fwd} = |\mathcal{C}(T_{\mathrm{fwd}}, R)|$
        and $\mathrm{cuts\_rev} = |\mathcal{C}(T_{\mathrm{rev}}, R)|$.
  \item Use $\min(1,\mathrm{cuts\_fwd} / \mathrm{cuts\_rev})$ as the
        acceptance probability when $\mathrm{cuts\_rev}>0$, and reject when
        the reverse tree has no balanced cuts.
\end{enumerate}

The key insight is that the reverse sample exposes whether the proposed split
would be easy or hard to regenerate from the proposed state.
Since both trees are sampled uniformly from the same merged region, the ratio
is a one-sample stochastic correction for the balanced-cut-count asymmetry
between the forward and reverse moves.
It is not a determinant computation and it is not the full finite-chain
Metropolis-Hastings ratio when district-pair counts differ.

\begin{remark}[Exact vs.\ approximate reversibility]
\label{rem:approx}
The \mergesplit{} ratio is a random, per-step correction, not the exact
matrix-tree ratio used by determinant-style \forestrecom{}
variants~\citep{mccartan2020,dellaLibera2026forestRecom}.
The chain still has a stationary distribution on the finite valid-plan graph
under the usual irreducibility and laziness conditions, but this paper does not
claim that the implemented chain exactly samples uniformly from all valid
plans.
Its audit claim is narrower and useful: the balanced-cut-count correction,
reverse zero-cut rejection, seed schedule, and accepted-state selection are
declared and reproducible.
\end{remark}
