\documentclass[12pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{amsmath, amssymb, amsthm}
\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}
\newtheorem{proposition}{Proposition}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{remark}{Remark}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage{booktabs}
\usepackage{xcolor}
\usepackage{multirow}
\usepackage{array}
\usepackage{microtype}
\usepackage{setspace}
\usepackage{titlesec}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{lmodern}
\usepackage[T1]{fontenc}

%% SIAM-adjacent style: numbered sections, upright headings
\titleformat{\section}{\normalfont\large\bfseries}{\thesection.}{0.5em}{}
\titleformat{\subsection}{\normalfont\normalsize\bfseries}{\thesubsection.}{0.5em}{}
\titleformat{\subsubsection}{\normalfont\normalsize\itshape}{\thesubsubsection.}{0.5em}{}

%% Spacing
\setstretch{1.4}

%% Algorithm macros
\newcommand{\cvd}{\textsc{CVD}}
\newcommand{\metis}{\textsc{Metis}}
\newcommand{\sa}{\textsc{SimulatedAnnealing}}
\newcommand{\ec}{\ensuremath{\mathrm{EC}}}
\newcommand{\pp}{\ensuremath{\mathrm{PP}}}
\newcommand{\be}{\textsc{BisectionEnsemble}}
\newcommand{\ar}{\textsc{ApportionRegions}}
\newcommand{\geosec}{\textsc{GeoSection}}
\newcommand{\areasec}{\textsc{AreaSection}}
\newcommand{\recom}{\textsc{ReCom}}
\newcommand{\cs}{\textsc{ConvergenceSweep}}
\newcommand{\bfsgrowth}{\textsc{BfsGrowth}}
\newcommand{\dia}{\textsc{Districting Integrity Act}}

%% Math shorthands
\newcommand{\ECnorm}{\mathrm{EC}_{\mathrm{norm}}}
\newcommand{\EC}{\mathrm{EC}}

%% -----------------------------------------------------------------------

\title{%
  \textbf{Centroidal Voronoi Districts: Geometric Packing via}\\
  \textbf{Graph-Medoid Iteration}\\[0.8em]
  \normalsize\textit{T.10 --- Algorithm Design / Geometric Structure}%
}

\author{%
  Gio Della-Libera\\
  \textit{Apportionment Research Group}\\
  \small\texttt{giodl@microsoft.com}
}

\date{May 2026}

\begin{document}

\maketitle

\begin{abstract}
\noindent
We introduce Centroidal Voronoi Districts (\cvd{}), a Structure-layer
algorithm for redistricting that constructs each bisection node by proximity
packing rather than edge-cut minimisation.
The implemented graph-distance mode assigns each census tract to its nearest
seed by BFS hop count and iteratively moves seeds to graph-distance medoids.
The implemented geographic mode instead uses Albers-projected tract centroids
and updates seeds through population-weighted Euclidean means snapped back to
real tracts.
\cvd{} is the discrete redistricting analogue of Lloyd's algorithm for
optimal quantisation~\citep{lloyd1982}, but with redistricting-specific
constraints: finite tract units, population balance, contiguity repair, and
deterministic seed derivation.

Two seeds are placed using farthest-first selection: the first seed is chosen
from a SHA-256-derived per-node seed; the second is the tract farthest from the
first under the selected metric.
This initialisation shows the algorithm's purpose visually: one seed grabs one
end of the state, the other grabs the opposite end, and the boundary moves
until the two proximity regions stabilise.
The current implementation is a binary splitter inside the standard
\texttt{BisectionTree}; it rebalances each split toward a half-population
target, so odd-seat ratio-aware targeting remains an implementation/evidence
ceiling for publication claims.

Empirically, draft \cvd{} runs on North Carolina ($k=14$) show
Polsby-Popper compactness competitive with \metis{} despite not optimising
edge cuts, but the numerical comparison tables in this paper should be read
as benchmark targets unless paired with archived RPLAN/RCTX run packages.
\cvd{} has a simple verifier story: given the input graph or centroids, node
path, base seed, metric, iteration budget, and rebalance rule, an auditor can
replay the split tree.
We compare \cvd{} with \metis{} and \sa{} on compactness, edge cut, runtime,
and evidence requirements.
\end{abstract}

\tableofcontents
\newpage

\input{sections/01-introduction}
\input{sections/02-background}
\input{sections/03-algorithm}
\input{sections/04-comparison}
\input{sections/05-phase2}
\input{sections/06-conclusion}

\bibliographystyle{plainnat}
\bibliography{references}

\end{document}
