\section{Benchmark Targets and Evidence Status}
\label{sec:results}

\subsection{What the Current Evidence Proves}

The implemented code proves that geographic CVD is a real BISECT split mode,
not just a paper design. The L0/L1 surface covers:

\begin{itemize}
\item parsing \texttt{--cvd-metric geographic};
\item the distinct \texttt{"CVD\_GEO\_INIT\_"} seed domain;
\item successful k=2 geographic CVD splits on synthetic centroid grids;
\item deterministic output for the same base seed and node path;
\item a hard error when geographic CVD is requested without centroid data; and
\item the Albers central-meridian invariant used by the projection helper.
\end{itemize}

That evidence is sufficient for an implementation note. It is not sufficient
for a final empirical claim such as ``geographic CVD improves Florida by
6 percent.'' Those claims require archived input packages, seed manifests,
run directories, output hashes, and same-metric analysis artifacts.

\subsection{Publication Benchmark Plan}

Table~\ref{tab:main-results} is therefore a benchmark target table. Its rows
state what the publication package should measure once NC/FL/WA centroid
packages are archived and reproducible. Values are intentionally marked
\texttt{TBD}; a future evidence pass should replace them with run-derived
numbers and cite the package hashes.

\begin{table}[t]
\centering
\caption{%
  Phase~1 vs.\ geographic CVD benchmark matrix. Values are targets for an
  archived reproduction package, not current paper claims.
}
\label{tab:main-results}
\renewcommand{\arraystretch}{1.25}
\begin{tabular}{@{}llrrrrr@{}}
\toprule
State & Algorithm & PP & EC & Pop dev. & Iters & Package \\
\midrule
NC ($k=14$) & Graph-distance CVD & TBD & TBD & TBD & TBD & pending \\
NC ($k=14$) & Geographic CVD     & TBD & TBD & TBD & TBD & pending \\
\midrule
FL ($k=28$) & Graph-distance CVD & TBD & TBD & TBD & TBD & pending \\
FL ($k=28$) & Geographic CVD     & TBD & TBD & TBD & TBD & pending \\
\midrule
WA ($k=10$) & Graph-distance CVD & TBD & TBD & TBD & TBD & pending \\
WA ($k=10$) & Geographic CVD     & TBD & TBD & TBD & TBD & pending \\
\bottomrule
\end{tabular}
\end{table}

\subsection{What to Compare}

The benchmark should compare three effects separately.

\paragraph{Metric effect.}
Run graph-distance CVD and geographic CVD with the same base seeds, the same
bisection tree, the same population tolerance, and the same centroid package.
Only the distance metric should differ.

\paragraph{Geometry effect.}
Use state families where graph distance and geographic distance are expected
to diverge differently: a regular inland state such as North Carolina, a
panhandle/coastal state such as Florida, and a water-barrier state such as
Washington. The claim should be visual as well as numeric: show the seed
frontier on at least one node where the two metrics split the tract grid
differently.

\paragraph{Odd-seat effect.}
For $k=7$, $k=13$, $k=27$, and similar odd-seat trees, record whether the local
CVD split is being forced toward half-population rather than the tree's
$k_L:k_R$ child ratio. Until ratio-aware CVD splits are implemented, this
effect should be disclosed beside any statewide compactness numbers.

\subsection{Failure Modes to Archive}

A useful geographic CVD package should include failures and boundary cases, not
only successful runs:

\begin{itemize}
\item missing centroid companion file, expected configuration error;
\item malformed centroid JSON, expected load error;
\item centroid length mismatch, expected load error before split execution;
\item synthetic grid with identical centroids, expected deterministic tie
  behavior; and
\item real state with island or water-barrier geometry, expected visible
  metric divergence.
\end{itemize}

These cases are the bridge between an attractive geometric algorithm and an
auditable election-administration tool: users need to know not just when the
metric works, but how it fails.
