\label{sec:algorithm}

\subsection{Commuting Destination Distribution}

For each home census tract $h$, the commuting destination distribution
is a sparse vector:
\begin{equation}
  D(h) = \bigl\{ w : F[h, w] > 0 \bigr\}
  \quad\text{with values } D(h)[w] = F[h, w]
  \label{eq:distribution}
\end{equation}
where $F[h, w]$ is the number of workers residing in tract $h$ and employed
in tract $w$, aggregated from the LODES OD block-level records as described
in Section~\ref{sec:background}.

The distribution is stored in unnormalized count form.
Normalization to a probability distribution (dividing by $\sum_w D(h)[w]$)
is performed at similarity computation time, not at storage time,
to avoid floating-point precision issues in the sparse representation.

\subsection{Weighted Jaccard Similarity}

The community similarity between two home tracts $h_1$ and $h_2$ is
the weighted Jaccard similarity of their commuting destination distributions:
\begin{equation}
  \mathrm{sim}(h_1, h_2)
  = \frac{\sum_{w} \min\!\bigl(D(h_1)[w],\, D(h_2)[w]\bigr)}
         {\sum_{w} \max\!\bigl(D(h_1)[w],\, D(h_2)[w]\bigr)}
  \label{eq:jaccard}
\end{equation}
where the sums run over all work tracts $w$ that appear in either $D(h_1)$
or $D(h_2)$.
The $\min$ and $\max$ functions operate element-wise: for $w$ present in
$D(h_1)$ but not $D(h_2)$, $\min(\cdot) = 0$ and $\max(\cdot) = D(h_1)[w]$.

\paragraph{Properties of weighted Jaccard similarity.}
\begin{itemize}
  \item \textbf{Bounded.} $\mathrm{sim}(h_1, h_2) \in [0, 1]$ for all pairs.
  \item \textbf{Symmetric.} $\mathrm{sim}(h_1, h_2) = \mathrm{sim}(h_2, h_1)$
    because $\min$ and $\max$ are symmetric.
  \item \textbf{Self-similarity.} $\mathrm{sim}(h, h) = 1.0$ for any tract with
    at least one commuting flow.
  \item \textbf{Disjoint zero.} $\mathrm{sim}(h_1, h_2) = 0.0$ when $D(h_1)$ and
    $D(h_2)$ have no destination tract in common ($\sum \min = 0$).
\end{itemize}

\paragraph{Interpretation.}
The weighted Jaccard similarity measures the overlap of commuting flows
rather than binary destination co-membership.
If $h_1$ sends 100 workers to downtown Milwaukee and 20 to Brookfield,
and $h_2$ sends 80 workers to downtown Milwaukee and 60 to Brookfield:
\begin{align}
  \sum_w \min &= \min(100, 80) + \min(20, 60) = 80 + 20 = 100 \\
  \sum_w \max &= \max(100, 80) + \max(20, 60) = 100 + 60 = 160 \\
  \mathrm{sim}(h_1, h_2) &= 100 / 160 = 0.625
\end{align}
This is a meaningful community bond: both tracts primarily commute to the same
Milwaukee employment areas, even though the proportions differ.

If the same $h_1$ is compared to a tract $h_3$ that sends workers to
Port of Milwaukee logistics hubs rather than downtown Milwaukee:
$\mathrm{sim}(h_1, h_3) \approx 0$ because the destination sets are
disjoint.
The algorithm correctly identifies the community boundary.

\subsection{Computational Complexity}

For a home tract $h$, the commuting distribution $D(h)$ has at most $W$
non-zero entries, where $W$ is the number of distinct work tracts
receiving commute flows from $h$.
Empirically, $W \approx 15$--$30$ per home tract based on LODES OD
density analysis for mid-sized states.

The weighted Jaccard for a pair $(h_1, h_2)$ requires iterating over the
union of their destination sets, which has at most $W_1 + W_2$ elements.
For all $n$ adjacent tract pairs in a state graph (approximately $4n$--$6n$
edges for typical tract adjacency graphs), the total computation is:
\begin{equation}
  O\!\bigl(n \times (W_1 + W_2)\bigr) \approx O(n \times 2 \bar{W})
\end{equation}
For NC ($n = 2{,}655$ tracts, approximately 7{,}000 adjacency edges,
$\bar{W} \approx 20$):
\begin{equation}
  \text{operations} \approx 7000 \times 40 = 280{,}000
\end{equation}
This is trivially fast; the dominant cost is the initial block-to-tract
aggregation and sparse matrix construction.
The Phase~2 performance target of under 60 seconds wall time for NC on
8 workers is expected to be achieved comfortably.

\subsection{Sparsity Handling}

\paragraph{Tracts with no OD data.}
Some census tracts have no OD records in the LODES file.
This occurs for:
\begin{itemize}
  \item Very low-population tracts (fewer than 3 workers to protect privacy;
    LODES suppresses records with cell counts below 3).
  \item Institutional population tracts: prisons, dormitories, military
    barracks where residents are present but UI-covered commuting is not
    recorded.
  \item Tracts that are entirely non-residential (airports, industrial parks,
    commercial cores with no residential population).
\end{itemize}
For pairs where one or both tracts have empty distributions, the weighted
Jaccard is undefined (denominator is zero).
The implementation handles this with a neutral default:
\begin{equation}
  \mathrm{sim}(h_1, h_2) = 0.0 \quad \text{if } D(h_1) = \emptyset \text{ or } D(h_2) = \emptyset
  \label{eq:sparsity}
\end{equation}
A similarity of 0.0 corresponds to a neutral edge: the M.4 weight falls to
$\alpha \times w_{\mathrm{geo}}(h_1, h_2)$, meaning geographic boundary
length governs with the M.4 blend factor applied.
This is the correct behavior: a tract with no commuting data should not be
treated as similar or dissimilar to its neighbors; it should simply not
participate in the commuting shed signal.

\paragraph{Low-count tracts.}
Tracts with very few commuting records ($\sum_w D(h)[w] < 10$) have
statistically unreliable distributions.
The implementation applies a minimum count threshold: tracts with fewer
than 10 total workers in the OD file are treated as empty distributions
(similarity defaults to 0.0).

\subsection{Edge Weight Formula}

The M.4 edge weight applies the M-track blend formula:
\begin{equation}
  w_{\mathrm{cs}}(h_1, h_2)
  = w_{\mathrm{geo}}(h_1, h_2) \times
    \bigl(\alpha + (1 - \alpha)\,\mathrm{sim}(h_1, h_2)\bigr)
  \label{eq:weight}
\end{equation}
with default blend factor $\alpha = 0.5$.

The blend preserves geographic boundary length as a baseline constraint and
adds commuting shed community signal on top.
For tracts with zero commuting similarity (disjoint destination sets),
the edge weight is $\alpha \times w_{\mathrm{geo}}$, reducing the boundary
cost by half.
For tracts with perfect commuting similarity (identical destination
distributions), the edge weight is $w_{\mathrm{geo}}$, preserving the
full geographic cost.

\paragraph{CLI invocation.}
\begin{verbatim}
bisect state --state NC --year 2020 \
    --weights-override commuting-shed \
    --version v_commute_test
\end{verbatim}

\subsection{Test Invariants}

The following L0 unit test invariants must hold for any correct implementation:

\begin{description}
  \item[\texttt{same\_tract\_similarity\_is\_one}.]
    For any home tract $h$ with $\sum_w D(h)[w] > 0$:
    $\mathrm{sim}(h, h) = 1.0$.
    (Minimum and maximum are identical, so the ratio is 1.0.)
  \item[\texttt{disjoint\_destinations\_similarity\_is\_zero}.]
    For two tracts $h_1, h_2$ where $D(h_1)$ and $D(h_2)$ have no
    destination tract in common: $\mathrm{sim}(h_1, h_2) = 0.0$.
  \item[\texttt{jaccard\_symmetric}.]
    For all tract pairs: $\mathrm{sim}(h_1, h_2) = \mathrm{sim}(h_2, h_1)$.
    (Follows from commutativity of $\min$ and $\max$.)
\end{description}
