\documentclass[12pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{amsmath, amssymb, amsthm}
\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}
\newtheorem{proposition}{Proposition}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{remark}{Remark}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage{booktabs}
\usepackage{xcolor}
\usepackage{multirow}
\usepackage{array}
\usepackage{microtype}
\usepackage{setspace}
\usepackage{titlesec}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{lmodern}
\usepackage[T1]{fontenc}

%% SIAM-adjacent style: numbered sections, upright headings
\titleformat{\section}{\normalfont\large\bfseries}{\thesection.}{0.5em}{}
\titleformat{\subsection}{\normalfont\normalsize\bfseries}{\thesubsection.}{0.5em}{}
\titleformat{\subsubsection}{\normalfont\normalsize\itshape}{\thesubsubsection.}{0.5em}{}

%% Spacing
\setstretch{1.4}

%% Shared algorithm macros
\newcommand{\metis}{\textsc{Metis}}
\newcommand{\sa}{\textsc{SimulatedAnnealing}}
\newcommand{\be}{\textsc{BisectionEnsemble}}
\newcommand{\ar}{\textsc{ApportionRegions}}
\newcommand{\recom}{\textsc{ReCom}}
\newcommand{\cs}{\textsc{ConvergenceSweep}}
\newcommand{\smc}{\textsc{SMC}}
\newcommand{\ilp}{\textsc{ILP}}

%% U.7-specific macros
\newcommand{\pareto}{\textsc{Pareto}}
\newcommand{\nsga}{\textsc{NSGA-II}}
\newcommand{\ecm}{\mathrm{EC}}      % math-mode only
\newcommand{\ppm}{\mathrm{PP}}      % math-mode only
\newcommand{\reock}{\mathrm{Reock}} % math-mode only

%% Math/text dual macros (ensuremath for use in both modes)
\newcommand{\EC}{\ensuremath{\mathrm{EC}}}
\newcommand{\PP}{\ensuremath{\mathrm{PP}}}
\newcommand{\Dseats}{\ensuremath{D_{\mathrm{seats}}}}
\newcommand{\VRA}{\ensuremath{\mathrm{VRA}_{\mathrm{deficit}}}}

%% Single-run dagger (avoid redefining existing \dagger)
\newcommand{\singledagger}{\ensuremath{{}^\dagger}}

%% -----------------------------------------------------------------------

\title{%
  \textbf{Pareto-Optimal Redistricting:}\\
  \textbf{Transparent Trade-offs via Multi-Objective Genetic Algorithms}\\[0.8em]
  \normalsize\textit{U.7 --- Algorithm Design / Multi-Objective Optimisation}%
}

\author{%
  Gio Della-Libera\\
  \textit{Apportionment Research Group}\\
  \small\texttt{giodl@microsoft.com}
}

\date{May 2026}

\begin{document}

\maketitle

\begin{abstract}
\noindent
All current redistricting optimisation uses \emph{weighted-sum objectives}: practitioners
minimise $w_1 \cdot \EC + w_2 \cdot \text{partisan\_deviation} + w_3 \cdot \VRA$,
tuning weights until a desired plan emerges.
This is opaque: the weight choice determines the outcome, yet the relationship
between weights and plans is non-linear and non-transparent.

We introduce \pareto{}-optimal redistricting, which instead returns a
\emph{frontier} of plans --- the NSGA-II Pareto front --- where no plan is
strictly better than any other on all three objectives simultaneously.
Practitioners choose from the frontier with full visibility of trade-offs:
``this plan gives more compact districts at the cost of two fewer
proportionally-owed seats.''

The three objectives are (1) \EC: graph edge cuts (compactness),
(2) $\Dseats$: distance from proportional seat allocation (partisan
neutrality), and (3) $\VRA$: minority VAP shortfall below 50\% in
formerly majority-minority districts (VRA compliance).
All three are minimised.

The central legal application is the \emph{enacted-plan dominance test}:
if the NSGA-II frontier contains a plan that strictly beats the enacted plan
on all three objectives simultaneously, the legislature chose a provably worse
option on its own stated criteria.
This is direct, non-statistical evidence of gerrymandering, distinct in
character from ensemble-based arguments.

We present NSGA-II adapted for redistricting plans as chromosomes, with a
ReCom-style crossover operator and a single boundary-tract flip mutation.
Empirical results on NC ($k=14$, 14 congressional districts) and WI ($k=8$)
demonstrate meaningful Pareto frontiers of typically 18--24 plans and characterise
enacted-plan dominance for both states.

\textbf{D\_seats note:} With NC $k=14$, $\Dseats$ takes only $\sim$14 discrete
values.  The Pareto ordering may then collapse to a two-dimensional
$\EC$--$\VRA$ frontier when many plans share the same $\Dseats$ value;
this is acknowledged in output metadata (\texttt{d\_seats\_discrete: true}).
Phase~2 will introduce vote margin as a continuous proxy.
\end{abstract}

\tableofcontents
\newpage

\input{sections/01-introduction}
\input{sections/02-background}
\input{sections/03-algorithm}
\input{sections/04-results}
\input{sections/05-legal}
\input{sections/06-conclusion}

\bibliographystyle{plainnat}
\bibliography{references}

\end{document}
