\label{sec:results}

All results in this section are single-seed runs using
\texttt{--search single}, reported with a dagger~($^\dagger$) per
the \textsc{bisect} empirical reporting convention.

\subsection{Texas: Spanish Section~203 Coverage and County Splits}

Texas has 67 Section~203-covered jurisdictions under the 2023 determination,
primarily for Spanish. Of these, 54 are counties with Spanish coverage,
concentrated in the Rio Grande Valley and South Texas border region
but extending northward into major metropolitan areas including Bexar
County (San Antonio), El Paso County, and Webb County (Laredo).

\textbf{Enacted 2022 map}: The enacted Texas 38-district congressional
plan splits 17 of the 54 Spanish-covered counties between two or more
districts, producing 17 covered county splits.

\textbf{\textsc{bisect} standard geographic mode}: Using
\texttt{--weights-override geographic}, \textsc{bisect} produces a plan
with 15 covered county splits$^\dagger$---2 fewer than the enacted map,
a 12\% reduction.

\textbf{\textsc{bisect} county-weight mode}: Using
\texttt{--weights-override county}, \textsc{bisect} reduces covered county
splits to 11$^\dagger$---a 35\% reduction relative to the enacted map,
and 27\% fewer than the standard geographic mode.

The county-weight mode's improvement is concentrated in the border region,
where contiguous covered counties can often be combined into whole districts
without violating population balance requirements. Webb County (Laredo),
which is covered for Spanish and was split in the enacted plan, is kept
intact under the county-weight mode.

\textbf{Vietnamese community in Harris County}: Texas also has three
Section~203-covered jurisdictions for Vietnamese (Harris County, Fort Bend County,
and Collin County). The Vietnamese community in Harris County is
concentrated in southwest Houston, particularly in the Alief area.
Under the enacted plan, the Alief area is split between two congressional
districts (the 7th and 22nd). Under the \textsc{bisect} county-weight plan,
the Vietnamese community tracts in Alief achieve a community integrity
score of 0.73$^\dagger$ (8 of 11 tracts in the same district), up from
0.45 under the enacted plan.

\subsection{California: Multiple Language Coverage}

California has 29 Section~203-covered jurisdictions under the 2023
determination, with coverage in Spanish, Chinese (including Cantonese
and Mandarin), Vietnamese, Korean, Filipino (Tagalog), and Japanese.
Santa Clara County (San Jose metropolitan area) is covered for Vietnamese,
Chinese, and Korean, making it one of the most linguistically complex
redistricting contexts in the country.

\textbf{Vietnamese community in Santa Clara County}: The Vietnamese community
in San Jose is concentrated in a cluster of census tracts in the
Eastside and Berryessa neighborhoods. The enacted 2022 California
52-district congressional plan assigns these tracts primarily to the
17th and 18th congressional districts, splitting the core Vietnamese
community cluster across district lines.

\textsc{bisect} with geographic weights assigns 3 of the 4 identified
Vietnamese-majority tracts ($\geq 25\%$ Vietnamese-American population)
to the same district$^\dagger$ (community integrity = 0.75), compared
with 1 of 4 under the enacted plan (community integrity = 0.25).
With county-weight mode, the result is the same: 3 of 4 tracts in one
district$^\dagger$, with the fourth in a geographically adjacent district.

\textbf{Spanish coverage in Los Angeles County}: Los Angeles County is
covered for Spanish and contains 18 of California's 38 and later 52
congressional districts. The large number of districts necessarily
produces county splits; the question is whether Section~203-covered
communities within the county are kept intact.

\textsc{bisect} county-weight mode achieves a 9\% reduction in total
county splits in California relative to the enacted plan$^\dagger$,
but because Los Angeles's splits are largely driven by population
arithmetic rather than mapmaker choice, the improvement is more modest
than in Texas.

\subsection{Summary}

\begin{table}[ht]
\centering
\caption{Section~203 County Splits: \textsc{bisect} vs.\ Enacted Maps (2022, single-seed$^\dagger$)}
\label{tab:results-203}
\begin{tabular}{lcccc}
\toprule
State & Enacted & \textsc{bisect} & \textsc{bisect} & Reduction \\
      & Splits  & Geographic & County-Weight & (County-Wt.\ vs.\ Enacted) \\
\midrule
TX (Spanish) & 17 & 15$^\dagger$ & 11$^\dagger$ & 35\% \\
CA (all) & 22 & 21$^\dagger$ & 20$^\dagger$ & 9\% \\
\bottomrule
\end{tabular}
\begin{tablenotes}
\small
\item Section~203 covered county splits (counties split between 2+ districts).
      All single-seed runs ($^\dagger$). TX: 54 Spanish-covered counties;
      CA: 29 jurisdictions across 6 languages, condensed to county level.
\end{tablenotes}
\end{table}
