\section{Conclusion}

Congressional district compactness is negatively associated with member
ideological extremism ($\hat\beta = -0.071$ per 0.10 PP, $p < 0.05$) after
controlling for member fixed effects, district partisan composition, state
fixed effects, and congress fixed effects.
The effect is consistent with the safe-seat polarization mechanism: more
compact districts face more competitive general elections, producing less
extreme winning candidates.

The counterfactual projection suggests that replacing the five study states'
enacted gerrymanders with \textsc{bisect} algorithmic plans would reduce
mean absolute DW-NOMINATE by approximately 0.072 points --- a 13\% reduction.
This represents the O-track's most theoretically significant finding:
redistricting methodology does not merely affect the geometric properties
of districts or the partisan composition of legislatures;
it may affect the quality of legislative representation itself by changing
the ideological character of the members elected.

Two caveats deserve emphasis.
First, the within-member variation in compactness across redistricting cycles
is small (typically 0.05--0.15 PP), limiting statistical precision.
Second, the projection assumes the within-member slope applies to
between-plan comparisons.
\textbf{Formal statement of the identifying assumption}: we assume
$\mathbb{E}[|\text{NOMINATE}|_{\text{bisect plan}}] - \mathbb{E}[|\text{NOMINATE}|_{\text{enacted}}]
= \hat\beta \times \Delta\mathrm{PP}$,
where $\hat\beta$ is the within-member OLS coefficient.
This assumption fails if members' ideological choices are endogenous to district
compactness beyond what the member fixed effects absorb (e.g., if members anticipate
redistricting changes and adjust ideology in advance), or if between-plan PP differences
are confounded by unobserved district-level factors not present in within-member
variation.
The projection is therefore labeled preliminary${}^\dagger$;
a Phase~2 analysis using redistricting cycle boundaries as an instrumental
variable for compactness change would provide the causal identification
this projection cannot.
