Housing stock character---the mix of single-family and multifamily units,
ownership versus rental tenure, and the vintage of the built environment---
encodes information about community structure that is absent from occupational,
economic, or administrative data.
Single-family homeowner neighborhoods share zoning politics, property tax
interests, school district bond capacities, HOA governance structures, and a
common stake in neighborhood stability that is fundamentally different from
the interests of high-rise rental corridors or postwar apartment complexes.
Yet no existing redistricting algorithm operationalizes this distinction.

This paper introduces M.3, the housing character implementation of the
M-track community character framework~\citep{m0framework}.
Using three American Community Survey (ACS) 5-year tables---B25024
(units in structure), B25003 (tenure by occupancy status), and B25035
(median year structure built)---we construct a four-component housing
character vector for each census tract:
percentage single-family units, percentage large-multifamily units,
owner-occupancy rate, and housing vintage score.
The cosine similarity of adjacent tracts' housing character vectors
modifies METIS edge weights via the standard M-track blend formula.

M.3 has a distinctive implementation advantage: the ACS tables it requires are
available through the same Census API infrastructure already used by
\texttt{bisect fetch}. The implemented \texttt{acs-housing} fetch path writes
the four derived columns, and \texttt{--weights-override housing-character}
applies the edge-weight blend.

Implementation mechanics are backed by a hash-bound smoke package; empirical
outcome results remain future work until full plan runs are archived.
The empirical design targets North Carolina (14 congressional districts,
2020 vintage): expected outcomes include within-district homogenization of
housing type and clustering of pre-1960 historic neighborhoods.
The legal foundation draws on neighborhood stability jurisprudence,
historic preservation district doctrine, and long-tenured homeowner
representation interests recognized in state redistricting criteria.
