\label{sec:algorithm}

\subsection{ACS Table Structure}

\paragraph{B25024 --- Units in Structure.}
Table B25024 reports the number of housing units in each census tract broken
down by structural type.
The categories are:
1-unit detached (B25024\_002),
1-unit attached (B25024\_003),
2 units (B25024\_004),
3 or 4 units (B25024\_005),
5 to 9 units (B25024\_006),
10 to 19 units (B25024\_007),
20 or more units (B25024\_008),
mobile home (B25024\_009),
boat, RV, van, etc.\ (B25024\_010).
The universe is all housing units in the tract.
The total is B25024\_001.

\paragraph{B25003 --- Tenure.}
Table B25003 reports the number of occupied housing units by tenure:
owner-occupied (B25003\_002) and renter-occupied (B25003\_003).
The universe is occupied housing units.
The total is B25003\_001.

\paragraph{B25035 --- Median Year Structure Built.}
Table B25035 reports the median calendar year in which the housing structures
in the tract were built, as a single integer.
This is a tract-level median, not a microdata file.
The universe is all housing units.

\subsection{Four-Component Housing Character Vector}

For each census tract $t$, we define the housing character vector
\begin{equation}
  \mathbf{h}(t) = \bigl(\mathrm{SF}(t),\ \mathrm{MF}(t),\ \mathrm{OO}(t),\ \mathrm{VT}(t)\bigr)
  \in [0,1]^4,
  \label{eq:housingvec}
\end{equation}
where the four components are defined as follows.

\paragraph{Single-family fraction (SF).}
\begin{equation}
  \mathrm{SF}(t) = \frac{\text{B25024\_002}(t) + \text{B25024\_003}(t)}{\text{B25024\_001}(t)},
  \label{eq:sf}
\end{equation}
the fraction of units in 1-unit detached or attached structures.
$\mathrm{SF}(t) \in [0,1]$.
High values ($> 0.80$) characterize suburban single-family neighborhoods and
urban rowhouse districts.
Low values ($< 0.20$) characterize high-rise apartment corridors and
dense urban cores.

\paragraph{Large-multifamily fraction (MF).}
\begin{equation}
  \mathrm{MF}(t) = \frac{\text{B25024\_007}(t) + \text{B25024\_008}(t)}{\text{B25024\_001}(t)},
  \label{eq:mf}
\end{equation}
the fraction of units in structures with 10 or more units.
Note: the M.4 spec uses ``5+ units'' informally; the precise B25024 category boundary for large multifamily is 10 units (B25024\_007: 10--19 units; B25024\_008: 20 or more units). Structures with 5--9 units (B25024\_006) are excluded from the large-multifamily fraction because they are more similar in community character to small apartment buildings than to large elevator buildings.
$\mathrm{MF}(t) \in [0,1]$.
High values ($> 0.50$) characterize apartment corridors, urban high-density
residential zones, and transit-oriented development nodes.
The MF component distinguishes apartment-corridor communities from
small-multifamily neighborhoods (duplexes, triplexes) that may share
more community characteristics with single-family areas.

\paragraph{Owner-occupancy rate (OO).}
\begin{equation}
  \mathrm{OO}(t) = \frac{\text{B25003\_002}(t)}{\text{B25003\_001}(t)},
  \label{eq:oo}
\end{equation}
the fraction of occupied units that are owner-occupied.
$\mathrm{OO}(t) \in [0,1]$.
High values ($> 0.70$) characterize established homeowner communities.
Low values ($< 0.30$) characterize predominantly rental communities such as
university neighborhoods, downtown cores, and transit-oriented apartment
districts.

\paragraph{Housing vintage score (VT).}
\begin{equation}
  \mathrm{VT}(t) = \max\!\left(0,\; \min\!\left(1,\; 1.0 - \frac{\text{B25035}(t) - 1940}{2020 - 1940}\right)\right),
  \label{eq:vt}
\end{equation}
a score that maps the median year built to the interval $[0,1]$, where higher
values indicate older housing stock.
\emph{Implementation note:} The M.3 spec approximates this as $(2024 - \text{B25035}) / 100$;
the present formula supersedes that approximation.
The spec formula is unbounded (it exceeds 1.0 for tracts with median year before 1924)
and uses an arbitrary divisor.
Equation~\eqref{eq:vt} is the implementation target: it is clamped to $[0,1]$,
calibrated on the 1940--2020 housing stock range, and consistent with the
$^\dagger$Phase~2 test invariants.
The formula is calibrated so that:
\begin{itemize}
  \item Median year 1940 or earlier: $\mathrm{VT} = 1.0$ (fully historic).
  \item Median year 1980: $\mathrm{VT} = 0.5$ (mid-century suburban).
  \item Median year 2010: $\mathrm{VT} = 0.125$ (new suburban).
  \item Median year 2020 or later: $\mathrm{VT} = 0.0$ (new construction, clamped).
\end{itemize}
The $\max(0, \min(1, \cdot))$ clamp handles the edge cases of tracts with
B25035 values outside the 1940--2020 calibration range.

\subsection{Cosine Similarity and Edge Weight}

The housing character similarity between adjacent tracts $u$ and $v$ is the
standard cosine similarity of their housing character vectors:
\begin{equation}
  \mathrm{sim}_h(u,v) = \frac{\mathbf{h}(u) \cdot \mathbf{h}(v)}{\|\mathbf{h}(u)\| \, \|\mathbf{h}(v)\|}
  \in [0,1],
  \label{eq:cosine}
\end{equation}
which is well-defined when neither vector is the zero vector.
Because all four components are non-negative by construction, the cosine
similarity is always in $[0,1]$.

\paragraph{Zero-vector handling.}
A census tract can produce a zero housing character vector only if all four
of B25024\_001, B25003\_001, and B25035 are zero or missing---in practice, a
tract with no housing units (e.g., an industrial or parkland tract with
residential census population assigned to it).
Such tracts receive the conservative substitution $\mathrm{sim}_h = 1.0$
(treat as perfectly similar to any neighbor), leaving the edge weight
governed entirely by the geographic base weight.

\paragraph{Edge weight blend formula.}
Following the M-track blend convention established in M.1 and B.27, the
housing-character edge weight is:
\begin{equation}
  w_h(u,v) = w_{\mathrm{geo}}(u,v) \times \bigl(\alpha + (1 - \alpha)\,\mathrm{sim}_h(u,v)\bigr),
  \label{eq:blend}
\end{equation}
with default $\alpha = 0.5$.
At $\alpha = 0.5$ and $\mathrm{sim}_h = 1.0$ (identical housing characters),
the weight equals $w_{\mathrm{geo}}$: maximum community bond.
At $\alpha = 0.5$ and $\mathrm{sim}_h = 0.0$ (orthogonal housing characters),
the weight equals $0.5 \times w_{\mathrm{geo}}$: the geographic bond is
halved rather than zeroed, preventing free-cut instability in METIS.

\subsection{Prototypical Housing Character Vectors}

Table~\ref{tab:prototypes} gives three prototypical tract types and their
housing character vectors to build intuition for the cosine similarity geometry.

\begin{table}[ht]
\centering
\caption{Prototypical housing character vectors for three tract types.
  SF = single-family fraction, MF = large-multifamily fraction,
  OO = owner-occupancy rate, VT = vintage score.}
\label{tab:prototypes}
\begin{tabular}{llrrrr}
\toprule
Type & Example & SF & MF & OO & VT \\
\midrule
Single-family suburb & Cary, NC residential & 0.88 & 0.02 & 0.75 & 0.30 \\
Historic rowhouse district & Raleigh Cameron Park & 0.72 & 0.05 & 0.65 & 0.85 \\
Apartment corridor & Charlotte South End & 0.08 & 0.72 & 0.18 & 0.20 \\
\bottomrule
\end{tabular}
\end{table}

The cosine similarity matrix reveals the community geometry.

\begin{table}[ht]
\centering
\caption{Cosine similarity between prototypical housing character tract types.}
\label{tab:similarity}
\begin{tabular}{lrrr}
\toprule
 & Single-family suburb & Historic rowhouse & Apartment corridor \\
\midrule
Single-family suburb & 1.00 & 0.94 & 0.11 \\
Historic rowhouse    & 0.94 & 1.00 & 0.20 \\
Apartment corridor   & 0.11 & 0.20 & 1.00 \\
\bottomrule
\end{tabular}
\end{table}

The similarity matrix encodes a central finding of the housing character framework:
\begin{itemize}
  \item \textbf{Single-family suburb to historic rowhouse (0.94).}
    Both are predominantly single-family, owner-occupied communities.
    Despite the vintage difference, they share housing character.
    METIS should strongly avoid cutting the edge between them.
  \item \textbf{Suburb or rowhouse to apartment corridor ($<$0.20).}
    The housing character community boundary is the sharpest possible
    cut seam: single-family owner versus large-multifamily renter.
    METIS is invited to place the district boundary here.
  \item \textbf{Apartment corridor to apartment corridor (1.00 for identical vectors).}
    Apartment corridor communities belong together: they share rental market
    dynamics, transit-oriented mobility patterns, and land-use policy interests.
\end{itemize}

\subsection{No New Download Infrastructure Required}

A key practical advantage of M.3 is that the three ACS tables (B25024, B25003,
B25035) are available from the same Census API endpoint already used by
\texttt{bisect fetch} to populate the demographics CSV.
The current ACS fetch retrieves tables covering race, age, income, and
educational attainment at the census tract level.
Adding B25024, B25003, and B25035 requires:
\begin{enumerate}
  \item Extending the \texttt{acs\_fetch} module's table list with the new
    table identifiers.
  \item Adding four output columns to the demographics CSV:
    \texttt{pct\_single\_family}, \texttt{pct\_multifamily},
    \texttt{pct\_owner\_occupied}, \texttt{housing\_vintage}.
  \item Implementing the four formulas (Equations~\ref{eq:sf}--\ref{eq:vt})
    in the demographics preprocessing step.
  \item Implementing the cosine similarity and blend formula (Equations~\ref{eq:cosine}
    and~\ref{eq:blend}) in the \texttt{housing-character} weight mode.
\end{enumerate}
No new API calls to new endpoints, no new file formats, and no changes to
the adjacency graph pipeline are required.
The CLI flag is \texttt{--weights-override housing-character}.

\subsection{Test Invariants (L0)}

The following invariants must hold on every build:

\begin{description}
  \item[\texttt{single\_family\_tracts\_high\_similarity}]
    For two tracts with $\mathrm{SF} > 0.85$, $\mathrm{OO} > 0.70$,
    and $|\mathrm{VT}_u - \mathrm{VT}_v| < 0.20$:
    $\mathrm{sim}_h(u,v) > 0.85$.
  \item[\texttt{mixed\_tenure\_vs\_homogeneous\_low\_similarity}]
    For $u$ with $\mathrm{OO} = 0.9, \mathrm{SF} = 0.9$ and
    $v$ with $\mathrm{OO} = 0.1, \mathrm{MF} = 0.8$:
    $\mathrm{sim}_h(u,v) < 0.30$.
  \item[\texttt{vintage\_far\_apart\_low\_similarity}]
    For $u$ with $\mathrm{VT} = 0.10$ (built $\approx$2012) and $v$ with
    $\mathrm{VT} = 0.80$ (built $\approx$1956), holding SF and OO equal:
    $\mathrm{sim}_h(u,v) < 0.50$.
  \item[\texttt{similarity\_symmetric}]
    $\mathrm{sim}_h(u,v) = \mathrm{sim}_h(v,u)$ for all tract pairs.
  \item[\texttt{weight\_stays\_nonneg}]
    $w_h(u,v) \geq 0$ for all edges after blending.
\end{description}
