\label{sec:conclusion}

This paper has introduced M.3, the housing character implementation of the
M-track community character framework.
Three ACS tables---B25024 (units in structure), B25003 (tenure), and B25035
(median year structure built)---produce a four-component vector that captures
the single-family/multifamily dimension, the owner/renter dimension, and
the historic/new-construction dimension of residential built environment
character.
The cosine similarity of housing character vectors modifies METIS edge weights
via the M-track blend formula, causing district boundaries to fall preferentially
at the transitions between single-family homeowner communities and multifamily
rental corridors, and between historic pre-war neighborhoods and postwar
suburban tracts.

Several features make M.3 distinctive within the M-track family.

\paragraph{Zero implementation cost.}
M.3 requires no new data download infrastructure.
The three ACS tables are available from the same Census API endpoint already
used by \texttt{bisect fetch}.
Four new columns extend the existing demographics CSV without restructuring
the pipeline.
This makes M.3 the fastest M-track signal to move from specification to
implementation.

\paragraph{Legal recognition depth.}
The housing character signal has unusually deep legal recognition as a
community-of-interest indicator.
Homeownership creates a legally enforceable fiscal bond through property
taxation.
Historic district designation creates an administrative recognition of
community cohesion.
The homevoter hypothesis~\citep{fischel2001} provides theoretical grounding
for why housing character predicts shared political interests.
Expert testimony from redistricting proceedings in Oregon and Colorado has
accepted historic neighborhood cohesion as a community of interest.

\paragraph{Orthogonality to existing signals.}
M.3 is orthogonal to M.1 (economic character): M.1 describes the workplace
side of a census tract, while M.3 describes the residential built environment.
A residential tract can have low economic character (few jobs) but high housing
character homogeneity.
M.3 is also largely orthogonal to M.6 (administrative zone co-membership),
though the two signals correlate in the school district dimension.
Composing M.3 and M.6 weights would capture both the physical built environment
and the administrative fiscal bond dimensions of community cohesion simultaneously.

\paragraph{Future work.}
Three natural extensions follow from this paper.
First, empirical validation using the implemented ACS housing pipeline will test
the within-district homogeneity hypotheses (H1--H3) on North Carolina.
Second, a multi-state sweep comparing housing-character weights to geographic
baseline across all 50 states will identify which states have sufficient
housing character variation to benefit from M.3.
Third, the composition of M.3 with M.1 and M.6---a three-signal composite
community character weight---is the natural next experiment for states with
strong communities-of-interest criteria and diverse built environments.
