Every ten years, following the decennial census, the United States Congress
apportions 435 House seats among the 50 states in proportion to population.
This apportionment step --- not redistricting, which comes after --- determines
\emph{how many} seats each state receives before a single district boundary
is drawn.
Getting the apportionment step right is a precondition for any legitimate
redistricting: a pipeline that draws 15 districts for North Carolina
when that state is entitled to 14 has introduced an error at the foundation
that no amount of subsequent algorithmic sophistication can correct.

The constitutional mandate is clear in form but silent on method.
Article~I, \S2 reads: ``Representatives\ldots shall be apportioned among
the several States according to their respective Numbers,'' and
the Fourteenth Amendment, \S2, refined this to ``the whole number of persons
in each State.''
The Supreme Court confirmed in \textit{Wesberry v.\ Sanders}, 376 U.S.\ 1
(1964), that congressional districts must be as equal in population as
practicable \citep{wesberry1964}, establishing that mathematical precision
in apportionment and districting is a constitutional value.
Yet Congress has changed the apportionment method eleven times since 1789
\citep{balinski1982fair}, and each change has reflected different
mathematical and political judgements about what ``proportional'' means
when the exact quotas are not integers.

\subsection{The Rounding Problem}

The core mathematical difficulty is that exact proportional shares are
generally not integers.
If the U.S.\ apportionment population is $N = 331{,}108{,}434$ and the
House size is $H = 435$, the national divisor is $d = N/H \approx 761{,}169$
persons per seat.
California's apportionment population of approximately $39{,}538{,}223$
yields an exact quota of $q_{\text{CA}} = 39{,}538{,}223 / 761{,}169
\approx 51.94$ seats --- which must be rounded to an integer.
Wyoming's population of approximately $576{,}851$ yields a quota of
$q_{\text{WY}} \approx 0.758$, which must be rounded to at least 1
(the constitutional minimum).
Different rounding rules for these fractional parts define different
apportionment methods, and they disagree in close cases.

\subsection{Five Methods}

The five methods that have received serious legislative and academic
attention are:
\begin{enumerate}
  \item \textbf{Huntington-Hill (equal proportions)}: round at the geometric
    mean of consecutive integers; used since 1941 \citep{huntington1928method}.
  \item \textbf{Webster (major fractions)}: round at the arithmetic mean
    (i.e., at 0.5); used 1842 and 1911--1940 \citep{balinski1982fair}.
  \item \textbf{Adams (smallest divisors)}: always round up (ceiling);
    never used federally but studied for its small-state properties
    \citep{balinski1982fair}.
  \item \textbf{Jefferson (greatest divisors)}: always round down (floor);
    used 1790--1840 under Jefferson's recommendation \citep{schmeckebier1941congressional}.
  \item \textbf{Hamilton (largest remainders)}: allocate integer parts first,
    then distribute remaining seats to states with largest fractional remainders;
    used 1850--1900 \citep{balinski1982fair}.
\end{enumerate}
Methods 1--4 belong to the \emph{divisor family}; method 5 is a \emph{quota method}.
This distinction is the most important structural division in apportionment theory.

\subsection{Scope and Contribution}

This paper makes three contributions to the J-track.
First, we develop the divisor vs.\ quota taxonomy formally, showing that
the two families differ not just in computation but in their fundamental
mathematical properties (Section~\ref{sec:taxonomy}).
Second, we characterise all five methods along three fairness dimensions ---
the population pair test (which divisor method treats each pair of states
most equitably), the quota rule (which methods guarantee that each state
receives a seat count within 1 of its exact quota), and bias direction
(which states gain or lose seats relative to strict proportionality)
--- and provide the 2020 Census seat allocation table for all five methods
and all 50 states (Section~\ref{sec:fairness}).
Third, we document the \textsc{bisect-apportion} crate's implementation
of all five methods and its SHA-256-verified reproduction of the official
2020 apportionment (Section~\ref{sec:implementation}).

\subsection{Data}

All computations use 2020 Census Bureau apportionment population figures
as published in the April~26, 2021 release \citep{census2020apportionment}.
The apportionment population for each state counts all residents including
overseas federal employees allocated to a home state; it excludes the
District of Columbia (which has no House representation) and U.S.\ territories.
Total apportionment population: 331{,}108{,}434.
House size: 435.
