\label{sec:implementation}

The \textsc{bisect-apportion} crate implements all five apportionment
methods in Rust with a unified interface.
It is invoked before each redistricting run to determine the seat
count for each state, and its Huntington-Hill output is SHA-256-verified
against the official 2020 Census Bureau apportionment.

\subsection{Architecture}

The crate exposes a single primary function:

\begin{verbatim}
pub fn apportion(
    populations: &[(String, u64)],
    house_size: usize,
    method: ApportionMethod,
) -> Vec<(String, usize)>
\end{verbatim}

\noindent
where \texttt{ApportionMethod} is an enum with variants
\texttt{HuntingtonHill}, \texttt{Webster}, \texttt{Adams},
\texttt{Jefferson}, and \texttt{Hamilton}.
The function returns a sorted vector of (state name, seat count) pairs.

All divisor methods share a priority-queue implementation:
\begin{enumerate}
  \item Assign each state 1 seat (constitutional floor).
  \item Build a max-heap of (priority, state\_index) pairs.
  \item For each of the remaining $H - n$ seats, pop the state with
    highest priority, increment its seat count, and push its new
    priority back onto the heap.
\end{enumerate}
The priority formula differs by method:
for Huntington-Hill, $P_i(s) = p_i^2 / (s \cdot (s+1))$
(squared to avoid floating-point square roots --- equivalent to
$p_i / \sqrt{s(s+1)}$ but computed with integer arithmetic).

\subsection{Integer Arithmetic for Exact Comparison}

A critical implementation detail is tie-breaking.
When two states have exactly equal priority values, the method of
comparison determines which state receives the seat.
The crate represents all priority values as rational numbers
$P_i(s) = p_i^2 / (s(s+1))$ stored as pairs of \texttt{u128}
integers, enabling exact comparison via cross-multiplication
without floating-point rounding.
This is essential for the SHA verification: floating-point priority
comparison can produce wrong results in near-tie cases,
whereas integer comparison is exact.

\subsection{SHA-256 Verification}

The 2020 Census Bureau published its official apportionment on
April~26, 2021 \citep{census2020apportionment}.
The \textsc{bisect-apportion} verification protocol computes
the SHA-256 hash of the string formed by concatenating
``\texttt{StateA:seats\textbackslash n}'' lines sorted alphabetically,
and compares this to the hash of the same string constructed
from the Census Bureau table.
The two hashes match exactly, confirming zero discrepancies
across all 50 states.

\subsection{CLI Usage}

\begin{verbatim}
bisect apportion --year 2020 --method huntington-hill --verify
bisect apportion --year 2020 --method webster
bisect apportion --year 2020 --method adams
bisect apportion --year 2020 --method jefferson
bisect apportion --year 2020 --method hamilton
\end{verbatim}

The \texttt{--verify} flag triggers the SHA comparison and exits
non-zero if the hash does not match the committed reference.
This makes apportionment correctness a CI-enforced invariant.
