\label{sec:architecture}

\subsection{Public API}

The crate exposes two primary apportionment functions:

\begin{verbatim}
/// Huntington-Hill apportionment (2 U.S.C. sec 2a).
/// `populations`: map from state code to total population.
pub fn huntington_hill(
    populations: &HashMap<String, u64>,
    total_seats: u32,
) -> HashMap<String, u32>

/// General divisor-method apportionment.
/// `rule` selects HuntingtonHill, Webster, Adams, or Jefferson.
pub fn apportionment_divisor(
    populations: &[(String, u64)],
    house_size: u32,
    rule: RoundingRule,
) -> Vec<(String, u32)>

pub enum RoundingRule {
    HuntingtonHill,
    Webster,
    Adams,
    Jefferson,
}
\end{verbatim}

Both functions guarantee at least 1 seat per state (constitutional floor)
and a total seat count equal to \texttt{total\_seats} / \texttt{house\_size}.

Secondary utilities:

\begin{verbatim}
/// Check if Hamilton produces the Alabama paradox at this H.
pub fn check_alabama_paradox(
    populations: &[(String, u64)],
    house_size: usize,
) -> Option<(String, usize, usize)>
\end{verbatim}

\subsection{Priority Queue Implementation}

The divisor methods (HH, Webster, Adams, Jefferson) share a
priority-queue core in \texttt{apportionment\_divisor}:

\begin{enumerate}
  \item Assign each state 1 seat (constitutional floor).
  \item Build a max-heap keyed by each state's \texttt{f64} priority
    for the next seat.
  \item For each of the remaining $H - n$ seats, pop the state with
    highest priority, increment its seat count, and push its updated
    priority back onto the heap.
\end{enumerate}

The key loop (simplified):
\begin{verbatim}
// Seed heap: all states start at 1 seat
for (idx, (name, pop)) in populations.iter().enumerate() {
    heap.push(Entry {
        priority: priority_for(*pop, 1, rule),
        idx,
        name: name.clone(),
    });
}
// Award remaining seats
let remaining = house_size - n_states;
for _ in 0..remaining {
    let Entry { idx, .. } = heap.pop().unwrap();
    seats[idx] += 1;
    heap.push(Entry {
        priority: priority_for(populations[idx].1, seats[idx], rule),
        idx,
        name: populations[idx].0.clone(),
    });
}
\end{verbatim}

\subsection{Priority Representation}

Priority values are stored as \texttt{f64} (double-precision floating-point):

\begin{verbatim}
let priority_for = |pop: u64, n: u32, rule: RoundingRule| -> f64 {
    let n = n as f64;
    let p = pop as f64;
    match rule {
        RoundingRule::HuntingtonHill => p / (n * (n + 1.0)).sqrt(),
        RoundingRule::Webster        => p / (n + 0.5),
        RoundingRule::Adams          => p / n,
        RoundingRule::Jefferson      => p / (n + 1.0),
    }
};
\end{verbatim}

Double-precision floating-point arithmetic (\texttt{f64}) is used for
priority comparisons.
Huntington-Hill priorities involve square roots ($P / \sqrt{n(n+1)}$) and
are therefore irrational; \texttt{f64} introduces rounding error.
The question is whether this error can flip a priority ordering.
The relative error of \texttt{f64} is $\varepsilon_{\rm rel} \approx 2^{-52}
\approx 2.2 \times 10^{-16}$.
For a typical priority value of $10^6$, the absolute rounding error is
$\approx 2.2 \times 10^{-10}$.
For the 2020 Census apportionment, the minimum priority gap (closest two
states ever were to exchanging a seat) was approximately $8.3$ priority
units (Montana's 2nd seat vs.\ North Carolina's 14th seat, a gap of
$\approx 8.3$ computed from the published Census Bureau priority table).
This gap is $\sim 10^{10}\times$ larger than the \texttt{f64} rounding error,
confirming that \texttt{f64} is sufficient for the 2020 data.
\textbf{Caveat}: for hypothetical populations engineered to produce
priority gaps smaller than $\sim 10^{-9}$, \texttt{f64} could give wrong
results. For robustness, a production implementation could fall back to
128-bit rational arithmetic when the minimum gap falls below a configurable
threshold (not implemented in the current version).
Ties are broken deterministically by lexicographic state name order.

\subsection{Hamilton Implementation}

Hamilton does not use the priority queue.
It is implemented separately:

\begin{verbatim}
fn apportion_hamilton(
    populations: &[(String, u64)],
    house_size: usize,
) -> Vec<(String, usize)> {
    let total: u64 = populations.iter().map(|(_, p)| p).sum();
    // Compute exact quotas as rationals
    let quotas: Vec<(u128, u128)> = populations.iter()
        .map(|(_, &p)| (p as u128 * house_size as u128, total as u128))
        .collect();
    // Assign lower quotas
    let mut seats: Vec<usize> = quotas.iter()
        .map(|&(n, d)| (n / d) as usize)
        .collect();
    // Apply constitutional floor
    for s in seats.iter_mut() { *s = (*s).max(1); }
    // Distribute remainder seats by fractional part
    let remainder = house_size.saturating_sub(seats.iter().sum());
    let mut fracs: Vec<(u128, usize)> = quotas.iter()
        .enumerate()
        .map(|(i, &(n, d))| (n % d, i))
        .collect();
    fracs.sort_by(|a, b| b.0.cmp(&a.0));
    for k in 0..remainder {
        seats[fracs[k].1] += 1;
    }
    populations.iter().zip(seats.iter())
        .map(|((name, _), &s)| (name.clone(), s))
        .collect()
}
\end{verbatim}
