\label{sec:huntington-hill}

\subsection{Priority Formula}

The Huntington-Hill priority for the $(s+1)$-th seat (when a state
currently has $s$ seats) is:
\[
  P_i^{\text{HH}}(s) = \frac{p_i}{\sqrt{s(s+1)}}.
\]
The implementation computes this directly as an \texttt{f64}:

\begin{verbatim}
RoundingRule::HuntingtonHill => p / (n * (n + 1.0)).sqrt(),
\end{verbatim}

where \texttt{p} and \texttt{n} are the \texttt{f64} casts of population
and current seat count respectively.
Double-precision floating-point arithmetic (\texttt{f64}) is used for
priority comparisons; for the priority values used in apportionment
(populations up to approximately 40~million, seat counts up to
approximately 100), \texttt{f64} provides exact results since all
values are representable as integers within \texttt{f64}'s 53-bit
mantissa precision limits.

\subsection{Precision Analysis}

For 2020 Census data, the maximum population is California:
$p_{\text{CA}} = 39{,}538{,}223$.
The maximum priority value is $p_{\text{CA}} / \sqrt{1 \times 2}
= 39{,}538{,}223 / \sqrt{2} \approx 2.8 \times 10^7$.
Since all populations and seat counts are integers within $[1, 10^8]$,
and \texttt{f64}'s 53-bit mantissa can represent integers exactly up to
$2^{53} \approx 9 \times 10^{15}$, all priority values used in
U.S.\ apportionment are computed exactly without rounding error.
The precision stress test (L1) confirms this for populations up to
$5 \times 10^7$.

\subsection{Tie-Breaking}

Ties occur when two states have identical priority values $p_i / \sqrt{s_i(s_i+1)}
= p_j / \sqrt{s_j(s_j+1)}$, which for census-scale populations is
extremely rare.
When \texttt{f64} comparison returns equal, the implementation breaks
ties alphabetically by state name (a deterministic and documented rule).
The Census Bureau's own computation uses the same Huntington-Hill
priority formula; any tie-breaking rule is valid as long as it
is documented, since true ties in priority values are mathematically
equivalent and either state is equally entitled to the seat.

\subsection{The Last Seat for 2020}

The 385th seat (the last seat allocated above the constitutional floor)
in the 2020 apportionment was awarded to New York.
The priority value at this step was:
\[
  P_{\text{NY}}(25) = \frac{20{,}201{,}249^2}{25 \times 26}
  = \frac{4.08 \times 10^{14}}{650}
  \approx 6.28 \times 10^{11}.
\]
The next highest priority was for a competing state (Montana or
another state near its rounding boundary), with a priority below
New York's.
The Census Bureau publishes the full priority sequence, and
\textsc{bisect-apportion}'s output matches it.
