\label{sec:other-methods}

\subsection{Webster}

Webster's priority for the $(s+1)$-th seat is $p_i/(s+0.5)$.
The implementation computes this as:

\begin{verbatim}
RoundingRule::Webster => p / (n + 0.5),
\end{verbatim}

where \texttt{p} and \texttt{n} are the \texttt{f64} casts of population
and current seat count.

\subsection{Adams}

Adams uses $P_i^{\text{Ad}}(s) = p_i/s$:

\begin{verbatim}
RoundingRule::Adams => p / n,
\end{verbatim}

Each state starts at 1 seat (constitutional floor), so the first
Adams priority is $p_i/1 = p_i$.

\subsection{Jefferson}

Jefferson uses $P_i^{\text{Jeff}}(s) = p_i/(s+1)$:

\begin{verbatim}
RoundingRule::Jefferson => p / (n + 1.0),
\end{verbatim}

Jefferson initialises with 1 seat per state (constitutional floor),
and the priority formula $p/(n+1.0)$ means the next seat priority
for a state at 1 seat is $p/2$, equivalent to the d'Hondt divisor sequence
under a U.S.-style one-seat floor.

\subsection{Hamilton Boundary}

Hamilton/largest remainder is not currently implemented as a public
\textsc{bisect-apportion} apportionment function.  It remains in the J-track as
the canonical source of the Alabama paradox and as a future comparison target.
Any paper table that mentions Hamilton should therefore be read as mathematical
context or a target fixture, not as current crate output.

\subsection{Edge Cases}

All four implemented divisor methods handle:
\begin{itemize}
  \item \textbf{Single state}: Returns $H$ seats (trivially).
  \item \textbf{Exact quotas}: If $q_i$ is an integer for all $i$
    (and $\sum q_i = H$), the divisor methods agree on the tested fixtures.
  \item \textbf{Constitutional floor}: States with exact quota $< 1$
    receive 1 seat; the priority sequence allocates the remaining seats.
  \item \textbf{Seat totals}: Unit tests assert that all four divisor methods
    return the requested house size on representative synthetic inputs.
\end{itemize}
