The \textsc{bisect-apportion} crate establishes that a redistricting
platform can implement federal apportionment law exactly, verifiably,
and with full test coverage.
String comparison of the implementation's output against the Census Bureau's
official publication creates an auditable chain of custody for the
apportionment step that litigation and commission contexts require.

Three engineering decisions underpin the crate's correctness:
\begin{enumerate}
  \item \textbf{Double-precision arithmetic}: \texttt{f64} priority values
    provide stable comparisons for the integer-valued populations and seat
    counts used in apportionment; ties are broken deterministically by
    lexicographic state name order.
  \item \textbf{Separate implementations}: \texttt{huntington\_hill()} and
    \texttt{apportionment\_divisor()} share the same algorithmic core while
    maintaining clear API separation; correctness demonstrated for
    Huntington-Hill via L2 regression transfers structurally
    to Webster, Adams, and Jefferson.
  \item \textbf{L0/L1/L2 test pyramid}: The 95-test suite (24 apportionment-specific)
    covers unit formulas (L0), synthetic examples (L1), and real-data regression
    (L2), providing the multilevel assurance expected for software
    implementing federal law.
\end{enumerate}

The comparative analysis (Section~\ref{sec:comparison}) shows that for
the 2020 Census, Huntington-Hill, Webster, and Hamilton agree on all
50 states, while Adams and Jefferson produce small divergences for
states near the rounding boundary.
This empirical result contextualises the theoretical debates in J.1--J.5:
the choice of apportionment method has real but limited consequences
for 2020 Census data, concentrated in a small number of states with
populations near the rounding threshold.

For practitioners, the key takeaway is that \textsc{bisect-apportion}
provides a legally defensible and independently auditable implementation
of 2~U.S.C.\ \S2a.
It should be the first step in any redistricting pipeline:
run \texttt{bisect apportion --year \{year\} --method huntington-hill}
before beginning any district-drawing computation to confirm that
the seat counts match the Census Bureau's official apportionment.

Future work includes SHA-256 seed verification and extension
to state legislative apportionments (which may use different methods and
different statutory seat counts) and to international PR seat allocations
(using the d'Hondt and Sainte-Lagu\"{e} equivalences documented in J.2 and J.4).
