\label{sec:introduction}

When courts and commissions turn to algorithmic redistricting as an alternative to
partisan line-drawing, one of the first objections raised by incumbent legislators is
straightforward: the algorithm does not know where I live. This objection is technically
correct. The \textsc{bisect} redistricting system~\citep{bisect2024} receives as inputs
census tract geometries, population counts, and optionally precinct-level election data
for edge weighting---but never incumbent addresses, party registration at the home
address, or any other signal tied to the specific location of sitting legislators.
The algorithm is, in the terminology of this paper, incumbency-blind.

The paradox is that incumbency-blind redistricting is not incumbency-neutral in its
consequences. When a recursive bisection algorithm draws a new district boundary, it
creates a district that may or may not contain an incumbent's home census tract.
Whether it does depends entirely on the geographic coincidence between district boundaries
optimised for population balance and compactness, and the residential addresses of
legislators whose homes were not an input to the optimisation. These coincidences are not
random in the sense of being uniformly distributed across all possible maps; they are random
in the sense of being drawn from the distribution implied by the algorithm's geographic
constraints, which we call the incumbency-neutral baseline.

This paper synthesises four companion papers---I.1 (incumbent pairing), I.2 (safe-seat
creation), I.3 (open-seat effects), and I.4 (legal analysis)---into a unified framework
for understanding how algorithmic redistricting affects incumbent legislators. Our goals
are three:

\begin{enumerate}
  \item \textbf{Descriptive}: Document the incumbency consequences of \textsc{bisect}
    algorithmic maps in NC, WI, and TX on three measurable dimensions (pairing rate,
    safe-seat fraction, open-seat count), and compare these consequences to enacted
    2022 congressional maps and to the incumbency-neutral baseline.
  \item \textbf{Explanatory}: Identify the three mechanisms through which an
    incumbency-blind algorithm produces incumbency outcomes: geographic restructuring,
    partisan composition change, and constituent-base reassignment.
  \item \textbf{Legal}: Establish that \textsc{bisect}'s constitutional silence on
    incumbency is legally sound under existing federal doctrine, and that the
    algorithm's outcomes are indistinguishable from what a neutral random redistricting
    process would produce given the same geographic constraints.
\end{enumerate}

\subsection{Three Mechanisms}

The paradox of incumbency-blind redistricting producing incumbency consequences is
resolved by recognising that district boundaries affect incumbents through three
geometric mechanisms that operate independently of any incumbent-specific intent:

\paragraph{Geographic restructuring.}
When district boundaries shift, two incumbents who previously lived in separate districts
may find themselves in the same new district. This \textit{incumbent pairing} forces at
least one incumbent to retire, seek an open seat in a different district, or face a
primary against a fellow incumbent. The algorithm does not intend to pair them; the
pairing is a consequence of population-balance optimisation drawing a boundary through
a region where two incumbents happened to live close enough together to fall on the
same side of the cut.

\paragraph{Partisan composition change.}
Even when an incumbent remains the sole member of Congress residing in a new district,
the district's partisan lean may change dramatically if the algorithm's geographic
boundary reassigns neighbourhoods that were heavily partisan in one direction. An
incumbent who held a safe seat under the enacted plan may find herself in a competitive
seat under the algorithmic plan---not because the algorithm targeted her, but because
the compactness-optimal boundary happened to include or exclude a suburban ring that
shifted the district's presidential vote share by 10 percentage points.

\paragraph{Constituent-base reassignment.}
Even with stable partisan lean, boundary shifts reassign specific communities of
constituents from one incumbent to another. Incumbents invest heavily in constituent
service relationships within their current districts; losing a county or a congressional
district's share of a city disrupts those relationships and may create a different
electoral environment regardless of aggregate partisan numbers.

\subsection{Paper Organisation}

Section~\ref{sec:background} reviews the legal background on incumbency as a
redistricting criterion, covering the key federal cases and the VRA tension.
Section~\ref{sec:methodology} introduces the three metrics (pairing rate, safe-seat
fraction, open-seat count) and the incumbency-neutral baseline.
Section~\ref{sec:results} presents the empirical overview across NC, WI, and TX.
Section~\ref{sec:discussion} develops the neutrality argument and its legal implications.
Section~\ref{sec:conclusion} concludes.
