\label{sec:conclusion}

This paper has presented a unified framework for understanding how \textsc{bisect}
algorithmic redistricting interacts with congressional incumbency. Three mechanisms---
geographic restructuring, partisan composition change, and constituent-base
reassignment---explain how an algorithm with no incumbency inputs nonetheless produces
measurable incumbency consequences. Three metrics---incumbent pairing rate, safe-seat
fraction, and open-seat count---provide the empirical vocabulary for comparing algorithmic
outcomes to enacted maps and to the incumbency-neutral baseline.

The empirical overview across NC, WI, and TX establishes the central quantitative
finding: \textsc{bisect} maps fall within or marginally above the incumbency-neutral
baseline range on all three metrics (the NC open-seat count of 4 slightly exceeds the
$[0,3]$ 95\% CI, a consequence of the pairing effect documented in I.1), while enacted
2022 congressional maps produce lower pairing rates and higher safe-seat fractions than
the baseline. These comparisons use single-seed bisect results ($^\dagger$); multi-seed
ensemble validation is required before treating the gap as a general property of the
algorithm. This gap cannot be explained by
geographic constraints; it requires the hypothesis that enacted mapmakers used incumbent
address data and partisan information to engineer protective boundaries.

The legal framework governing this finding is settled. Courts have consistently held
that incumbency protection is a permissible redistricting criterion, but no court has
required it. Algorithmic indifference to incumbency is legally sound. The \textsc{bisect}
algorithm's failure to protect incumbents is not a constitutional defect; it is a
deliberate design choice to treat incumbency the same way a neutral geographic process
would treat it---as one of the many consequences of drawing compact, population-balanced
districts from a census tract graph, rather than as a target to be optimised.

The companion papers in this series provide the detailed empirical and legal analysis
summarised here. I.1 derives the mathematical expectation for incumbent pairing under
random redistricting and tests whether \textsc{bisect} and enacted maps differ
significantly from that expectation. I.2 documents the safe-seat creation and destruction
dynamics in detail, including the party-specific decomposition. I.3 analyses open-seat
creation and its implications for electoral accountability. I.4 provides a comprehensive
legal survey of incumbency as a redistricting criterion across federal doctrine and state
constitutional frameworks. Together, these papers establish \textsc{bisect} as a
credible, incumbency-neutral baseline for redistricting litigation and legislative
proceedings.
