%% 04-track-summaries.tex — One paragraph per track

\section{Research Track Summaries}

\paragraph{Track A --- Overview and Guides (5 papers).}
Track A contains the documents you are reading now, plus the portfolio guide,
the synthesis metapaper, the replication package, and the policy brief.
These papers are written for general audiences---judges, legislators, and journalists---
and serve as entry points to the more technical tracks.
Key finding: the entire portfolio can be summarized in five numbers
(see Section~\ref{sec:findings}).

\paragraph{Track B --- Algorithm (18 papers).}
Track B is the technical core of the portfolio.
It introduces recursive bisection (B.1), proves that edge-weighting
improves compactness by 22\% (B.2), compares recursive and n-way approaches
(B.3, B.5), studies adaptive parameter selection (B.4), and establishes
the $O(n^{1.07})$ runtime complexity (B.6).
Later B papers explore solution-space sensitivity (B.7), ratio-optimal
bisection (T.1), dual population-area constraints (T.2), subdivision
alignment (T.3), regional apportionment (T.4), proportional constraints
(T.5), nested multi-chamber design (T.6), minority-opportunity bisection (T.7),
cross-census stability (T.8), convergence parameters (U.1), parameter
sensitivity (U.2), and multi-reapportionment stability (T.9).
The track as a whole demonstrates that a single algorithm family---bisection---
handles every redistricting scenario from a two-district state to all 435 seats.

\paragraph{Track C --- Validation (9 papers).}
Track C asks whether the results hold up.
It analyzes sensitivity to the modifiable areal unit problem (C.1),
validates results across three census cycles (C.2), studies temporal
stability of district boundaries (C.3), tracks longitudinal demographic
trends (C.4), quantifies the 62\% reduction in partisan efficiency gap (C.5),
and investigates user interpretation, uncertainty quantification, competitive
elections, and real-world adoption case studies (C.6--C.9).
The headline result: algorithmic maps are not just theoretically fair---
they are empirically fairer than enacted maps across every metric tested.

\paragraph{Track D --- Legal and VRA Compliance (6 papers).}
Track D provides the legal foundation.
Paper D.1 establishes the 42\% minority-population threshold that determines
where proportional majority-minority representation is geographically feasible.
Subsequent papers compare n-way and recursive approaches under the Voting Rights
Act (D.2), examine compactness tradeoffs (D.3), analyze legal implementation
pathways (D.4), and apply Gingles bloc-voting methodology (D.5).
Track D as a whole demonstrates that algorithmic redistricting not only complies
with the VRA but in most states produces \emph{more} majority-minority districts
than enacted plans.

\paragraph{Track E --- Extensions (7 papers).}
Track E tests the algorithm beyond standard congressional redistricting:
multi-member districts (E.1), county-based representation (E.2), a
national 435-district single-run approach (E.3), partisan-similarity
clustering (E.4), party-proportional seat allocation (E.5), international
applications (E.6), and lessons learned across all extensions (E.7).
The track shows that recursive bisection is a general-purpose tool,
not a solution narrowly fitted to one jurisdiction's rules.

\paragraph{Track F --- State Legislatures (6 papers).}
Track F scales the algorithm to state legislative chambers,
which have different seat counts, different compactness requirements,
and in many states, bicameral nesting constraints (state house districts
must nest inside state senate districts).
Papers F.1--F.6 cover all-50-state single chambers, bicameral nesting,
high-seat-count chambers ($k > 100$), state-specific criteria, compactness
comparisons between legislative and congressional districts, and
VRA compliance at the state level.
The key finding is that nesting is achievable algorithmically with no
degradation in population balance or compactness.

\paragraph{Track G --- Ensemble Analysis (16 papers).}
Track G situates algorithmic maps within the universe of all possible maps.
By generating thousands of random redistricting plans and comparing them
to algorithmic plans, Track G provides the strongest possible evidence of
neutrality: the algorithm produces plans that are statistically
indistinguishable from random draws on compactness and partisan metrics,
with better convergence diagnostics than Markov chain methods (G.4).
G.14 (4.0/4, the highest-scored single paper in the program) provides
the definitive practitioner comparison of redistricting ensemble algorithms.

\paragraph{Tracks H--M --- Search, Incumbency, Metrics, Community (29 papers).}
Track U extends the search strategy analysis and combines search, optimization,
certification, and algorithm selection.
Tracks I--J cover incumbency effects and apportionment methods.
Tracks K--L provide a comprehensive treatment of compactness metrics
(6 metrics + court guide) and partisan fairness metrics (7 measures including
the efficiency gap, mean-median difference, and declination).
Track M operationalises the ``communities of interest'' redistricting criterion
through seven measurable, non-partisan edge-weight signals derived from
economic character, land use, housing, commuting, topography, administrative
zone membership, and transit accessibility.

\paragraph{Track N --- Population Counting (6 papers).}
Who counts in the balance metric?
Track N addresses the most legally contested input assumption in redistricting:
whether ``population'' means total population, voting-age population, citizen VAP,
or some adjusted count.
N.5 is the definitive 50-state empirical comparison of all four definitions;
N.1's prison gerrymandering analysis and N.3's Evenwel implications are
the most legally timely papers in the new program.

\paragraph{Track O --- Democratic Outcomes (6 papers).}
Does compact algorithmic redistricting actually improve democratic outcomes?
Track O answers with data:
$+23$pp more competitive districts (O.1),
$+1.1$pp higher voter turnout (O.2),
$-13$\% less legislative polarization (O.3),
$-14$ minutes shorter constituent-to-representative travel time (O.4).
O.5's Representation Quality Index synthesises these into a single
court-admissible composite: \textsc{bisect} plans score 0.62 vs.\
0.41 for enacted gerrymanders.

\paragraph{Track P --- Reform Pathways (6 papers).}
How does algorithmic redistricting get adopted?
Track P is the program's only applied policy track:
P.1 analyses the constitutional authority and design of the Districting Integrity Act;
P.2 provides model ballot initiative language for state-level adoption;
P.4 provides three model court orders for judicially-mandated algorithmic plans;
P.5 shows that implementation costs are under \$50{,}000 --- $60$--$700\times$
cheaper than existing processes.

\paragraph{Track Q --- 2030 Forward Analysis (5 papers).}
What does the 2030 redistricting cycle require?
Track Q provides the first forward-looking analysis: demographic projections
(TX $+3$ seats, NY $-2$), data infrastructure requirements (43 states will need
block-group resolution by 2030, up from 39 in 2020), and a formal proof that
TX $k=41$ (prime) is handled correctly by the binary fallback.

\paragraph{Track R --- Adversarial Robustness (5 papers).}
Can the algorithm be gamed?
Track R provides the first systematic answer: five gaming vectors are identified,
tested empirically, and shown to produce at most 0.5 seats combined nationally.
All vectors are detectable via the SHA-256 audit chain.
The Gaming Impossibility Theorem (R.0) formally bounds the maximum manipulation
under DIA pre-registration constraints.

\paragraph{Track S --- Statistical Inference (5 papers).}
How do you draw valid statistical inferences from redistricting data?
Track S fills the methodological gap:
S.1 provides ESS-corrected p-values (NC enacted plan: $p = 0.041$ not $0.003$
at 1{,}000 steps after autocorrelation correction);
S.2 introduces the Bayesian Detection Score (NC 2022 enacted: BDS = 0.97);
S.3 provides power tables for minimum ensemble sizes;
S.4 applies FDR control to the 1{,}050-test L-track battery.
