\section{Package Model}

An RCOUNT package is a directory whose files separate source evidence,
normalized records, reconciliation declarations, proof sketches, status events,
and transcripts. The current Rust implementation writes these layers as JSON or
NDJSON files so they can be diffed, hashed, and replayed with ordinary tools.

\begin{center}
\small
\begin{tabularx}{\textwidth}{llX}
\toprule
Layer & Required & Purpose \\
\midrule
\texttt{manifest.json} & Yes & Jurisdiction, election metadata, version, and package hash \\
\texttt{sources/} & Yes & Raw or synthetic source evidence and package-relative source index hashes \\
\texttt{normalized/contests.ndjson} & Yes & Contest definitions and valid selections \\
\texttt{normalized/reporting-units.ndjson} & Yes & Precincts, batches, jurisdiction totals, districts \\
\texttt{normalized/summaries.ndjson} & Yes & Contest totals by reporting unit, status, and optional batch \\
\texttt{normalized/batches.ndjson} & Optional & Batch manifests and accepted/counted/rejected ballots \\
\texttt{normalized/lineage.ndjson} & Optional & Reporting-unit split, merge, and unchanged events \\
\texttt{proofs/} & Optional & Privacy-safe inclusion proof sketches and package hashes \\
\texttt{status/events.ndjson} & Yes & Canvass, correction, recount, and certification events \\
\texttt{transcripts/} & Yes & Verifier output and replay evidence \\
\bottomrule
\end{tabularx}
\end{center}

The central record is a summary: a contest id, reporting unit id, optional batch
id, count status, selection totals, undervotes, overvotes, blank contests, and
counted ballots. This is intentionally not a cast-vote-record schema. Summaries
are the public claims being verified; source files and future adapters can
explain how those summaries were derived from vendor exports, statements of
votes, or canvass reports.

Hashes use domain-separated prefixes so that a record hash, source hash, file
hash, proof hash, and package hash cannot be silently substituted for one
another. A package content hash binds the normalized model. Source hashes bind
the raw evidence bytes used to justify that model. This distinction lets a
verifier say whether arithmetic failed, source evidence was tampered with, or a
manifest no longer matches the normalized records.
