\section{Status, Schema, and Hash Contract}

RCOUNT records count status as data rather than prose. A status says what public
claim is being replayed; a status event says why the claim changed.

\begin{center}
\small
\begin{tabularx}{\textwidth}{lXX}
\toprule
Status & Typical evidence & What RCOUNT can verify \\
\midrule
Unofficial & Election-night export or preliminary report & Local arithmetic and source hash stability for the preliminary claim. \\
Canvassed & Canvass report, board minutes, corrected public export & Arithmetic, jurisdiction totals, and declared correction events. \\
Recounted & Recount report or updated statement of votes & Arithmetic and event lineage for the recount snapshot. \\
Amended & Amended certification or court-ordered correction & Arithmetic and public event history for the amended snapshot. \\
Certified & Official certification artifact & Mechanical consistency of records marked certified, not the legal validity of certification. \\
\bottomrule
\end{tabularx}
\end{center}

The schema contract is intentionally small. A package declares
\texttt{rcount\_version}; the current crates use a draft version while the
format is incubating. Required layers are the manifest, contests, reporting
units, summaries, status events, source index, source bytes, package hashes, and
transcripts. Optional layers include batch manifests, lineage events, inclusion
proof sketches, and district aggregation transcripts. Source index paths are
package-relative and must remain under \texttt{sources/}.

Hashes are over canonical JSON projections or exact source bytes with
domain-separated prefixes. The current prefix families are source, record, file,
package, event, and proof. This prevents a hash from one layer from being
silently reused as evidence for another. A district aggregation transcript adds
the RCOUNT package content hash, the RPLAN plan hash, and, when present, the
RCTX context hash. Those hashes bind the district total to both the count
package and the assignment used to project reporting units into districts.

