\section{Verifier Contract}

The verifier contract is deliberately layered. A package may fail because local
contest arithmetic is wrong, because jurisdiction totals do not equal the sum of
their reporting units, because a batch summary has no batch manifest, because a
lineage event references a missing precinct, because a proof exposes voter
choices, or because source bytes no longer match the source index. These are
different failures and should not collapse into a generic ``invalid election''
message.

\begin{center}
\small
\begin{tabularx}{\textwidth}{lXX}
\toprule
Check & Inputs & Failure caught \\
\midrule
\texttt{contest\_selection\_sum} & Contest, summary & Votes plus residuals mismatch ballots \\
\texttt{jurisdiction\_contest\_total} & Summaries & Jurisdiction total disagrees with units \\
\texttt{status\_event\_declared} & Status events & Missing or malformed public event history \\
\texttt{canvass\_correction\_event} & Status events, summaries & Changed canvass lacks correction event \\
\texttt{accepted\_ballots} & Batch manifest & Accepted ballots mismatch counted plus rejected \\
\texttt{batch\_summary\_total} & Batch manifest, summary & Batch summary lacks accounting evidence \\
\texttt{lineage\_conservation} & Reporting units, lineage & Split or merge references missing units \\
\texttt{proof\_privacy\_gate} & Inclusion proof sketches & Proof reveals choices or linkable identity \\
\texttt{source\_hash\_match} & Source index, source bytes & Raw evidence changed after packaging \\
\texttt{district\_aggregation\_total} & RCOUNT, RPLAN & District totals do not replay from units \\
\bottomrule
\end{tabularx}
\end{center}

The current CLI exposes two verifier paths. \texttt{rcount verify} reads a
package directory and emits a pass/fail transcript with per-check results.
\texttt{rcount aggregate-districts} reads an RCOUNT package and an RPLAN
assignment, then emits district totals and an aggregation transcript. The latter
path is intentionally optional: base RCOUNT packages do not require districts,
districting graphs, or RPLAN contexts.

The contract is not a certification standard. A passing transcript means the
package satisfied the declared mechanical checks. It does not prove that every
eligible voter was registered, that every ballot was accepted or rejected
lawfully, that voting-system software was uncompromised, or that a court would
uphold the election. Those questions belong to law, administration, security,
and audit practice. RCOUNT contributes a replayable evidence layer underneath
those decisions.
