\section{Seed Sensitivity Analysis}
\label{sec:seed-sensitivity}

The South Carolina state house result---28 majority-minority districts at state house scale versus 0 possible at congressional scale---constitutes this paper's headline finding. To assess reproducibility, we ran VRASection for seeds $s \in \{42, 43, 44, 45, 46\}$ for all five covered states. A result is considered stable if all five seeds produce the same majority-minority district count.

\subsection{South Carolina Seed Sweep}

Table~\ref{tab:sc-seeds} reports VRASection majority-minority district counts for South Carolina across five seeds.

\begin{table}[h]
\centering
\caption{South Carolina state house VRASection results, 5-seed sweep. Target: 28 majority-minority districts (Black $\geq 50\%$).}
\label{tab:sc-seeds}
\begin{tabular}{lcc}
\toprule
Seed & MM Districts & Target \\
\midrule
42 & 28 & 28 \\
43 & 28 & 28 \\
44 & 28 & 28 \\
45 & 28 & 28 \\
46 & 28 & 28 \\
\midrule
\textbf{Stability} & \textbf{5/5} & \\
\bottomrule
\end{tabular}
\end{table}

South Carolina achieves 5/5 seed stability at exactly 28 majority-minority districts. The stability is expected: VRASection's threshold-based architecture is deterministic given minority population geography and the 42\% threshold. Seed variation affects only the METIS partitioning of non-threshold-sensitive tracts, not the identification and preservation of minority community cores.

\subsection{All Covered States}

Table~\ref{tab:all-seeds} reports seed stability for all five covered states.

\begin{table}[h]
\centering
\caption{5-seed sweep results for all covered states. MM$_{\min}$ and MM$_{\max}$ are the minimum and maximum majority-minority district counts across seeds 42--46.}
\label{tab:all-seeds}
\begin{tabular}{lccccc}
\toprule
State & $k$ & Target MM & MM$_{\min}$ & MM$_{\max}$ & Stable? \\
\midrule
Alabama       & 105 & 27 & 27 & 27 & 5/5 \\
Georgia       & 180 & 47 & 47 & 47 & 5/5 \\
Louisiana     & 105 & 30 & 30 & 30 & 5/5 \\
Mississippi   & 122 & 44 & 44 & 44 & 5/5 \\
South Carolina & 124 & 28 & 28 & 28 & 5/5 \\
\midrule
\textbf{Total} & & & & & \textbf{5/5 all states} \\
\bottomrule
\end{tabular}
\end{table}

The 5/5 covered-state result is fully seed-stable: all five covered states produce identical majority-minority district counts across all five seeds. This confirms that the headline finding is not seed-dependent.

\paragraph{Why such stability?}
The minority population concentrations in the five covered states are geographically robust --- they are not near-threshold cases. The Black population in South Carolina is concentrated in Charleston, Columbia, and the Low Country coastal communities, none of which are near the 42\% threshold; all cores are well above 50\% minority. METIS seed variation affects only how non-concentrated tracts at the margins are assigned, which does not change which cores qualify as majority-minority.
