\section{Lineage Model}

An RCOUNT lineage event relates reporting units across cycles. The core fields
are:

\begin{center}
\small
\begin{tabularx}{\textwidth}{lX}
\toprule
Field & Meaning \\
\midrule
\texttt{lineage\_id} & Stable event identifier within the package. \\
\texttt{kind} & Administrative relation: unchanged, split, or merge in the current draft. \\
\texttt{prior\_cycle}, \texttt{current\_cycle} & Package-level cycle labels connected by the event, such as \texttt{SYN-2024-general}. \\
\texttt{prior\_reporting\_unit\_ids} & Units from the earlier cycle. \\
\texttt{current\_reporting\_unit\_ids} & Units in the later cycle. \\
\texttt{authority} & Public office, board, or source asserting the change. \\
\texttt{explanation} & Human-readable reason, such as growth, consolidation, or redrawn precinct boundaries. \\
\bottomrule
\end{tabularx}
\end{center}

The verifier checks structure before interpretation. A split must name one prior
unit and multiple current units. A merge must name multiple prior units and one
current unit. Every referenced unit must exist in the package's reporting-unit
records. Those checks do not prove that the county drew the new units correctly;
they prove that the declared cross-cycle relationship is internally replayable.
The current crate uses unchanged, split, and merge events; rename and boundary
adjustment events are future lineage kinds, not implied by this draft.

Cycle ids are scoped to the package family being compared. They are stable
labels for audit transcripts, not national election identifiers. Reporting-unit
ids are likewise scoped to the RCOUNT package and its source adapters.

Lineage is distinct from count totals. The lineage event says how units relate;
the summaries still carry contest, status, residual, and counted-ballot
arithmetic. This separation keeps an administrative crosswalk from silently
rewriting votes. The check named \texttt{lineage\_conservation} conserves the
declared unit relationship: it validates split/merge cardinality and referenced
unit existence. It does not assert that vote totals are conserved across cycles.

Public source artifacts for lineage can vary by jurisdiction. Examples include
precinct-change orders, board records, county GIS releases, public precinct
notices, statements of votes, and source crosswalks. RCOUNT normalizes the
lineage record and binds public source references where they are available.
