\section{Multi-Election Harness}

The L2 synthetic harness spans three cycles: 2024, 2026, and 2028. Each cycle
has its own RCOUNT package, RPLAN plan hash, unit universe hash, and district
aggregation transcript. The point is not that the synthetic elections are
realistic. The point is that every cross-cycle comparison names the exact count
package and plan universe that produced it.

\begin{center}
\small
\begin{tabularx}{\textwidth}{lrrX}
\toprule
Cycle & Units & Lineage events & Unit universe \\
\midrule
2024 & 3 & 0 & \texttt{P-001}, \texttt{P-002}, \texttt{P-003} \\
2026 & 4 & 1 & \texttt{P-001A}, \texttt{P-001B}, \texttt{P-002}, \texttt{P-003} \\
2028 & 3 & 1 & \texttt{P-001A}, \texttt{P-001B}, \texttt{P-023} \\
\bottomrule
\end{tabularx}
\end{center}

For district aggregation, each cycle binds three hashes:
\[
  H_{\mathrm{count\ package}},\quad
  H_{\mathrm{rplan\ plan}},\quad
  H_{\mathrm{unit\ universe}}.
\]
In the current transcript, the unit-universe hash binds the sorted reporting
unit ids that the aggregation used for that cycle. It is not a substitute for
the package hash or plan hash; it is a compact guard that says which unit set
the bridge actually saw.
That triple prevents a stale crosswalk from being silently reused against a
different set of reporting units. If the plan expects P-002 but the current
package only contains P-023 after a merge, the failure belongs to the plan/unit
boundary, not to vote arithmetic.

This L2 harness is where RCOUNT and RPLAN meet. Base lineage verification does
not require RPLAN. District aggregation does, because the district totals depend
on a particular plan assignment over a particular reporting-unit universe.
