\section{Boundaries}

RCOUNT lineage does not prove why voters moved, whether turnout changed, or
whether a precinct alteration had political intent. It only states the
administrative relationship between reporting units and verifies that the
declared relationship is structurally consistent. A passing lineage check is
not legal approval of a precinct change.

It also does not require a county to publish voter-level movement records.
Lineage works at the reporting-unit level, not the individual-voter level. When
future adapters use voter-file, assignment, or registration evidence, they must
preserve the V.0 privacy boundary: public verification should not expose voter
identity, ballot style, timestamps, or small-cell facts that allow
reidentification. Private support evidence can be hashed and retained by an
authority without being published; a hash does not sanitize sensitive source
data.

Finally, lineage is not a substitute for RPLAN. RCOUNT can declare how reporting
units changed; RPLAN declares how those units aggregate into districts for a
specific plan and cycle. The bridge between them must remain hash-bound and
optional.

\begin{center}
\small
\begin{tabularx}{\textwidth}{XX}
\toprule
RCOUNT lineage enables & RCOUNT lineage does not prove \\
\midrule
Replayable split, merge, and unchanged-unit evidence. & Legal validity or political neutrality of a precinct change. \\
Cycle-specific unit-universe checks. & Individual voter movement. \\
Failure attribution for bad lineage, stale plans, and source tampering. & That turnout or vote shifts have any particular cause. \\
\bottomrule
\end{tabularx}
\end{center}
