\documentclass[12pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{lmodern}
\usepackage[T1]{fontenc}
\usepackage{amsmath, amssymb, amsthm}
\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}
\newtheorem{proposition}{Proposition}
\newtheorem{corollary}{Corollary}[proposition]
\newtheorem{lemma}{Lemma}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage{booktabs}
\usepackage{xcolor}
\usepackage{algorithm}
\usepackage{algpseudocode}

\title{H.1: BisectionEnsemble:\\
       Integrating Local Feasibility Sampling into the Redistricting Bisection Tree}

\author{Gio Della-Libera}

\date{Draft --- \today}

\begin{document}

\maketitle

\begin{abstract}
Markov-chain redistricting ensemble methods (ReCom, GerryChain) sample the full
plan space by running $k$-way recombination steps, but exhibit catastrophic
bipartition failure at large $k$: Pennsylvania ($k=17$, prime) requires a
17-way bipartition of the whole state; Texas ($k=38=2\times 19$) requires a
19-way bipartition at the first ApportionRegions split.
These large-arity bipartitions have no efficient graph-theoretic decomposition
and cause GerryChain cold-start failures. We introduce \emph{BisectionEnsemble}, which resolves this failure
by repositioning ensemble sampling inside the bisection tree rather than around
it. At each node of the bisection tree---which decomposes a $k$-district state
into a hierarchy of 2-way cuts---we run a local $T$-step ReCom chain on the
$\approx n/k$-tract subregion and select the bisection cut at percentile $p$ of
the collected edge-cut distribution. Since every local ensemble is always 2-way
($k_{\text{local}} = 2$), bipartition failure at the local ReCom level is structurally impossible.
The bisection tree structure (ApportionRegions, GeoSection) is fully preserved.
We implement \texttt{SeedCompositor::BisectionEnsemble\{p, ensemble\_steps\}} in
\texttt{redist-cli} using \texttt{split\_subgraph\_bisection\_ensemble()}, which
seeds a METIS bisection as the Markov chain's initial state, runs
\texttt{ensemble\_steps} ReCom steps via \texttt{redist\_ensemble::recom::RecomChain},
collects accepted plans with their edge cuts, and returns the plan at rank
$\lfloor p \times \text{accepted\_count} \rfloor$.
Empirical evaluation on NC ($k=14$), WI ($k=8$), and TX ($k=38$) shows that
BisectionEnsemble(p=0.5, T=100) improves Polsby-Popper compactness by
5--12\% over standard METIS bisection, matches or improves partisan proportionality,
and crucially, succeeds on TX where direct GerryChain $k=38$ fails with bipartition errors.
Runtime is $O(T \cdot n \log n)$ total and trivially parallelisable across tree nodes.
\end{abstract}

\input{sections/01-introduction}
\input{sections/02-related-work}
\input{sections/03-algorithm}
\input{sections/04-results}
\input{sections/05-legal-properties}
\input{sections/06-conclusion}

\bibliographystyle{plainnat}
\bibliography{references}

\end{document}
