\section{Evaluation Plan}

The current evidence ladder is fixture and staged exact-contract evidence.
Required L0/L1 checks include formulation construction, cut-report emission,
fallback status clarity, and RPLAN sidecar verification. A public U.16 golden
package records the branch-and-cut family handoff in the package corpus. A
synthetic path8 benchmark package now adds a solved exact branch-and-cut
artifact, including the CPLEX-LP model hash, solve report, audit certificate,
and verifier manifest. Real-data exact performance and broad runtime claims
remain future benchmark evidence.

\begin{table}[htbp]
\centering
\small
\begin{tabular}{p{0.26\linewidth}p{0.31\linewidth}p{0.29\linewidth}}
\toprule
Claim & Current evidence & Publication gap \\
\midrule
Formulation reports are reproducible & L0 formulation fixtures and the path8
benchmark LP/report artifact & larger model corpus \\
Fallback status is explicit & L0/L1 status tests & failure-mode examples \\
RPLAN sidecars verify lineage & L1 audit package checks and public golden
package plus solved path8 benchmark & larger solver-produced instance \\
Exact performance scales & future benchmark placeholder & real-data benchmark protocol \\
\bottomrule
\end{tabular}
\caption{Evidence ladder for U.16 claims.}
\end{table}

\begin{table}[htbp]
\centering
\small
\begin{tabular}{p{0.30\linewidth}p{0.56\linewidth}}
\toprule
Command or package & Evidence produced \\
\midrule
Path8 generator example
& Builds the public path8 benchmark package with LP artifact, solve report,
manifest, RPLAN/RCTX pair, and audit certificate. \\
RPLAN certificate verifier
& Checks that the exported plan and context satisfy the declared certificate
profile and lineage expectations. \\
BISECT manifest verifier
& Checks method transcript, ILP audit summary digest, solve-report directory,
and manifest-sidecar consistency. \\
\bottomrule
\end{tabular}
\caption{Audit commands that define the current U.16 evidence path.}
\end{table}

The concrete generator command is:
\begin{verbatim}
cargo run -p bisect-cli --no-default-features \
  --example u16_branch_cut_path8_benchmark_package
\end{verbatim}
The corresponding verification commands target the generated certificate and
manifest JSON files.

The path8 package is intentionally small. It is a smoke-test instance for the
exact-search contract, not a scaling result. A paper that claims real-data
branch-and-cut performance should add solver name and version, hardware, time
limit, optimality gap, instance metadata, full transcripts, and archived output
directories for every reported instance.
