\section{Implementation Contract}

The implementation contract distinguishes formulation-only paths, active
branch-and-cut paths, solver fallback paths, and completed proof/bound paths.
Each run summary should name the model, objective, constraint profile,
separation routine status, solve status, incumbent plan source, and any proof
or bound metadata.

The report is the boundary between implementation success and exactness. A run
can be successful as software because it produced a valid audited plan, while
still failing to support an exact optimization claim because the solver timed
out, fell back, or never produced a bound. U.16 therefore requires every output
to carry a status label before any exactness language is used.

The atlas framing makes the separated candidate concrete. A candidate incumbent
must be inspectable as a graph assignment; if a district appears in disconnected
components, the separation routine emits the violated connectivity cut and the
model is resolved. The paper's exactness language therefore belongs to the
solver report, while final-plan validity belongs to the RPLAN/RCTX certificate.

For example, on a five-vertex path with two districts, the assignment
\[
  [0,1,0,1,1]
\]
places district 0 on the separated vertices 0 and 2, with vertex 1 assigned to
the other district between them. A separation round does not describe this as
``probably bad'' or silently repair it. It emits a structured round: schema
\texttt{branch-and-cut-separation-round-v1}, status
\texttt{violated-connectivity-cuts}, cut count 4, and disconnected districts
\([0,1]\). The next solve is then performed under the added cuts. If the
process ends with a proof, the run may claim a model-relative exact result; if
it ends by timeout or fallback, the same final plan may still be auditable but
the exactness claim is withheld.

\begin{table}[htbp]
\centering
\small
\begin{tabular}{p{0.25\linewidth}p{0.31\linewidth}p{0.30\linewidth}}
\toprule
Field & Purpose & Failure mode avoided \\
\midrule
Formulation id & Names variables and constraints & Ambiguous model claim \\
Cut summary & Records separation activity & Hidden non-branch-and-cut path \\
Solver status & Distinguishes solved, timed out, fallback & False exactness \\
Proof/bound & Carries exact evidence when available & Valid plan mistaken for proof \\
\bottomrule
\end{tabular}
\caption{Minimum branch-and-cut report fields.}
\end{table}

\begin{table}[htbp]
\centering
\small
\begin{tabular}{p{0.24\linewidth}p{0.30\linewidth}p{0.32\linewidth}}
\toprule
Artifact field & Current value family & Interpretation \\
\midrule
Report schema & \texttt{ilp-solve-report-v1} & Stable JSON envelope for solver evidence \\
Model format & \texttt{cplex-lp} & Rebuildable model serialization \\
Model digest & SHA-256 digest & Tamper check for the LP file \\
Proof status & proved-optimal, bounded-gap, timed-out, not-solved & Exactness vocabulary \\
Fallback flag & true/false & Whether a non-ILP plan source was needed \\
Search strategy & e.g. \texttt{k2-branch-and-bound} & Solver/search family actually used \\
\bottomrule
\end{tabular}
\caption{Concrete solve-report fields used by the U.16 implementation.}
\end{table}

The package-level lineage repeats the same boundary. A BISECT manifest records
the \texttt{bisect-ilp-lineage-v1} schema, the ILP solve-report directory, the
audit summary path, the audit-summary digest, the fallback flag, and the
configured optimality gap. Those fields make the branch-and-cut path observable
after the PDF, plan, and context have been copied away from the machine that
produced them.

\begin{table}[htbp]
\centering
\small
\begin{tabular}{p{0.24\linewidth}p{0.34\linewidth}p{0.28\linewidth}}
\toprule
Status & Meaning & Allowed claim \\
\midrule
Formulation-only & Model built, no active separation proof & reproducible model contract \\
Cut-active & Separation ran with cut metadata & branch-and-cut behavior observed \\
Bounded & Solver reports incumbent and bound & scoped bound under model \\
Proven & Solver reports proof under assumptions & scoped exact optimum/infeasibility \\
Fallback & Heuristic or prior plan returned & valid audited fallback only \\
\bottomrule
\end{tabular}
\caption{Status vocabulary for branch-and-cut outputs.}
\end{table}
