\subsection{Partisan Similarity Metric}

For each pair of adjacent census tracts $(u, v)$, we define the partisan
similarity edge weight as:
\begin{equation}
  w_{uv}^{\text{PSD}} = \exp\!\bigl(-\alpha \cdot |D_u - D_v|\bigr)
  \label{eq:psd-weight-full}
\end{equation}
where $D_u \in [0,1]$ is the Democratic share of the two-party presidential
vote in tract $u$ and $\alpha \geq 0$ is the similarity scaling parameter.

When $\alpha = 0$: $w_{uv}^{\text{PSD}} = 1$ for all edges --- pure
unweighted partitioning, which serves as the lower bound on partisan
homogeneity.  As $\alpha$ increases, the weight function penalises
edges connecting tracts with different partisan leans.  METIS minimises
total weighted edge cuts, so it prefers to cut edges with small weights (i.e.,
dissimilar tracts) and preserve edges with large weights (similar tracts).
The result is districts where all member tracts have similar partisan lean.

Note that equation~\eqref{eq:psd-weight-full} attracts \emph{similar} tracts
regardless of which party they lean toward.  A heavily Republican tract (e.g.,
$D_u = 0.30$) will have high similarity weight to another Republican-leaning
tract ($D_v = 0.35$) and low similarity weight to a Democratic-leaning tract
($D_v = 0.65$).  The resulting PSD map will contain both heavily Republican
and heavily Democratic safe districts; the distribution of seat types between
the parties is determined by geographic sorting, not by the algorithm's
parameters.

\subsection{Parameter Selection}

We test four values of $\alpha$: 0 (unweighted), 5, 20, and 50.  At $\alpha
= 5$, a pair of tracts with a 10-percentage-point difference in Democratic
vote share ($|D_u - D_v| = 0.10$) receives weight $\exp(-5 \cdot 0.10) =
e^{-0.5} \approx 0.61$, while a pair with a 30-point difference receives
$e^{-1.5} \approx 0.22$.  At $\alpha = 50$, the same pairs receive weights
$e^{-5} \approx 0.007$ and $e^{-15} \approx 3 \times 10^{-7}$---near-zero,
effectively prohibiting the edge from being preserved in the partition.

The transition from ``moderate'' ($\alpha = 5$) to ``strong'' ($\alpha = 20$)
partisan similarity represents the approximate threshold at which the within-
district homogeneity benefit exceeds the compactness cost, as measured by
mean Polsby-Popper.

\subsection{Implementation Details}

PSD weighting is implemented in the bisect \texttt{weights-override} framework
under the key \texttt{partisan-similarity}.  The edge-weight computation
requires tract-level partisan data, which is loaded from the election data
layer alongside census demographics.

\begin{verbatim}
bisect state --state NC --year 2020 --version E4_psd \
  --weights-override partisan-similarity \
  --partisan-alpha 20.0 --seed 42
\end{verbatim}

The partisan data is the 2020 presidential election results disaggregated to
census tracts from the MIT Election Data Science Lab \citep{MEDSL2020}.

\subsection{Metrics}

We evaluate PSD maps on four metrics:

\paragraph{Within-district partisan standard deviation ($\sigma_D$).}
For each district, compute the standard deviation of tract-level Democratic
vote shares weighted by tract population.  A lower $\sigma_D$ indicates a
more homogeneous district.  We report the mean $\sigma_D$ across all districts
in the map.

\paragraph{Safe seat count.}
The number of districts where the district-level Democratic vote share exceeds
60\% (safe Democratic) or falls below 40\% (safe Republican).  We report
counts separately for each party.

\paragraph{Efficiency Gap (EG).}
Computed as described in Section~2, using 2020 presidential election results
as the vote-share input.

\paragraph{Polsby-Popper compactness (PP).}
Mean Polsby-Popper across all districts, as a measure of the compactness cost
of PSD weighting.

\subsection{States}

We apply PSD weighting to North Carolina ($k=14$) and Texas ($k=38$), chosen
because both have been the subject of major partisan gerrymandering litigation
and have well-documented enacted maps with significant partisan asymmetry.
We compare PSD maps to: (1) the 2020 enacted congressional plans, and (2)
neutral bisect maps generated with geographic edge-weighting (standard mode).

All results are single-run (seed 42).\singrun
