\section{Results}

\subsection{Five-State Comparison}

Table~\ref{tab:perm-results} reports ESS-corrected p-values for both
the enacted 2022 congressional maps and the \textsc{bisect} algorithmic plans,
across all five study states, at two ensemble sizes.

\begin{table}[h]
\centering
\caption{ESS-corrected p-values from permutation test on compactness metric.
``Sig.'' = significant at $\alpha = 0.05$ (ESS-corrected).
G.1: 1{,}000-step ensemble; G.4: 10{,}000-step ensemble.}
\label{tab:perm-results}
\begin{tabular}{lcccccc}
\toprule
& \multicolumn{3}{c}{\textbf{Enacted 2022}} & \multicolumn{3}{c}{\textbf{Bisect}} \\
State & $\hat{p}$ & $p_{\rm corr}$ (1K) & $p_{\rm corr}$ (10K) & $\hat{p}$ & $p_{\rm corr}$ (1K) & $p_{\rm corr}$ (10K) \\
\midrule
NC  & 0.003 & 0.041$^*$ & 0.004$^{**}$ & 0.50 & 0.52 & 0.51 \\
WI  & 0.002 & 0.027$^*$ & 0.003$^{**}$ & 0.50 & 0.51 & 0.50 \\
GA  & 0.001 & 0.018$^*$ & 0.002$^{**}$ & 0.50 & 0.51 & 0.50 \\
PA  & 0.002 & 0.032$^*$ & 0.004$^{**}$ & 0.50 & 0.52 & 0.51 \\
TX  & 0.004 & 0.11     & 0.03$^*$     & 0.40 & 0.43 & 0.41 \\
\bottomrule
\end{tabular}

\smallskip
\footnotesize
$^*$ $p < 0.05$; $^{**}$ $p < 0.01$. All ESS from G.4 corrected table.
1{,}000-step ESS: NC=70, WI=81, GA=64, PA=58, TX=42.
10{,}000-step ESS: NC=695, WI=811, GA=638, PA=582, TX=417.
\end{table}

\paragraph{Key findings.}
\begin{enumerate}
  \item \textbf{ESS correction matters substantially for NC at 1K steps.}
    The unadjusted p-value ($\hat{p} = 0.003$) suggests extreme significance,
    but the ESS-corrected $p = 0.041$ is barely significant at 5\%.
    Courts and practitioners using 1{,}000-step ensembles for NC must account
    for this correction.
  \item \textbf{10{,}000-step ensembles provide reliable significance.}
    At $n = 10{,}000$, all four clear gerrymanders (NC, WI, GA, PA) achieve
    $p < 0.01$ with ESS correction. Texas achieves $p = 0.03$ --- borderline
    but significant.
  \item \textbf{Bisect plans are consistent with neutral draws.}
    All five \textsc{bisect} plans have ESS-corrected $p > 0.40$, consistent
    with the null hypothesis (no evidence the plan is systematically extreme).
  \item \textbf{Texas is the hardest case.}
    TX needs 10{,}000 steps for reliable significance. This confirms S.3's
    power analysis recommendation of $\geq 30{,}000$ steps for large-$k$ states.
\end{enumerate}
