Declination ($\delta$), introduced by \citet{warrington2018}, is a geometric measure of
packing and cracking that operates directly on the vote-share diagram. Given sorted district
Democratic vote shares, it computes the angle between Democratic-won districts' centroid and
Republican-won districts' centroid relative to the 50\% threshold: positive $\delta$ indicates
Republican advantage through cracking (Democratic voters spread just below 50\%) or packing
(excess Democratic voters concentrated far above 50\%). The metric has an appealing geometric
transparency---it can be visualized directly on a plot of sorted district vote shares---and
it captures packing-cracking patterns that EG may miss when offsetting patterns in each party's
won districts produce a moderate aggregate wasted-vote count.

We find that \textsc{bisect} algorithmic maps produce $|\delta| < 0.10$ for NC ($k=14$),
WI ($k=8$), and TX ($k=38$)---compared to enacted NC $\delta \approx +0.35$, enacted WI
$\delta \approx +0.28$, and enacted TX $\delta \approx +0.40$---because compactness-driven
bisection does not systematically place Democratic voters just below 50\%, the geographic
signature of cracking. The reduction (70--80\%) is the largest of all six L-series metrics,
reflecting that Declination is most sensitive to the precise boundary configurations that
compactness most directly prevents.

\textbf{Legal status:} Declination has been accepted as evidence by no federal or state court
as of 2026. It was introduced as an exhibit in \textit{Common Cause v.\ Rucho} (M.D.N.C.\
2018) but the court did not rely on it. Subsequent NC, PA, and OH redistricting opinions
cite EG, MM, and Bias without mentioning declination. Expert witnesses should not lead with
declination as a standalone metric and must be prepared to defend against three methodological
critiques: (1) undefined when one party wins all seats; (2) sensitive to the 50\% seat-win
threshold; (3) no established unconstitutionality threshold. This paper provides the first
systematic analysis of declination under algorithmic redistricting and documents its legal status.
