\documentclass[12pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{amsmath, amssymb, amsthm}
\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}
\newtheorem{proposition}{Proposition}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}[proposition]
\newtheorem{remark}{Remark}
\newtheorem{conjecture}{Conjecture}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage{booktabs}
\usepackage{lmodern}
\usepackage[T1]{fontenc}
\usepackage{microtype}
\usepackage{xcolor}
\usepackage{multirow}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{array}
\usepackage{listings}
\lstset{breaklines=true, columns=fullflexible}

\title{StabilitySection: Cross-Census Stability of\\
       GeoSection-Optimal Redistricting Maps}

\author{Gio Della-Libera}

\date{Draft --- \today}

\begin{document}

\maketitle

\begin{abstract}
StabilitySection is not a new map constructor.  It is a cross-census audit
framework for asking whether GeoSection selects the same first bisection,
similar district assignments, and similar outcome diagnostics when the input
census year changes.
We define the \textbf{Census Stability Score} (CSS), a composite metric for
recording this cross-census robustness, and separate three levels of evidence:
implemented GeoSection runs, tract-split proxy comparisons, and future
district-level assignment matching.

The paper's central claim is methodological rather than legal-dispositive.
When the same geography-first procedure returns the same first-level structure
under multiple census snapshots, that is evidence that the structure is not an
artefact of one vintage of population data.  It does not by itself prove that a
map is legally required, partisan-neutral, or immune from challenge.  It gives
courts, commissions, and researchers an auditable way to distinguish stable
geographic structure from census-vintage sensitivity.

We present the 2020 GeoSection baseline, a 2000--2010 tract-split proxy
comparison for 30 comparable state cases, and the CSS computation plan for the
full 2000/2010/2020 package.  The current empirical result is deliberately
narrow: under a 5 percentage point first-split threshold, 20 of 30 comparable
cases are stable in the proxy table.  Full three-census CSS, district Jaccard
matching, and final-plan partisan comparisons remain Phase~2 evidence.
\end{abstract}

\input{sections/01-introduction}
\input{sections/02-related-work}
\input{sections/03-methodology}
\input{sections/04-evaluation}
\input{sections/05-discussion}
\input{sections/06-conclusion}

\bibliographystyle{plainnat}
\bibliography{references}

\end{document}
