\section{Attack Surface 3: ACS Demographic Tables}
\label{sec:acs}

\subsection{The Data Format}

American Community Survey (ACS) 5-year tables provide demographic
estimates for census tracts, including:
\begin{itemize}
\item B02001: Race alone
\item B03002: Hispanic or Latino origin by race
\item B05001: Nativity and citizenship status
\end{itemize}

\bisect{} uses ACS tables to compute the VRA boost weight $w_{\rm vra}$
for census tracts in states with significant minority populations.
The boost weight increases the edge weight between tracts that are both
in VRA-eligible minority communities, encouraging the bisection to keep
such communities intact.

\subsection{The Attack}

An adversary could manipulate ACS minority population estimates to:
\begin{enumerate}
\item Artificially inflate minority percentages in tracts in non-competitive
  areas, spreading VRA boost weights broadly and diluting their effect
  in areas where minority community protection matters.
\item Artificially deflate minority percentages in tracts in competitive
  areas with high minority concentrations, weakening the VRA boost and
  allowing the bisection to divide minority communities.
\end{enumerate}

\subsection{Effect Magnitude}

From U.2's $w_{\rm vra}$ sweep~\citep{b17paper}, the partisan effect
of VRA boost weight variation across its full range is $\Delta D \leq 0.2$
seats nationally.
An ACS manipulation that effectively sets $w_{\rm vra}$ to zero for
specific tracts would be at most as consequential as removing VRA boost
entirely, which U.2 shows produces $\leq 0.2$ seats of shift.

More precisely: ACS manipulation affects only the weight assigned to
VRA-eligible edges.
From U.2 Table~5, the total VRA boost effect ranges from $w_{\rm vra} = 0.2$
(minimal) to $w_{\rm vra} = 0.6$ (strong).
ACS manipulation can simulate any value in this range.
The maximum effect is therefore $\Delta D \leq 0.2$ seats.

Combining with the saturation argument, ACS manipulation contributes
$\Delta D \leq 0.1$ seats to the combined attack, as listed in
Table~\ref{tab:attacks}.

\subsection{Detection}

\texttt{bisect fetch} computes SHA-256 hashes for all ACS tables used.
ACS tables are also available for independent download at:
\begin{verbatim}
https://www2.census.gov/programs-surveys/acs/
\end{verbatim}
Detection probability: $\approx 1.00$.
