\section{Combined Effect Analysis}
\label{sec:effect}

\subsection{Individual Bounds}

From the preceding sections:
\begin{itemize}
\item Population manipulation ($\pm 1\%$ per tract): $\Delta D \leq 0.4$ seats
\item Shapefile adjacency manipulation: $\Delta D \leq 0.4$ seats
\item ACS demographic manipulation: $\Delta D \leq 0.1$ seats
\end{itemize}

\subsection{Why the Combined Effect Does Not Exceed 0.4 Seats}

All three attacks share the same geometric bottleneck: competitive-district
geography.
A competitive district is one where the Democratic vote share is within
3--5 percentage points of 50\%.
In 2020, there were approximately 45 competitive congressional districts
nationally.

Each of the three attacks can, at most, flip one or two of these districts.
But flipping the same district via two different attacks does not double
the effect---a district either flips or it doesn't.
The attacks are therefore largely substitutes, not complements.

Furthermore, the three attacks require different kinds of data access:
population manipulation requires write access to P.L.\ 94-171 files;
shapefile manipulation requires write access to TIGER/Line files;
ACS manipulation requires write access to ACS tables.
An adversary who has write access to all three simultaneously has
effectively compromised the entire data pipeline---which is already
detected by the audit chain.

\subsection{The $\pm 1\%$ Threshold}

We use $\pm 1\%$ as the attack magnitude because:
\begin{itemize}
\item It is within the Census Bureau's acknowledged margin of error
  for tract-level population estimates (differential privacy noise
  at the block level, aggregated to tracts).
\item It is small enough that a manipulation of this magnitude might be
  attributed to legitimate data correction rather than intentional gaming.
\item It is precisely the threshold at which the effect is still within
  B.7's noise floor.
\end{itemize}

For manipulations $> 1\%$ per tract, the effect grows approximately
linearly until saturation at competitive-district thresholds.
At $\pm 5\%$ per tract, $\Delta D$ might reach 1--2 seats nationally,
but such a manipulation would be easily detectable through comparison
with independently-sourced Census data.

\subsection{Below the Noise Floor}

B.7~\citep{b7paper} establishes that the seed-variance noise floor for
\bisect{} is approximately $\sigma \approx 0.8$ seats nationally.
The combined V2 attack maximum of 0.4 seats is within $0.5\sigma$ of zero---
indistinguishable from natural algorithmic variation.

This means that even a successful (undetected) combined V2 attack would
produce partisan outcomes that cannot be confidently attributed to the
attack rather than to legitimate algorithmic variation.
The attack is both bounded and non-attributable.
