\section{Background}
\label{sec:background}

\subsection{Fair Division and the Moving-Knife Protocol}

Fair-division theory addresses the problem of allocating a divisible resource
among multiple claimants such that each claimant regards the allocation as
acceptable by some criterion --- most commonly \emph{proportionality} (each
party receives at least $1/n$ of the whole) or \emph{envy-freeness} (no party
prefers another party's share to its own).

The moving-knife protocol was introduced by Dubins and
Spanier~\citeyear{dubins1961} as a continuous-time generalisation of the
``cut-and-choose'' two-party procedure.
In cut-and-choose, one party divides the resource into two pieces they value
equally, and the other party chooses the piece they prefer.
The procedure is proportional but not envy-free when the ``cutter'' and
``chooser'' value the resource differently.

Dubins and Spanier's moving-knife protocol achieves both properties
simultaneously for two parties:
\begin{enumerate}
\item A referee moves a knife across the resource at a uniform rate.
\item The first party calls ``stop'' when the portion to the left of the knife
  reaches $1/2$ of that party's total valuation.
\item The left portion is awarded to the calling party; the right portion
  goes to the other party.
\end{enumerate}
The protocol is \emph{procedurally impartial}: the knife sweep is mechanical
and neither party controls its rate or direction.
Rawlsian fairness requires that procedures be chosen from behind a ``veil of
ignorance'' about which position one will occupy; the moving-knife protocol
embodies this principle, since neither party knows in advance whether they
will be the caller or the recipient.

Puppe and Tasnadi~\citeyear{puppe2026} adapt the moving-knife framework to
redistricting by reinterpreting ``the resource'' as the state's geographic
territory and ``the knife'' as a directional sweep line.
Instead of moving across the state at a fixed rate, the knife rotates through
angular orientations.
At each orientation, the ``stop'' criterion is replaced by the population
balance condition: the split point is wherever the cumulative population
first reaches $P_{\mathrm{total}}/2$.
The ``value'' being optimised is Reock compactness: the algorithm stops
not when either party calls stop, but at the orientation that maximises
the geometric quality of the split.

\subsection{Reock Compactness}
\label{subsec:reock}

Reock compactness was introduced by Ernest Reock in 1961 as a
``measure of the shape of geographic districts''~\citep{reock1961}.
The measure is defined for a district $D$ as:
\begin{equation}
  \reock{}(D) = \frac{A(D)}{A(\mathrm{MEC}(D))},
  \label{eq:reock}
\end{equation}
where $A(D)$ is the area of district $D$ and $A(\mathrm{MEC}(D))$ is the
area of the smallest circle that encloses $D$ (the minimum enclosing circle,
or MEC).
Since $D \subseteq \mathrm{MEC}(D)$ by definition, we have $\reock{}(D) \in (0, 1]$,
with equality iff $D$ is a perfect circle.

\begin{remark}[Centroid approximation]
\label{rem:centroid-approx}
The precise Reock score of a district requires the MEC of the district's
full polygon boundary.
In the Phase~1 implementation described in this paper, we compute the MEC of
the projected tract centroid points only (not the full polygon vertices).
This approximation may yield $\reock{}_{\mathrm{approx}}(D) > 1.0$ for districts
that contain large-area boundary tracts whose polygon vertices extend beyond
the centroid-MEC.
The implementation clamps all Reock scores to $[0.0, 1.0]$: \linebreak
$\reock{}_{\mathrm{clamped}} = \min(\reock{}_{\mathrm{approx}}, 1.0)$.
The invariant $\reock{} \le 1$ is enforced by clamping, not by structural
guarantee of the centroid approximation.
Phase~2 will implement the polygon-boundary MEC.
\end{remark}

\paragraph{Legal prominence.}
Reock is the most widely cited compactness measure in redistricting
litigation for two reasons.
First, it is intuitively interpretable: a district with Reock $= 0.40$ uses
40\,\% of the smallest circle that could contain it.
Second, elongation and protrusion --- the most common forms of partisan
packing and cracking --- directly reduce Reock, because elongated shapes
have large MECs relative to their area.
Polsby-Popper is sensitive to perimeter irregularity (``jagged'' edges from
following county or precinct boundaries) even in compact districts; Reock is
not, because it depends only on the overall extent of the district, not its
boundary smoothness.

\subsection{Welzl's Minimum-Enclosing-Circle Algorithm}
\label{subsec:welzl}

Given a set of $m$ points in the plane, Welzl's algorithm~\citep{welzl1991}
computes the smallest enclosing circle in expected $O(m)$ time via a
randomised incremental procedure.
The algorithm maintains a candidate MEC and processes points one at a time.
If a new point lies outside the current MEC, the point is added to the
\emph{basis} (the boundary-defining set) and the MEC is recomputed from the
updated basis.
In the redistricting context, the input points are the projected centroids
of the tracts in the half being evaluated; the basis has at most three points
(the MEC of a set of 2D points has at most 3 boundary points).

\begin{proposition}[Welzl Runtime~\protect\citep{welzl1991}]
\label{prop:welzl}
Welzl's algorithm computes the MEC of $m$ points in $\mathbb{R}^2$ in
expected $O(m)$ time.
\end{proposition}

The expected-$O(m)$ bound makes Welzl's algorithm well-suited to the inner
loop of \mka{}: it is called $2 \times n_{\mathrm{orient}}$ times per bisection
node (once for each half at each orientation), contributing $O(n_{\mathrm{orient}} \times m)$
total work.

\subsection{Relationship to Existing Platform Algorithms}
\label{subsec:related}

Table~\ref{tab:algorithm-compare} places \mka{} in the context of the four
main structure algorithms in the redistricting platform.

\begin{table}[t]
\centering
\caption{%
  Structure algorithm comparison.
  \mka{} is the only algorithm that directly targets Reock; the others
  optimise different metrics and achieve Reock as a secondary consequence.
  Runtime is per bisection node with $m$ tracts.
}
\label{tab:algorithm-compare}
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{@{}lllll@{}}
\toprule
Algorithm & Optimises & Distance & Reock & Runtime \\
\midrule
\metis{}  & Edge cuts & Graph topology & Low & $O(m \log m)$ \\
\cvd{}    & Proximity & BFS hop-count  & Medium & $O(n_{\mathrm{iter}} \times m)$ \\
\cvdgeo{} & Proximity & Euclidean      & Medium--high & $O(n_{\mathrm{iter}} \times m)$ \\
\bfsg{}   & Balance speed & Euclidean seeds & Medium & $O(m \log m)$ \\
\textbf{\mka{}} & \textbf{Reock} & \textbf{Euclidean sweep} & \textbf{Highest} & $O(n_{\mathrm{orient}} \times m)$ \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{\mka{} as the third geometric algorithm.}
\cvdgeo{} (B.22b) and \bfsg{} (B.23~\citep{dellaLibera2026bfs}) each
introduced a new geometric structure algorithm targeting a distinct court-relevant
metric: \cvdgeo{} targets Polsby-Popper via geographic proximity clustering;
\bfsg{} targets geographic coverage balance.
\mka{} is the third, targeting Reock via angular sweep.
The three algorithms form a family: each uses Albers EPSG:5070 projection for
geographic coordinates, each shares the 200-iteration boundary-swap rebalance,
and each provides a distinct plain-language justification for the submitted plan.

\paragraph{Why not optimise Polsby-Popper in \mka{}?}
Polsby-Popper requires perimeter computation, which in turn requires the
polygon boundary data from TIGER/Line shapefiles.
Reock requires only the positions of tract centroids and the MEC geometry,
both available from the existing centroid companion file.
\mka{} therefore defaults to Reock; an optional \texttt{--mka-metric polsby}
flag is supported for contexts where Polsby-Popper is the legally operative
metric, but Reock is the preferred default because it requires no polygon
boundary data beyond what \cvdgeo{} already uses.
