\label{sec:mechanism}

\subsection{The Neck Geometry}

The mechanism connecting low compactness to high constituent driving time operates
primarily through what we call the district ``neck''---the narrow corridor that
connects two geographically distant population centers in a non-compact district.
A canonical gerrymandered district has the following geometric structure:

\begin{enumerate}
\item A \textit{core} region containing one population group (typically urban, Democratic-leaning
  in a Republican gerrymander).
\item A \textit{neck} corridor of minimal width connecting the core to an arm.
\item An \textit{arm} region containing the other population group (typically suburban
  or exurban).
\end{enumerate}

This structure minimizes the perimeter-to-area ratio \textit{of the core and arm} while
maximizing political homogeneity within each subregion---but the neck dramatically
increases the overall perimeter, reducing Polsby-Popper. Simultaneously, the neck
creates a geographic stretch: the population-weighted centroid of the district falls
somewhere along the neck (because population is distributed between core and arm),
but no one lives on the neck itself, and the representative's district office is
located in either the core or the arm.

Formally, we define the \textit{neck width ratio} as:

\begin{equation}
\text{NWR} = \frac{w_{\text{min}}}{d_{\text{max}}}
\end{equation}

where $w_{\text{min}}$ is the minimum width of the district (the minimum diameter of
any cross-section perpendicular to the district's long axis) and $d_{\text{max}}$ is
the maximum internal distance (the diameter of the district's convex hull). A compact
district with a roughly convex shape has NWR $\approx 0.5$--$0.8$; a district with
a narrow neck has NWR $< 0.15$.

We compute NWR for all 432 districts using the following procedure:
(1) generate the district's convex hull; (2) find the diameter of the convex hull
(the pair of vertices with maximum distance); (3) project the district polygon onto
the axis perpendicular to the diameter direction; (4) identify the minimum width along
this projection. Computationally, this requires generating 180 rotated projections
per district (at 1-degree intervals) and taking the minimum.

The NWR distribution across the 432 districts has median 0.44 (SD = 0.21), with 41
districts (9.5\%) having NWR $< 0.15$ (indicating clear neck geometry). These
41 districts have a mean driving time of 68.3 minutes, compared to 38.1 minutes
for districts with NWR $> 0.30$.

\subsection{Office Location Choices Under Neck Geometry}

An important secondary effect of neck geometry is that representatives from non-compact
districts must make a choice of office location that systematically disadvantages one
sub-community. A representative whose district contains a urban core in the north and
a rural arm in the south, connected by a narrow corridor, will locate their primary
office in whichever sub-community contains more voters. The other sub-community faces
dramatically elevated driving times.

To measure this effect, we compute the \textit{office location bias}: for each district,
the ratio of the 90th-percentile tract driving time to the 10th-percentile tract driving
time. A perfectly central office location would produce a ratio near 1; an office
located at one end of a non-compact district produces a ratio substantially above 1.

\begin{table}[ht]
\centering
\caption{Office Location Bias by Compactness Quartile}
\label{tab:bias}
\begin{tabular}{lrrrr}
\toprule
Compactness Quartile & Mean PP & Mean Driving Time & P90/P10 Ratio & $n$ \\
\midrule
Q1 (least compact) & 0.09 & 56.2 & 3.8 & 108 \\
Q2 & 0.18 & 43.7 & 2.9 & 108 \\
Q3 & 0.28 & 37.4 & 2.4 & 108 \\
Q4 (most compact) & 0.41 & 31.8 & 1.9 & 108 \\
\bottomrule
\end{tabular}
\begin{tablenotes}
\small
\item P90/P10 Ratio: ratio of 90th-percentile tract driving time to 10th-percentile
  tract driving time within district. Higher values indicate more unequal access.
\end{tablenotes}
\end{table}

The P90/P10 ratio---a measure of within-district access inequality---rises monotonically
from 1.9 in the most compact quartile to 3.8 in the least compact quartile. Constituents
in non-compact districts not only face longer average travel times but face more unequal
access: in the least compact quartile, a constituent at the 90th percentile of travel
time faces a journey 3.8 times longer than a constituent at the 10th percentile within
the same district. This within-district inequality is a distinct harm from the aggregate
driving time effect.

\subsection{Comparison with \textsc{bisect} Plans}

The \textsc{bisect} algorithm's minimum-edge-cut objective naturally avoids neck geometry.
A neck corridor represents a small number of edges connecting large sub-communities;
cutting those edges is precisely what the METIS partitioner seeks to minimize. As a result,
\textsc{bisect} plans have substantially fewer districts with NWR $< 0.15$: only 8 of
180 \textsc{bisect} districts in the gerrymandered states have NWR $< 0.15$, compared to
28 of 180 enacted districts ($p < 0.001$, Fisher exact test).

The absence of neck geometry in \textsc{bisect} plans is not a design choice but an
emergent consequence of the objective function: minimum edge cuts on the geographic
adjacency graph produce districts where each pair of adjacent tracts shares a substantive
border, eliminating the single-edge connections that create geographic necks. This
provides a mechanism explanation for the driving time reduction in \textsc{bisect}
plans that does not require any additional assumptions about representative behavior.
