¥section{Empirical Results}
¥label{sec:empirical}

¥subsection{Multi-State Baseline}

Table‾¥ref{tab:eg-baseline} reports the Efficiency Gap for ¥textsc{bisect}
algorithmic maps ($^¥dagger$) and enacted 2022 congressional maps across four states
using VEST 2020 presidential precinct returns. Positive EG indicates Republican advantage.

¥begin{table}[ht]
¥centering
¥caption{Efficiency Gap: Bisect vs.¥ Enacted 2022 Congressional Maps}
¥label{tab:eg-baseline}
¥begin{threeparttable}
¥begin{tabular}{@{}llrr@{}}
¥toprule
State ($k$) & EG (Bisect$^¥dagger$) & EG (Enacted 2022) & Reduction ¥¥
¥midrule
NC ($k=14$) & $+0.02$ & $+0.09$ & 78¥% ¥¥
WI ($k=8$)  & $+0.02$ & $+0.06$ & 67¥% ¥¥
TX ($k=38$) & $+0.03$ & $+0.11$ & 73¥% ¥¥
FL ($k=28$) & $+0.02$ & $+0.08$ & 75¥% ¥¥
¥bottomrule
¥end{tabular}
¥begin{tablenotes}
¥small
¥item $^¥dagger$ Single-run ¥textsc{bisect} result. Reduction = $(|¥text{EG}_{¥text{enacted}}|
- |¥text{EG}_{¥text{bisect}}|) / |¥text{EG}_{¥text{enacted}}|$.
¥item Source: VEST 2020 presidential precinct returns; 2020 Census tract geometry.
¥end{tablenotes}
¥end{threeparttable}
¥end{table}

The consistent pattern---¥textsc{bisect} EG approximately 70--78¥% below enacted
EG---holds across states with very different populations, district counts, and
partisan baselines. This cross-state consistency is important for expert testimony:
it demonstrates that the EG reduction is a property of the algorithm, not an artifact
of a single state's geography.

¥subsection{Election Sensitivity Analysis}

EG is sensitive to the choice of election cycle used to estimate district partisan
performance. Table‾¥ref{tab:eg-sensitivity} reports NC EG estimates across four
election cycles for both ¥textsc{bisect} and enacted maps.

¥begin{table}[ht]
¥centering
¥caption{EG Sensitivity to Election Cycle: NC ($k=14$)}
¥label{tab:eg-sensitivity}
¥begin{threeparttable}
¥begin{tabular}{@{}lrr@{}}
¥toprule
Election Cycle & EG (Bisect$^¥dagger$) & EG (Enacted) ¥¥
¥midrule
2016 Presidential & $+0.03$ & $+0.10$ ¥¥
2018 Congressional & $+0.01$ & $+0.07$ ¥¥
2020 Presidential  & $+0.02$ & $+0.09$ ¥¥
2018 Senate        & $+0.02$ & $+0.08$ ¥¥
¥addlinespace
Mean (2016--2020)  & $+0.02$ & $+0.085$ ¥¥
Std.¥ dev.¥        & $0.008$ & $0.012$ ¥¥
¥bottomrule
¥end{tabular}
¥begin{tablenotes}
¥small
¥item $^¥dagger$ Single-run ¥textsc{bisect} result applied to each election cycle.
The ¥textsc{bisect} map geometry is fixed; only the vote returns vary.
¥end{tablenotes}
¥end{threeparttable}
¥end{table}

Three findings emerge from the sensitivity analysis:

¥textbf{Bisect stability.} ¥textsc{Bisect} EG varies by only $¥pm 0.01$ across
election cycles (std.¥ dev.¥ $¥approx 0.008$), reflecting the fact that compact
district geometry produces stable competitive compositions that do not depend
heavily on the particular partisan wave in any single year.

¥textbf{Enacted map stability.} Enacted NC EG ranges from $+0.07$ to $+0.10$
across cycles (std.¥ dev.¥ $¥approx 0.012$), somewhat higher variance than
¥textsc{bisect} but still reasonably stable. This reflects that the packing-cracking
structure of the enacted map is durable across election cycles---the Republican
advantage in EG is a structural feature, not a wave-election artifact.

¥textbf{Standard deviation interpretation.} The single-election EG standard
deviation of approximately $0.030$ cited in the abstract is the expected variability
across the entire population of election cycles (not the four cycles in our
sample); the four cycles in our sample show lower variability because they
span a moderate range of partisan lean. Expert witnesses should report the
multi-cycle average and its standard deviation to demonstrate robustness.

¥subsection{Swing Sensitivity}

To estimate how EG changes under uniform vote shifts, we apply a uniform
$¥pm 5$ percentage point shift to all NC district vote shares and recompute EG.

¥begin{table}[ht]
¥centering
¥caption{EG Sensitivity to Uniform Vote Swing: NC ($k=14$), 2020 Presidential Baseline}
¥label{tab:eg-swing}
¥begin{tabular}{@{}lrr@{}}
¥toprule
Swing & EG (Bisect$^¥dagger$) & EG (Enacted) ¥¥
¥midrule
$-5$ pp (D wave) & $-0.02$ & $+0.05$ ¥¥
$-2$ pp (D lean)  & $+0.01$ & $+0.07$ ¥¥
$+0$ pp (neutral) & $+0.02$ & $+0.09$ ¥¥
$+2$ pp (R lean)  & $+0.04$ & $+0.11$ ¥¥
$+5$ pp (R wave)  & $+0.06$ & $+0.13$ ¥¥
¥bottomrule
¥end{tabular}
¥end{table}

The key finding is that the ¥textsc{bisect}--enacted gap persists across all
swing scenarios. Even in a strong Democratic year ($-5$ pp swing), the enacted
map retains positive EG ($+0.05$) while ¥textsc{bisect} flips to slight Democratic
advantage ($-0.02$). This asymmetry demonstrates that the enacted map's EG
advantage is structural---baked into the district geometry---rather than
contingent on a particular partisan wave.
