\section{Legal Landscape}
\label{sec:legal}

\subsection{Academic Origin and the 8\% Threshold}

The Efficiency Gap was introduced by \citet{stephanopoulosMcGhee2015} in a
University of Chicago Law Review article arguing that EG provides a judicially
manageable standard for identifying partisan gerrymanders. The authors proposed
an 8\% threshold as the level at which EG indicates a presumptively unconstitutional
partisan advantage. This threshold was derived from historical data on congressional
and state legislative elections showing that EGs above 8\% in the first election
following redistricting tend to persist and compound across the decade, creating
durable partisan advantage.

The 8\% threshold is an academic proposal, not a legal standard. No federal or
state court has adopted it as a constitutional boundary. Expert witnesses citing
EG must disclose that (1) the threshold was proposed by the metric's inventors,
creating a potential conflict of interest; (2) it was derived from regression
analysis of historical data that may not apply to contemporary redistricting
environments; and (3) SCOTUS in \textit{Gill v.\ Whitford} and \textit{Rucho}
declined to adopt any EG threshold.

\subsection{\textit{Gill v.\ Whitford} (2018)}

\textit{Gill v.\ Whitford}\footnote{\textit{Gill v.\ Whitford}, 585 U.S.\ 48 (2018).}
was the first Supreme Court case to consider EG as evidence. Plaintiffs argued
that Wisconsin's 2012 state assembly maps were a partisan gerrymander, relying
on EG evidence showing Wisconsin's statewide EG exceeded 10\% in multiple election
cycles. The district court in \textit{Whitford v.\ Gill}\footnote{\textit{Whitford
v.\ Gill}, 218 F.Supp.3d 837 (W.D.\ Wis.\ 2016).} credited EG evidence and ruled
for the plaintiffs.

SCOTUS reversed on standing grounds. The Court held that plaintiff voters lacked
standing because EG is a statewide metric that does not speak to the particularized
injury of any individual voter in any specific district. The Court expressly declined
to reach the merits of whether EG or any other metric provides a judicially manageable
standard for partisan gerrymandering, reserving that question. The reservation was
rendered moot by \textit{Rucho} the following year.

The practical significance of \textit{Gill} for expert witnesses: SCOTUS scrutinized
EG more carefully than any other metric has ever been examined in federal court,
and the decision's focus on standing---not on the metric's validity---left EG's
methodological foundations intact. The metric's sensitivity to election-year choice,
its threshold ambiguity, and its proposed-but-not-adopted 8\% boundary were all
documented in the federal record and are fair targets for cross-examination.

\subsection{\textit{Rucho v.\ Common Cause} (2019)}

\textit{Rucho}\footnote{\textit{Rucho v.\ Common Cause}, 588 U.S.\ 684 (2019).}
directly addressed the North Carolina congressional map and held that federal courts
cannot adjudicate partisan gerrymandering claims. The majority's analysis of
available metrics---including EG---concluded that no metric or combination of
metrics provides a judicially manageable standard that would allow courts to
distinguish between permissible partisan line-drawing and unconstitutional excess
without imposing the court's own political preferences.

This holding does not invalidate EG as a measurement tool; it holds that EG
(and other metrics) cannot serve as the constitutional standard themselves.
Expert witnesses in state court proceedings can continue to use EG as evidence;
the constitutional question of what EG value crosses the state constitutional
line is a question of state law that \textit{Rucho} explicitly left to state
courts and state legislatures.

\subsection{State Court Status Through 2026}

Post-\textit{Rucho} state court proceedings have cited EG without adopting any
threshold:

\begin{itemize}
\item \textbf{North Carolina} (\textit{Harper v.\ Hall}, 2022): NC Supreme Court
  cited EG as one of several metrics demonstrating systematic partisan advantage.
  Enacted NC congressional EG exceeded 9\%; the court invalidated the map without
  specifying an EG threshold.
\item \textbf{Wisconsin}: Wisconsin state courts have faced redistricting challenges
  in which EG was cited; none has adopted a specific EG threshold.
\item \textbf{Pennsylvania}: PA expert testimony in legislative redistricting
  proceedings has used EG alongside MM; PA courts have focused on legislative
  intent evidence rather than specific metric thresholds.
\end{itemize}

\subsection{Expert Witness Protocol}

Based on the legal history, the recommended expert witness protocol for EG evidence is:

\begin{enumerate}
\item Report the multi-cycle average (at minimum: 2016, 2018, 2020 presidential;
  2018 Senate) with the standard deviation, to demonstrate stability.
\item Report EG for \textsc{bisect} algorithmic maps as the neutral reference
  distribution, to establish what EG the underlying geography would produce
  without deliberate manipulation.
\item Explicitly disclose that the 8\% threshold is an academic proposal not
  adopted by any court, and that the relevant legal question is whether the
  state constitutional standard is violated by the specific EG gap between
  the enacted map and the neutral reference.
\item Present EG alongside at least one other metric (MM or the seats-votes curve)
  to resist metric-shopping criticism.
\end{enumerate}
