\section{Conclusion}
\label{sec:conclusion}

The Efficiency Gap remains the most litigated partisan fairness metric in American
redistricting, and for good reason: it is mathematically transparent, intuitively
grounded in the wasted-vote concept, and sensitive to the packing-cracking operations
that define deliberate partisan manipulation. Its legal career has been extraordinary---
from a 2015 law review article to the central exhibit in a Supreme Court case within
three years---and its post-\textit{Rucho} status as a state court tool (rather than
a federal constitutional standard) preserves its evidentiary role while freeing it
from the constraints of a failed federal constitutional theory.

The core findings of this paper are three. First, \textsc{bisect} algorithmic
maps produce EG values approximately 70--78\% below enacted maps across NC, WI,
TX, and FL---a threefold reduction attributable to the same compactness constraint
that distinguishes the algorithm's geographic structure from deliberate packing and
cracking. Second, single-election EG estimates are sensitive to partisan wave
conditions, with standard deviation $\approx 0.030$ across cycles for NC; multi-election
averaging (four cycles minimum) is required for stable expert testimony. Third,
the 8\% threshold proposed by \citet{stephanopoulosMcGhee2015} has no legal force;
expert witnesses must frame EG as evidence (not as a standard) and explain the
gap between enacted map EG and the neutral \textsc{bisect} reference distribution
as the portion attributable to deliberate manipulation.

The \textsc{bisect} reference distribution provides state courts with what federal
litigation never had: a clear, non-partisan baseline that resolves the ``compared
to what?'' question at the center of every partisan gerrymandering claim. When
a neutral algorithm produces NC EG $\approx 0.02$ and the enacted map produces
NC EG $\approx 0.09$, the answer to ``compared to what?'' is not an abstract
normative standard but an empirical measurement of what the same state's geography
produces under a fully transparent, non-partisan process.

\bigskip

\noindent\textit{The author thanks the redistricting research team for data
access and algorithmic development. All errors are the author's own.}
