\section{Divergence Between Total Population and Citizen VAP}
\label{sec:divergence}

\subsection{The CVAP Special Tabulation}

The Census Bureau publishes Citizen Voting Age Population (CVAP) estimates at the census-tract level through its CVAP special tabulation, derived from the American Community Survey 5-year estimates. The tabulation provides estimates of citizen VAP by race and ethnicity, enabling analysis of noncitizen concentration patterns. Because the CVAP is ACS-derived rather than decennial-census-derived, it carries sampling error at the tract level; this methodological limitation is discussed in Section~\ref{sec:bisect}.

The CVAP is available for the redistricting-relevant 2019 ACS 5-year estimates (corresponding to the 2020 redistricting cycle) at the census-tract level. BISECT's \texttt{--population-source cvap} mode reads this tabulation and uses CVAP rather than total population as the balance constraint in recursive bisection.

\subsection{State-Level Divergence Patterns}

The gap between total population and citizen VAP is largest in states with large immigrant populations. Table~\ref{tab:state_divergence} shows the mean noncitizen fraction (total VAP minus citizen VAP, divided by total VAP) for the four highest-divergence states.

\begin{table}[htbp]
\centering
\begin{tabular}{lcccc}
\toprule
\textbf{State} & \textbf{Total VAP} & \textbf{Citizen VAP} & \textbf{Noncitizen Pop} & \textbf{Mean Noncit. Frac.} \\
\midrule
California & 29,500,000 & 24,800,000 & 4,700,000 & 15.9\% \\
Texas & 20,400,000 & 17,300,000 & 3,100,000 & 15.2\% \\
New York & 15,300,000 & 13,100,000 & 2,200,000 & 14.4\% \\
Florida & 16,200,000 & 14,000,000 & 2,200,000 & 13.6\% \\
\midrule
National & 258,300,000 & 232,400,000 & 25,900,000 & 10.0\% \\
\bottomrule
\end{tabular}
\caption{Total VAP, Citizen VAP, noncitizen adult population, and mean noncitizen fraction for the four highest-divergence states and the national total, based on 2019 ACS 5-year CVAP tabulation. State figures are approximate.}
\label{tab:state_divergence}
\end{table}

The national noncitizen fraction of approximately 10\% understates the redistricting relevance because noncitizens are concentrated in specific congressional districts. At the district level, noncitizen fractions exceeding 20\% are common in Los Angeles, South Texas, New York City, and Miami.

\subsection{District-Level Divergence: The 80-District Threshold}

For congressional redistricting, the relevant question is not the statewide noncitizen fraction but the distribution of noncitizens across congressional districts. A 5 percentage-point (pp) gap between total-population and citizen-VAP district shares---our threshold for redistricting relevance---is exceeded in approximately 80 of the 435 congressional districts.

These 80 districts are highly concentrated: approximately 25 are in California (primarily Los Angeles metro and the Central Valley), 15 in Texas (primarily South Texas and Houston), 12 in New York (primarily New York City), and 8 in Florida (primarily Miami-Dade). The remaining 20 are distributed across New Jersey, Illinois, Arizona, Nevada, and Washington.

This 80-district figure should be read as a tract-assignment point estimate from the CVAP rerun, not as an exact census of affected districts. The final paper should report the corresponding sensitivity band from the point-estimate/90\% confidence-interval comparison in Section~\ref{sec:bisect}, because ACS error can move a small number of boundary-adjacent districts in or out of the set.

The consequence of exceeding the 5pp threshold is that switching from total population to citizen VAP as the redistricting base forces these districts to draw substantially different boundaries. A district whose total population is at the state target ($\sim$760,000) but whose citizen VAP is only 600,000 (a noncitizen fraction of 21\%) would need to expand geographically under citizen-VAP balancing to reach the citizen-VAP target---absorbing additional territory with more citizens and fewer noncitizens, which typically means moving into more suburban and less immigrant-dense areas.

\subsection{Intrastate Geographic Concentration}

Within California, the total-population--citizen-VAP divergence is not uniform across counties. The highest divergence occurs in:
\begin{itemize}
    \item \textbf{Los Angeles County}: Several congressional district tracts have noncitizen fractions of 25--35\%, reflecting large unauthorized immigrant and legal permanent resident populations in East Los Angeles, Koreatown, and the San Gabriel Valley.
    \item \textbf{San Bernardino and Riverside Counties} (Inland Empire): Growing immigrant communities with noncitizen fractions of 15--20\% in many tracts.
    \item \textbf{Central Valley} (Fresno, Tulare, Kings Counties): Agricultural workforce with noncitizen fractions exceeding 20\% in many rural tracts.
\end{itemize}

In Texas, the highest divergence is in:
\begin{itemize}
    \item \textbf{South Texas} (Webb, Hidalgo, Cameron, Starr Counties): Border counties with noncitizen fractions of 20--30\% in urban tracts, reflecting large unauthorized and legal immigrant Mexican-American communities.
    \item \textbf{Harris County} (Houston): Diverse immigrant community with noncitizen fractions of 15--20\% in several congressional district tracts.
\end{itemize}

These geographic patterns are not random: the high-noncitizen-fraction areas are disproportionately Hispanic and Asian-American communities, reflecting the demographic composition of post-1965 immigration to the United States.
