\section{Conclusion}
\label{sec:conclusion}

The noncitizen-population redistricting debate is legally unresolved but empirically clear: switching from total population to citizen VAP as the redistricting base produces a consistently Republican-advantaging shift, concentrated in the 80 congressional districts where the total-population--citizen-VAP gap exceeds five percentage points. The four high-immigration states---California, Texas, New York, and Florida---account for the majority of this effect.

The legal framework is layered. For congressional redistricting, the textual argument from Article I's Apportionment Clause provides a strong basis for holding that total population is the constitutionally mandated base, insulating congressional maps from citizen-VAP challenges. For state legislative redistricting, \emph{Evenwel v. Abbott} holds that total population is \emph{permissible} but does not resolve whether citizen VAP is also permissible; the open question will be litigated in state-level proceedings over the coming decade.

BISECT's \texttt{--population-source cvap} implementation enables systematic empirical analysis of the redistricting consequences of any citizen-VAP proposal. The principal methodological finding is that CVAP-balanced congressional redistricting plans in high-immigration states would produce total-population disparities across districts that are likely to fail \emph{Wesberry v. Sanders} equal-population scrutiny---a self-defeating feature of mandatory CVAP congressional redistricting that has not received sufficient attention in the legal literature.

The deepest issue is representational theory. \emph{Evenwel}'s majority recognized that noncitizens are legitimate constituents whose representatives serve their interests. Citizen-VAP redistricting would systemically dilute that representation without eliminating it---noncitizens would still have a representative, but their representative would serve a larger and less immigrant-dense constituency. This is a form of representation dilution that the Equal Protection Clause may ultimately prohibit, though no court has yet said so.

\paragraph{Acknowledgments.}
The author thanks the Census Bureau for the CVAP special tabulation and the Brennan Center for Justice for its compilation of citizen-VAP redistricting litigation.
