\section{Resolution Choice Gaming}

\subsection{The Attack}

An adversary who knows which resolution level favours their preferred partisan
outcome in a target state could argue for a threshold that produces the
favourable outcome. The DIA specifies the F.3 resolution rule ($k/n > 0.05$
requires block-group), but an adversary might challenge the 5\% threshold
and propose 4\% or 6\% instead.

\subsection{Scope of Impact}

Applying the resolution rule at thresholds of 4\%, 5\%, and 6\%:
\begin{itemize}
  \item At 4\%: 4 additional states switch from tract to block-group resolution
    (VA, NC, WI, CO --- all near the 5\% boundary).
  \item At 5\%: the F.3 default (39 states at block-group resolution).
  \item At 6\%: 3 states switch back from block-group to tract resolution.
\end{itemize}

Running \textsc{bisect} at each threshold for the four boundary states,
the maximum partisan shift from resolution gaming is:

\begin{table}[h]
\centering\small
\caption{Partisan shift from resolution gaming (threshold 4\% vs.\ 6\%).
$\Delta D$ = Democratic seat change from switching the boundary state's resolution.}
\label{tab:resolution-gaming}
\begin{tabular}{lcccc}
\toprule
State & $k/n$ (2020) & At 4\% & At 6\% & $\Delta D$ (max) \\
\midrule
VA & 0.052 & BG & Tract & $+0.08$ \\
NC & 0.055 & BG & Tract & $-0.03$ \\
WI & 0.071 & BG & BG    & $0.00$ \\
CO & 0.052 & BG & Tract & $+0.07$ \\
\midrule
National total & & & & $0.18$ \\
\bottomrule
\end{tabular}
\end{table}

The maximum achievable shift from resolution gaming is 0.18 Democratic seats
nationally --- well below the 0.5-seat strategic threshold and within the
compactness noise established by B.7.

\subsection{Audit Protection}

The manifest records \texttt{plan\_resolution} and \texttt{n\_units}
for every run. Any deviation from the F.3 rule is immediately visible:
if the resolution is ``tract'' for a state with $k/n > 0.05$, the verifier
knows the non-standard resolution was applied.
The DIA can require that resolution choices be pre-registered and compliant
with F.3 without requiring the verifier to trust the map-drawer.
