\label{sec:legal}

\subsection{Communities of Interest Doctrine and Economic Communities}

The redistricting criterion of preserving ``communities of interest'' has
deep roots in American redistricting practice.
The National Conference of State Legislatures surveyed state redistricting
criteria in 2021 and found that 35 states include some form of communities
of interest as a redistricting criterion, either in state constitutions,
statutes, or commission guidelines \citep{ncsl2021criteria}.

Economic community membership---workers and residents sharing an economic
zone, an industrial employer, a commercial district, or a labor market
area---is among the most widely recognized subcategories.
The legal foundation is explicit in multiple states.

\paragraph{California Proposition 11 (2008).}
The Voters First Act established California's Citizens Redistricting
Commission and listed communities of interest as the second redistricting
criterion (after equal population and VRA compliance) \citep{caprop11}.
The Commission's 2011 and 2021 proceedings both accepted economic community
characterizations from public testimony, including testimony about
agricultural communities in the Central Valley and tech industry communities
in Silicon Valley.
The LODES WAC-based economic character vector operationalizes these
characterizations: a Central Valley agricultural tract ($\mathrm{IF} > 0.4$)
and an adjacent residential suburb ($\mathrm{IF} < 0.05$) have low cosine
similarity, making the district boundary between them more natural.

\paragraph{Arizona Proposition 106 (2000).}
Arizona's Independent Redistricting Commission criteria list ``communities
of interest'' including ``economic and social interests'' \citep{azircprop106}.
The Commission's proceedings have accepted testimony about economic communities
since its first redistricting cycle in 2001.
Arizona's ranching and farming communities, technology corridors in
Scottsdale and Tempe, and copper mining towns in southeastern Arizona are
all recognizable in the LODES WAC data.

\paragraph{Colorado Constitution, Article V, Section 44.}
Colorado's independent redistricting commission criteria explicitly list
``communities of interest'' and provide examples that include ``economic
interests'' and ``trade area'' community membership \citep{coart5s44}.
The trade area concept---communities that share a retail trade zone or
labor market area---maps directly onto the JPR and CI signals in the
economic character vector.

\paragraph{Washington RCW 44.05.090.}
Washington's redistricting law requires preservation of ``communities
of related and mutual interests'' and has been interpreted to include
economic communities in Commission practice \citep{warcw44}.
Washington's port communities (Seattle, Tacoma, Bellingham), agricultural
communities (Eastern Washington irrigated farmland), and technology
corridors (Redmond, Bellevue) are all distinguishable through the
LODES WAC economic character signals.

\subsection{Partisan Neutrality: The Legal Requirement}

Following \textit{Rucho v.\ Common Cause} \citeyearpar{rucho2019}, federal
courts cannot adjudicate partisan gerrymandering claims.
State courts in North Carolina, Wisconsin, Pennsylvania, and elsewhere
have increasingly applied state constitutional standards.
The most common standard is that maps must not have been drawn predominantly
to achieve a partisan advantage.

Economic character weights satisfy this standard structurally because no
electoral data enters the computation.
The inputs are:
\begin{itemize}
  \item NAICS-sector job counts (LODES WAC, Census Bureau administrative data)
  \item Total population (2020 decennial census)
  \item Tract-to-tract adjacency (TIGER/Line shapefiles)
\end{itemize}
None of these inputs contains voter registration data, election results,
precinct boundaries, or any other partisan signal.

Any partisan effect of economic character weights---such as the NC-14 gap
improvement demonstrated in Section~\ref{sec:results}---is a consequence of
the geographic correlation between economic activity and partisan sorting.
This correlation is a feature of the population landscape, not a product of
the algorithm's choices.
Under \textit{Shaw v.\ Reno} \citeyearpar{shaw1993} and \textit{Miller v.\
Johnson} \citeyearpar{miller1995}, a similar analysis applies to racial effects:
the relevant question is whether race (or partisanship) was the \emph{predominant
factor}, not whether effects exist.
Economic character weights with no electoral inputs cannot be characterized
as partisan redistricting tools any more than geographic boundary-length
weights can.

\subsection{Expert Witness Deployment Template}

Redistricting practitioners deploying M.1 economic character weights in a
proceeding should structure expert testimony around four elements:

\paragraph{1. Data provenance declaration.}
``I used the U.S.\ Census Bureau's LODES Workplace Area Characteristics data,
version~8, for census year 2020 (\url{https://lehd.ces.census.gov/data/}).
This is a federal administrative dataset derived from state Unemployment
Insurance wage records.
It records the number of jobs, by NAICS sector, in each census block.
No voter registration data, election results, or racial composition data
is present in this dataset or used in the computation.''

\paragraph{2. Criterion alignment.}
``The [state]'s redistricting criteria require preservation of communities
of interest, including economic communities.
I computed an economic character vector for each census tract representing
its commercial intensity, industrial fraction, and employment density.
District boundaries are drawn preferentially at the boundaries between
economically dissimilar tracts---at the transition from residential
neighborhoods to commercial corridors, or from commercial zones to
industrial parks.
This respects the [state]'s statutory criterion.''

\paragraph{3. Specific community identification.}
``In [state], the following economic communities are preserved by the
proposed plan: [specific examples from LODES data analysis, e.g.,
`the Research Triangle Park employment cluster in Durham, Orange, and
Wake counties is contained within a single district rather than split
across three districts as in the challenger's map'].''

\paragraph{4. Algorithmic transparency.}
``The blend formula is $w(u,v) = w_{\mathrm{geo}}(u,v) \times (\alpha +
(1-\alpha) \times \mathrm{sim}(\mathbf{e}_u, \mathbf{e}_v))$ where
$\alpha = 0.5$ and $\mathrm{sim}$ is the cosine similarity of the economic
character vectors.
The geographic boundary-length weight is preserved at full weight for
economically similar adjacent tracts and at half weight for the most
economically dissimilar adjacent tracts.
The formula is symmetric, non-negative, and bounded.
It has been implemented in the \textsc{bisect} platform and verified by
unit tests.''

\subsection{When Not to Use M.1}

Economic character weights should \emph{not} be used in the following contexts:

\begin{itemize}
  \item \textbf{States without a communities-of-interest criterion.}
    If the state's redistricting statute does not include communities of
    interest as a criterion, the weight modification lacks a statutory hook
    and may be challenged as an unauthorized departure from the statutory criteria.
  \item \textbf{States with diffuse economic geography.}
    As demonstrated by the Wisconsin null result, states without concentrated
    employment clusters will see no benefit.
    Presenting a zero-effect weight modification to a commission or court
    as a communities-of-interest tool is not credible.
  \item \textbf{When economic geography cuts against VRA compliance.}
    If economic character weights would reduce minority opportunity districts
    below VRA-required thresholds, geographic or VRA-aligned weights must
    take priority.
    The three-layer compositor in \textsc{bisect} supports combining
    \texttt{economic-character} with \texttt{vra-aligned} for contexts
    where both criteria apply.
\end{itemize}

The practitioner's checklist before deploying M.1:
\begin{enumerate}
  \item Confirm that the state has a communities-of-interest criterion.
  \item Confirm that the state has identifiable economic concentration clusters
    visible in the LODES data.
  \item Confirm that the resulting plan satisfies VRA and equal population constraints.
  \item Prepare LODES data documentation for the expert declaration.
\end{enumerate}
