\section{Audit Artifacts}

A selected frontier entry should be exportable as RPLAN/RCTX with certificate
and manifest sidecars. The audit package verifies the selected plan and lineage
under a declared profile; it does not certify that the frontier is complete or
that the selected trade-off is legally preferred.

The manifest must include the fixed seed, population and generation identifiers,
operator parameters, parent ids for generated descendants, selected entry id,
objective vector, constraints profile, RPLAN path, RCTX path, certificate path,
and verification transcript. If a candidate reached the frontier through a
fallback or repair path, that path remains part of the selected-entry lineage.

Selection criteria are separate audit inputs. U.19 can package a selected
frontier entry for verification, but preference among valid non-dominated
entries belongs to the U.14 selection record and the U.15 posture boundary.

The public golden package can be checked either through the RPLAN certificate
verifier or through the BISECT manifest verifier.

The selected package schema is
\texttt{bisect-pareto-selected-frontier-package-v1}. It writes
\texttt{selected-frontier.rplan}, \texttt{selected-frontier.rctx},
\texttt{audit-certificate.json}, and \texttt{manifest.json}. The manifest
records seed, selected index, frontier size, objective values, validity status,
certificate digest, certificate content hash, legal profile id, and context
hash. That package is the point where an exploratory evolutionary row becomes a
normal audit object.
