\subsection{Constitutional Status of MMDs}

The U.S. Constitution does not require single-member districts for the House
of Representatives.  Article I, Section 2 requires only that representatives
be ``chosen every second Year by the People of the several States,'' and
Section 4 gives Congress power to regulate the ``Time, Places and Manner''
of elections.  The Constitution is silent on district magnitude.

The statutory requirement for single-member districts derives from 2 U.S.C.
\S~2c, enacted in its current form in 1967.  This statute reads: ``In each
State entitled in the Ninety-first Congress or in any subsequent Congress to
more than one Representative under an apportionment made pursuant to the
provisions of [the apportionment statutes], there shall be established by
law a number of districts equal to the number of Representatives to which
such State is so entitled, and Representatives shall be elected only from
districts so established.''  This language clearly requires single-member
districts as a matter of current statutory law.  However, Congress could
repeal or amend Section 2c.  Bills to do so have been introduced repeatedly
\citep{fair_rep_act_2019}.

\subsection{Illinois Precedent (1870--1980)}

The most relevant domestic precedent for MMD redistricting is Illinois's
state legislative system from 1870 to 1980.  Under the Illinois Constitution
of 1870, the state Senate was elected from single-member districts while the
House was elected from three-member districts using Cumulative Voting
\citep{cole2001cumulative}.  Each voter cast three votes, allocable in whole
or fractional amounts among candidates.

Cumulative Voting in Illinois produced significant minority party
representation: Republicans routinely won one of three seats in safe
Democratic Chicago districts, and Democrats won one seat in safe Republican
downstate districts.  The system was abolished by referendum in 1980
(the ``Cutback Amendment'') under pressure from a taxpayer revolt seeking to
reduce the size of the legislature rather than because of dissatisfaction
with its representative outcomes \citep{cole2001cumulative}.

The Illinois precedent is directly relevant because: (a) it demonstrates
the constitutional validity of MMDs at the state legislative level; (b) it
provides 110 years of operational experience with three-member districts in
an American political context; and (c) the elimination of the system was
motivated by cost-cutting, not representational failure.

\subsection{Voting Rights Act Implications}

Section 2 of the Voting Rights Act (VRA) prohibits voting practices that
``result in a denial or abridgement of the right ... to vote on account of
race or color.''  Under \emph{Thornburg v.\ Gingles} \citep{thornburg1986},
a Section 2 violation requires showing that (1) the minority group is
sufficiently large and geographically compact to constitute a majority in a
single-member district, (2) the minority is politically cohesive, and (3) the
white majority votes sufficiently as a bloc to defeat minority-preferred
candidates.

The interaction between MMDs and the VRA is complex.  On one hand, MMDs may
satisfy VRA requirements without creating majority-minority single-member
districts: proportional allocation in a 4-member district where minorities
are 35\% of the population will typically elect 1--2 minority-preferred
candidates.  On the other hand, \emph{Thornburg}'s ``sufficiently compact''
prong has been interpreted to require geographic compactness in a
\emph{single-member} district.  Some courts have held that this prong is
inapplicable where no single-member structure exists \citep{vra_mmd_review}.
The law on this point is unsettled.

\subsection{Fair Representation Act}

The Fair Representation Act (FRA), introduced in multiple Congresses by
Rep.\ Don Beyer and others, would require states with more than one
congressional seat to use multi-member districts with ranked-choice voting
(STV) \citep{fair_rep_act_2019}.  The Act would supersede 2 U.S.C.\ \S~2c
and explicitly authorise districts of 3--5 members.  It would also amend
the VRA to clarify that minority representation in MMDs satisfies Section 2
obligations.

The computational framework developed in this paper is directly applicable
to implementing the FRA: bisect generates the district boundaries, and our
STV implementation provides the seat-allocation algorithm.  The case studies
presented here could serve as illustrative plans for congressional testimony
or commission evaluation.
