\section{Introduction}
\label{sec:introduction}

Partisan gerrymandering litigation ultimately presents courts with a normative
choice that no metric can resolve: what kind of representation should a democratic
system produce? Two fundamentally different answers have dominated Anglo-American
electoral theory:

\textbf{Proportionality}: a party's seat share should reflect its vote share.
If Democrats win 50\% of votes, they should win approximately 50\% of seats.
Proportionality requires that the translation from votes to seats be as accurate
as possible, minimizing the seat bonus that any party receives above its proportional
entitlement. It is the dominant norm in European electoral systems that use proportional
representation formulas (closed-list, open-list, mixed-member proportional).

\textbf{Majoritarianism}: the party winning a plurality of votes should receive
a seat-share bonus that magnifies its mandate, producing stable single-party
government and clear accountability. Majoritarianism holds that the purpose of
elections is to produce governing majorities, not proportional delegations.
It is the norm that Anglo-American single-member plurality systems have historically
embodied---the British and Canadian first-past-the-post traditions explicitly
accept high disproportionality as the price of stable majorities.

American congressional redistricting inhabits the tension between these frameworks.
The single-member district structure produces majoritarian seat bonuses by default:
any party winning more than half the votes in a state will tend to win more than
half the seats, because tight geographic distributions produce efficient seat
conversions above the 50\% threshold. But the redistricting decisions that determine
\textit{which} party receives the majority bonus, and \textit{how large} it is,
are made by legislatures with obvious self-interest in maximizing their own party's bonus.

This paper examines that tension through the Gallagher Index---the standard measure
of electoral disproportionality in comparative politics---and analyzes where
\textsc{bisect} algorithmic maps and enacted congressional maps fall on the
proportionality-majoritarianism spectrum. The practical contribution is a framework
for expert witnesses that acknowledges the normative contestability of the choice
(courts, not experts, decide which standard applies) while providing empirical
evidence relevant under either standard.

Section~\ref{sec:definition} defines the Gallagher Index and the majoritarian bonus.
Section~\ref{sec:behavior} analyzes how \textsc{bisect} maps behave under both
standards. Section~\ref{sec:empirical} presents empirical results across four states.
Section~\ref{sec:legal} analyzes the state constitutional provisions that may impose
proportionality constraints and the post-\textit{Rucho} federal foreclosure.
Section~\ref{sec:conclusion} provides the expert witness framework.
