\section{Behavior Under Algorithmic Redistricting}
\label{sec:behavior}

\subsection{Bisect and Proportionality}

\textsc{Bisect} algorithmic maps achieve Gallagher Index scores in the range
$[0.03, 0.08]$ across NC, WI, TX, and FL---significantly below enacted maps
($[0.10, 0.18]$) but not zero. The structural floor ($\text{LSq} > 0$) reflects
two sources documented in Section~\ref{sec:definition}: geographic sorting
and winner-take-all arithmetic. Neither can be eliminated by any redistricting
plan; they are inherent in the single-member plurality structure.

The practical implication is that proportionality is not a realistic standard
for evaluating individual redistricting plans. The question is not whether a plan
achieves $\text{LSq} = 0$ (impossible in a single-member system with geographic
sorting) but whether a plan achieves a $\text{LSq}$ consistent with what neutral
redistricting would produce for the same state. The \textsc{bisect} LSq provides
exactly this neutral benchmark.

\subsection{Bisect and Majoritarianism}

Under the majoritarian standard, \textsc{bisect} maps produce a seat bonus of
approximately $1.3$--$1.5$ for the plurality-winning party---within the historical
Anglo-American average. This reflects the structural majority-bonus of single-member
systems applied to NC's approximately $54\%$ Republican statewide presidential vote
in 2020: a 54\% plurality produces approximately $8/14 = 57\%$ of seats under
\textsc{bisect}, a bonus of $57/54 \approx 1.06$ in seat-share ratio terms.

Enacted NC maps produce a bonus of approximately $2.1$: Republicans winning $54\%$
of votes win $8/14 \approx 57\%$ of seats---but the packing and cracking of Democratic
voters ensures that this $57\%$ is structurally guaranteed rather than competitively
earned. The enacted bonus ($2.1$) far exceeds the structural majority-bonus ($1.3$--$1.5$)
of neutral maps, reflecting the redistricting contribution to the seat bonus.

\subsection{Why Compactness Approaches Both Standards}

The paradox of compactness is that it simultaneously improves proportionality
(reducing $\text{LSq}$) and reduces the unearned majoritarian bonus (the portion
of the bonus above the structural floor). This is not a coincidence: both
improvements flow from the same underlying property of compact maps---the
distribution of district vote shares is more uniform and less extreme.

A uniform distribution of district vote shares near the statewide mean produces
approximately proportional seat shares (the parties win districts roughly in
proportion to their total vote share) while also producing a competitive majority
bonus (the plurality winner earns its seats through competitive district outcomes
rather than structural safe-seat insulation). Gerrymandered maps with deliberately
extreme vote-share distributions are less proportional \textit{and} produce larger
unearned bonuses for the majority party---both failures reflect the same underlying
vote-concentration strategy.
