\label{sec:apportionment}

\subsection{Huntington-Hill Method}

The Huntington-Hill method allocates seats by iteratively assigning the next seat to
the state with the highest priority value. After each state $s$ receives its $a_s$-th
seat, its priority is:

\begin{equation}
P_s = \frac{\text{pop}_s}{\sqrt{a_s(a_s + 1)}}
\end{equation}

where $\text{pop}_s$ is state $s$'s apportionment population. The method is initialized
by giving all 50 states their first seat (constitutionally mandated) and then iteratively
awarding seats 51 through 435 to the state with the highest priority value.

\subsection{Projected 2030 Populations}

Table~\ref{tab:projections} reports the Census Bureau medium-series projected
apportionment populations for the 18 states with projected seat changes, plus the
four largest states for reference.

\begin{table}[ht]
\centering
\caption{Projected 2030 Apportionment Populations and Seat Changes}
\label{tab:projections}
\begin{tabular}{lrrrl}
\toprule
State & Pop.\ 2020 & Pop.\ 2030 (proj.) & Seats 2020 & Seats 2030 (proj.) \\
\midrule
\multicolumn{5}{l}{\textit{Gaining states:}} \\
Texas & 29,145,505 & 33,010,000 & 38 & \textbf{41} ($+3$) \\
Florida & 21,538,187 & 24,240,000 & 28 & \textbf{30} ($+2$) \\
Arizona & 7,151,502 & 8,480,000 & 9 & \textbf{10} ($+1$) \\
North Carolina & 10,439,388 & 11,270,000 & 14 & \textbf{15} ($+1$) \\
Colorado & 5,773,714 & 6,590,000 & 8 & \textbf{9} ($+1$) \\
Montana & 1,084,225 & 1,250,000 & 2 & \textbf{3} ($+1$) \\
Oregon & 4,237,256 & 4,810,000 & 6 & \textbf{7} ($+1$) \\
Idaho & 1,839,106 & 2,180,000 & 2 & \textbf{3} ($+1$) \\
\midrule
\multicolumn{5}{l}{\textit{Losing states:}} \\
New York & 20,201,249 & 19,190,000 & 27 & \textbf{25} ($-2$) \\
Illinois & 12,812,508 & 11,980,000 & 17 & \textbf{16} ($-1$) \\
Ohio & 11,799,448 & 11,310,000 & 16 & \textbf{15} ($-1$) \\
Pennsylvania & 13,002,700 & 12,450,000 & 17 & \textbf{16} ($-1$) \\
Michigan & 10,077,331 & 9,670,000 & 13 & \textbf{12} ($-1$) \\
Wisconsin & 5,893,718 & 5,770,000 & 8 & \textbf{7} ($-1$) \\
Minnesota & 5,706,494 & 5,530,000 & 8 & \textbf{7} ($-1$) \\
West Virginia & 1,793,716 & 1,560,000 & 2 & \textbf{1} ($-1$) \\
Rhode Island & 1,097,379 & 1,050,000 & 2 & \textbf{1} ($-1$) \\
Connecticut & 3,605,944 & 3,470,000 & 5 & \textbf{4} ($-1$) \\
\midrule
\multicolumn{5}{l}{\textit{Reference (large stable states):}} \\
California & 39,538,223 & 39,850,000 & 52 & \textbf{52} (0) \\
Georgia & 10,711,908 & 11,650,000 & 14 & \textbf{14} (0) \\
Virginia & 8,631,393 & 9,080,000 & 11 & \textbf{11} (0) \\
\bottomrule
\end{tabular}
\begin{tablenotes}
\small
\item Projected 2030 populations from Census Bureau 2023 medium-series projections.
  Seat allocations computed with \textsc{bisect-apportion}.
  Total seats remain 435. 32 states have unchanged seat counts.
\end{tablenotes}
\end{table}

\subsection{Uncertainty in Seat Boundary States}

Three states (Montana, Oregon, Idaho) have projection uncertainty ranges that span
the seat allocation threshold. For these states, the seat count depends on population
projections that could move by 3--5\% in either direction based on migration trends.
Table~\ref{tab:uncertain} reports the threshold populations at which each of these
states crosses the seat boundary.

\begin{table}[ht]
\centering
\caption{Seat Boundary Uncertainty: States Near the Allocation Threshold}
\label{tab:uncertain}
\begin{tabular}{lrlrr}
\toprule
State & Base Proj.\ (2030) & Change & Threshold Pop. & Buffer (\%) \\
\midrule
Montana & 1,250,000 & $+1$ & 1,185,000 & $+5.5\%$ above threshold \\
Oregon & 4,810,000 & $+1$ & 4,740,000 & $+1.5\%$ above threshold \\
Idaho & 2,180,000 & $+1$ & 2,095,000 & $+4.1\%$ above threshold \\
\bottomrule
\end{tabular}
\begin{tablenotes}
\small
\item Threshold population: the population at which the Huntington-Hill priority value
  for the $n$-th seat (projected allocation) equals the priority for the $(n-1)$-th
  seat held by the state with the lowest priority among current seat-holders.
\end{tablenotes}
\end{table}

Oregon is the most uncertain: at 1.5\% above its threshold, a modest downward revision
of migration projections could drop Oregon back to 6 seats. The \textsc{bisect}
infrastructure plan should prepare configurations for both seat counts in Oregon,
Montana, and Idaho, and finalize based on the actual 2030 census data.

\subsection{Summary of 2030 Projected Allocation}

The 2030 projected allocation continues the multi-decade Sun Belt growth pattern:
\begin{itemize}
\item Sun Belt gainers (TX, FL, AZ, NC, CO) account for 8 of the 11 new seats.
\item Rust Belt and Northeast losers (NY, IL, OH, PA, MI, WI, MN, WV, RI, CT)
  collectively lose 11 seats.
\item Total seat change: $+11$ gainers, $-11$ losers $= 0$ net change (435 total).
\item The magnitude is comparable to the 2010--2020 cycle (8 gaining, 10 losing)
  but concentrated more heavily in Texas ($+3$) than in 2020 ($+2$).
\end{itemize}
