\label{sec:stability}

\subsection{The County Disruption Metric}

Cross-census algorithmic stability measures how much redistricting plans change
between two redistricting cycles, holding constant the algorithmic parameters.
The T.9 paper introduced the \textit{county disruption metric} as the primary
stability measure: the fraction of county-seat assignments that change between
the two cycles.

\begin{definition}[County Disruption]
For state $s$ with county set $C_s$ and seat allocations $A^{(t)} = \{a_c^{(t)}\}$
for census years $t$ and $t+10$, the county disruption $D_s$ is:

\begin{equation}
D_s = \frac{|\{c \in C_s : a_c^{(t)} \neq a_c^{(t+10)}\}|}{|C_s|}
\end{equation}

where $a_c^{(t)}$ is the district assignment of county $c$'s plurality census tract
in cycle $t$.
\end{definition}

County disruption ranges from 0 (perfect stability: every county is in the same
district across both cycles) to 1 (complete disruption: every county changes district
assignment). For states where $k$ changes between cycles, disruption is computed
using the larger-$k$ cycle's district assignments as the baseline and asking how many
county assignments are maintained despite the structural change.

\subsection{Projected 2020--2030 Disruption}

We project county disruption for the 2020--2030 cycle using two inputs:
(1) the 2020 \textsc{bisect} plans (from the 50-state sweep), which provide the
baseline county-district assignments; and (2) Census Bureau projected population
distribution across counties in 2030, which determines how the optimal redistricting
changes.

The projection method uses a partial-equilibrium approach: for each state, we apply
the 2030 projected county-level populations to the 2020 adjacency graph and re-run
the \textsc{bisect} algorithm with the 2030 $k$-value. Counties that change district
assignment between the 2020 and projected-2030 optimal plans are counted as disrupted.

\begin{table}[ht]
\centering
\caption{Projected County Disruption: 2020--2030 vs.\ Observed 2010--2020}
\label{tab:disruption}
\begin{tabular}{llrrr}
\toprule
Category & $n$ states & Disruption 2010--2020 & Projected 2020--2030 & Difference \\
\midrule
Gaining states & 8 & 78\% & 72\% & $-6$ pp \\
Losing states & 10 & 71\% & 51\% & $-20$ pp \\
Stable states & 32 & 61\% & 54\% & $-7$ pp \\
\midrule
All states & 50 & 66\% & 58\% & $-8$ pp \\
\bottomrule
\end{tabular}
\begin{tablenotes}
\small
\item Disruption is fraction of counties changing district assignment.
  2010--2020 values from T.9 (observed). 2020--2030 values are projections.
\end{tablenotes}
\end{table}

The projected 2020--2030 disruption (58\%) is lower than the observed 2010--2020
disruption (66\%) for three reasons:

\textbf{Predictable Sun Belt growth}: Unlike the 2010 cycle, where several Sun Belt
states had unexpected population surges (Texas grew 15.9\% vs.\ a projected 13\%),
the 2030 projections extrapolate a well-established trend in Sun Belt growth. The
county-level population distribution in 2030 Texas, Arizona, and Florida is more
predictable from current trends, so the optimal county groupings change less
dramatically.

\textbf{Losing state contraction}: States losing seats (NY, IL, OH, PA, MI) are
contracting at predictable rates. County disruption in losing states falls from 71\%
to 51\% because the optimal redistricting plan for a state losing one seat typically
involves merging two districts that are adjacent in the current plan, preserving most
county assignments.

\textbf{Stable state inertia}: States with unchanged $k$-values and moderate growth
rates show modest disruption reduction. The largest within-cycle shifts are in
fast-growing stable states (Georgia, Virginia, Nevada), where internal population
redistribution changes optimal county groupings even without seat changes.

\subsection{State-Specific Stability Highlights}

\textbf{Texas}: Despite gaining 3 seats ($38 \to 41$), Texas shows projected disruption
of 68\%---lower than the 81\% observed disruption in the 2010--2020 cycle when Texas
also gained 4 seats. The lower disruption reflects more predictable growth patterns
and the fact that the new seats accommodate growth in already-identified growth corridors
(I-35 corridor from San Antonio to Austin, DFW exurbs).

\textbf{New York}: Losing 2 seats ($27 \to 25$) produces projected disruption of 44\%,
substantially below the 61\% observed disruption in the 2010--2020 cycle (when New York
lost 2 seats then as well, $29 \to 27$). The lower projected disruption reflects that
New York's population concentration in the New York City metro area makes the optimal
county groupings relatively invariant: the outer boroughs and surrounding counties must
be grouped together regardless of $k$.

\textbf{Montana}: Gaining 1 seat ($2 \to 3$) triggers substantial projected disruption
(88\%) because the current 2-seat configuration divides the state along a
east-west axis, while the 3-seat configuration must create a third district, likely
in the Billings corridor. The entire county assignment structure changes.
