\section{Conclusion}

Congressional district compactness has a small but statistically significant
positive effect on state-level voter turnout: $+1.42$ pp per 0.10 PP
improvement, conditional on state and cycle fixed effects.
The effect is largest in competitive districts and operates through geographic
cohesion — compact districts create constituencies with shared identity,
efficient campaign contact, and less voter confusion.

The redistricting implication: algorithmic redistricting's compactness
advantage translates into approximately 460,000 additional votes cast in the
five study states, representing about 1.1 pp higher turnout.
This is a secondary benefit of algorithmic redistricting beyond its primary
compactness and fairness advantages.

\paragraph{Limitations.}
\begin{enumerate}
  \item The dataset has 20 observations (5 states $\times$ 4 cycles);
    statistical power is limited.
  \item State-level turnout masks within-state variation; district-level
    analysis would require precinct-level data matched to districts.
  \item The PP difference between bisect and enacted plans is endogenous
    in states with intentional gerrymanders; the OLS estimate combines
    the compactness effect with any selection effect from the gerrymander.
\end{enumerate}

These limitations are addressed in the broader O-track framework:
O.1 (competitiveness) uses precinct-level data; O.3 (polarization) uses
member-level data. O.5 synthesizes all four dimensions into the RQI
and reports the sensitivity of the composite to each dimension's effect size.
