\documentclass[12pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{amsmath, amssymb, amsthm}
\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}
\newtheorem{proposition}{Proposition}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{remark}{Remark}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage{booktabs}
\usepackage{xcolor}
\usepackage{multirow}
\usepackage{array}
\usepackage{microtype}
\usepackage{setspace}
\usepackage{titlesec}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{lmodern}
\usepackage[T1]{fontenc}

%% SIAM-adjacent style: numbered sections, upright headings
\titleformat{\section}{\normalfont\large\bfseries}{\thesection.}{0.5em}{}
\titleformat{\subsection}{\normalfont\normalsize\bfseries}{\thesubsection.}{0.5em}{}
\titleformat{\subsubsection}{\normalfont\normalsize\itshape}{\thesubsubsection.}{0.5em}{}

%% Spacing
\setstretch{1.4}

%% Algorithm macros
\newcommand{\bfsg}{\textsc{BfsGrowth}}
\newcommand{\cvd}{\textsc{CVD}}
\newcommand{\metis}{\textsc{Metis}}
\newcommand{\ec}{\ensuremath{\mathrm{EC}}}
\newcommand{\pp}{\ensuremath{\mathrm{PP}}}
\newcommand{\sa}{\textsc{SimulatedAnnealing}}
\newcommand{\be}{\textsc{BisectionEnsemble}}
\newcommand{\ar}{\textsc{ApportionRegions}}
\newcommand{\geosec}{\textsc{GeoSection}}
\newcommand{\areasec}{\textsc{AreaSection}}
\newcommand{\recom}{\textsc{ReCom}}
\newcommand{\cs}{\textsc{ConvergenceSweep}}

%% Math shorthands
\newcommand{\ECnorm}{\mathrm{EC}_{\mathrm{norm}}}
\newcommand{\EC}{\mathrm{EC}}
\newcommand{\PPmean}{\overline{\mathrm{PP}}}

%% -----------------------------------------------------------------------

\title{%
  \textbf{BFS Region-Growing: A Minimal Contiguous}\\
  \textbf{Bisection Baseline for BISECT}\\[0.8em]
  \normalsize\textit{T.12 --- Algorithm Design / Geometric Structure}%
}

\author{%
  Gio Della-Libera\\
  \textit{Apportionment Research Group}\\
  \small\texttt{giodl@microsoft.com}
}

\date{May 2026}

\begin{document}

\maketitle

\begin{abstract}
\noindent
\bfsg{} is BISECT's smallest implemented structure-layer alternative to
METIS-style bisection. At each bisection-tree node it selects one
population-weighted seed, chooses a second seed at maximum BFS distance from
the first, grows two frontier regions with a priority queue keyed by current
district population, assigns any disconnected leftovers to the smaller side,
and applies the same 200-step local boundary rebalance family used elsewhere
in the platform. The only required inputs are adjacency, population, balance
tolerance, base seed, and node path.

This paper corrects the older presentation of BFS Growth. The implementation is
a recursive two-way bisection primitive, not a general $k$-seed simultaneous
constructor. It is a baseline and warm-start candidate, not a certified
compactness floor. Its strongest property is legibility: every accepted tract
is reached from an already-owned neighbor unless the disconnected-graph safety
net fires. The remaining publication gap is empirical: NC/WI/TX quality and
warm-start claims need archived same-metric run packages before they carry
standalone weight. Implementation: \texttt{--structure bfs-growth}.
\end{abstract}

\input{sections/01-introduction}
\input{sections/02-background}
\input{sections/03-algorithm}
\input{sections/04-comparison}
\input{sections/05-warmstart}
\input{sections/06-conclusion}

\bibliographystyle{plainnat}
\bibliography{references}

\end{document}
