\label{sec:background}

\subsection{Defining a Safe Seat}

We define a district as \textit{safe} if its expected Democratic presidential vote
share falls outside the competitive range $[0.35, 0.65]$---that is, if the expected
margin of victory for the dominant party exceeds 15 percentage points. Formally:

\begin{definition}[Safe Seat]
District $d$ is a \emph{safe-Democratic seat} if $v_d^D > 0.65$, a
\emph{safe-Republican seat} if $v_d^D < 0.35$, and \emph{competitive} if
$|v_d^D - 0.50| \leq 0.15$, where $v_d^D$ is the two-party Democratic presidential
vote share in district $d$ in 2020:
\[
  v_d^D = \frac{\sum_{t \in d} V_t^D}{\sum_{t \in d} (V_t^D + V_t^R)},
\]
and $V_t^D$, $V_t^R$ are the 2020 Democratic and Republican presidential votes in
tract $t$.
\end{definition}

The 15 percentage point threshold is conventional in the political science literature
on electoral competition~\citep{fiorina1980incumbency} and corresponds approximately
to the margin that insulates an incumbent from being defeated by a statewide wave of
moderate magnitude. We additionally classify districts as \textit{lean-Democratic}
($v_d^D \in (0.50, 0.65]$) and \textit{lean-Republican} ($v_d^D \in [0.35, 0.50)$)
to provide a more granular picture of the competitive landscape.

\subsection{Measurement Approach}

Presidential vote share is a standard proxy for district-level partisan lean because
it is relatively stable across election cycles, is not directly susceptible to
incumbency advantage (since the presidential candidates are the same in all districts),
and is available at fine geographic granularity through the VEST database.

We aggregate VEST 2020 presidential precinct returns to 2020 census tracts via an
area-weighted spatial join~\citep{vest2020}. Precinct boundaries do not align with
census tract boundaries; we apportion precinct votes to census tracts in proportion to
the share of each precinct's area that falls within each tract. This introduces
measurement error in tracts that are split across many precincts, particularly in
urban areas with small precincts. In NC and WI, the area-weighting error is estimated
to be less than 0.5 percentage points for 95\% of tracts; in TX, where some rural
precincts are very large, the error can reach 1--2 percentage points in sparsely
populated tracts.

\subsection{Geographic Sources of Safe Seats}

Not all safe seats are created equal. Two distinct mechanisms produce safe districts
under any redistricting plan, including algorithmic ones:

\paragraph{Geographic sorting.}
When voters of one party are heavily concentrated in a small geographic area---most
dramatically in urban cores---any contiguous district drawn around that area will be
heavily partisan regardless of how the district is shaped. Downtown Chicago, urban
Houston, and the Research Triangle in NC will all produce safe-Democratic districts
under any compact map that groups their tracts together. These geographically
unavoidable safe seats are present in both algorithmic and enacted maps.

\paragraph{Deliberate packing and cracking.}
Deliberate packing concentrates one party's voters into fewer, very safe seats to
``waste'' their votes in a small number of blowout victories. Deliberate cracking
disperses those voters across multiple districts to deny them majorities anywhere.
Both tactics increase the safe-seat count for the party doing the packing and cracking,
and both require information about voter partisan behaviour that goes beyond what compact
geography alone would imply.

The comparison between algorithmic and enacted safe-seat counts provides a measure of
how much of the enacted map's safe-seat count is attributable to geographic sorting
(which algorithmic maps also produce) versus deliberate packing and cracking (which only
enacted maps can produce, because they use partisan data).
