\section{Benchmark Targets}
\label{sec:comparison}

\subsection{Evidence Status}

The older T.12 draft contained specific NC/WI/TX compactness, edge-cut, runtime,
and partisan tables. Those values are useful hypotheses, but this paper should
not present them as final evidence until the corresponding run directories,
source hashes, seed manifests, and analysis outputs are archived.

The current verified evidence is narrower:
\begin{itemize}
\item small-grid BFS Growth returns a complete two-way split;
\item repeated runs with the same seed are deterministic;
\item assigned sides are contiguous on connected toy graphs before validation;
\item the \texttt{bfs-growth} structure mode parses through the CLI surface; and
\item \texttt{run\_all\_splits\_bfs} produces complete assignments for toy k=2 cases.
\end{itemize}

\subsection{Publication Matrix}

\begin{table}[t]
\centering
\caption{BFS Growth benchmark matrix. Values remain pending until archived
same-metric packages are available.}
\label{tab:main-comparison}
\renewcommand{\arraystretch}{1.2}
\begin{tabular}{llrrrr}
\toprule
State & Algorithm & PP & EC & Pop dev. & Package \\
\midrule
NC ($k=14$) & \bfsg{}  & TBD & TBD & TBD & pending \\
NC ($k=14$) & \metis{} & TBD & TBD & TBD & pending \\
NC ($k=14$) & \cvd{}   & TBD & TBD & TBD & pending \\
\midrule
WI ($k=8$) & \bfsg{}  & TBD & TBD & TBD & pending \\
WI ($k=8$) & \metis{} & TBD & TBD & TBD & pending \\
WI ($k=8$) & \cvd{}   & TBD & TBD & TBD & pending \\
\midrule
TX ($k=38$) & \bfsg{}  & TBD & TBD & TBD & pending \\
TX ($k=38$) & \metis{} & TBD & TBD & TBD & pending \\
TX ($k=38$) & \cvd{}   & TBD & TBD & TBD & pending \\
\bottomrule
\end{tabular}
\end{table}

\subsection{What to Measure}

The comparison should separate three questions.

\paragraph{Construction quality.}
How much compactness and edge-cut quality does the simplest adjacency-growth
rule provide before heavier optimization is added?

\paragraph{Validation cost.}
How often does boundary rebalance leave a plan that final validation must
repair or reject, especially on disconnected or island-heavy graphs?

\paragraph{Odd-seat behavior.}
For odd-seat bisection nodes, how different is a half-population BFS split from
the bisection tree's intended $k_L:k_R$ child ratio?

\subsection{Reader-Facing Example}

The benchmark package should include at least one worked tract-block example
showing every heap pop:

\begin{center}
\begin{tabular}{c@{\quad}c@{\quad}c@{\quad}c}
\textbf{Step} & \textbf{Pop 0} & \textbf{Pop 1} & \textbf{Assigned tract} \\
\hline
0 & 10 & 12 & seeds only \\
1 & 18 & 12 & side 0 grows \\
2 & 18 & 20 & side 1 grows \\
3 & 25 & 20 & side 0 grows \\
4 & 25 & 27 & side 1 grows \\
\end{tabular}
\end{center}

This is the visual standard for T.12: show the heap decision, not just the final
map.
