\section{Conclusion}
\label{sec:conclusion}

\bfsg{} is valuable because it is small. It gives the BISECT family a
minimal, seed-auditable, adjacency-only construction baseline: choose two
spread seeds, grow the smaller side first, fill disconnected leftovers, and
rebalance locally.

\paragraph{Correct role.}
\bfsg{} should be described as a baseline and warm-start candidate, not as a
legal certificate and not as a proven compactness floor. Strong compactness,
runtime, and partisan claims belong in archived benchmark packages.

\paragraph{Implementation caveats.}
The current implementation is recursive two-way bisection, not simultaneous
$k$-way growth. The disconnected-graph safety net and local rebalance step mean
the final plan still needs ordinary validation. Odd-seat nodes also need future
ratio-aware target handling if \bfsg{} is to make strong claims for bisection
trees with uneven child seat counts.

\paragraph{Next work.}
The next useful milestones are:
\begin{enumerate}
\item archive NC/WI/TX same-metric packages for \bfsg{}, \metis{}, and \cvd{};
\item add a worked heap-pop trace to the public algorithm atlas;
\item decide whether ratio-aware split targets should be shared across CVD,
  BFS Growth, and other recursive splitters; and
\item run paired warm-start chain experiments before making any MCMC
  convergence claims.
\end{enumerate}

In the algorithm family, \bfsg{} is the first thing to teach after a plain
graph: it shows how a district can grow from a seed by adjacency alone. That
clarity is its main contribution.
