\documentclass[12pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{amsmath, amssymb, amsthm}
\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}
\newtheorem{proposition}{Proposition}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{remark}{Remark}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage{booktabs}
\usepackage{xcolor}
\usepackage{multirow}
\usepackage{array}
\usepackage{microtype}
\usepackage{setspace}
\usepackage{titlesec}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{lmodern}
\usepackage[T1]{fontenc}

%% SIAM-adjacent style: numbered sections, upright headings
\titleformat{\section}{\normalfont\large\bfseries}{\thesection.}{0.5em}{}
\titleformat{\subsection}{\normalfont\normalsize\bfseries}{\thesubsection.}{0.5em}{}
\titleformat{\subsubsection}{\normalfont\normalsize\itshape}{\thesubsubsection.}{0.5em}{}

%% Spacing
\setstretch{1.4}

%% Algorithm macros
\newcommand{\mka}{\textsc{MKA}}
\newcommand{\reock}{\ensuremath{\mathrm{Reock}}}
\newcommand{\cvd}{\textsc{CVD}}
\newcommand{\cvdgeo}{\textsc{CVD-Geo}}
\newcommand{\bfsg}{\textsc{BfsGrowth}}
\newcommand{\metis}{\textsc{Metis}}
\newcommand{\ec}{\ensuremath{\mathrm{EC}}}
\newcommand{\pp}{\ensuremath{\mathrm{PP}}}
\newcommand{\sa}{\textsc{SimulatedAnnealing}}
\newcommand{\be}{\textsc{BisectionEnsemble}}
\newcommand{\ar}{\textsc{ApportionRegions}}
\newcommand{\areasec}{\textsc{AreaSection}}
\newcommand{\geosec}{\textsc{GeoSection}}
\newcommand{\recom}{\textsc{ReCom}}
\newcommand{\cs}{\textsc{ConvergenceSweep}}
\newcommand{\dia}{\textsc{Districting Integrity Act}}

%% Math shorthands
\newcommand{\ECnorm}{\mathrm{EC}_{\mathrm{norm}}}
\newcommand{\EC}{\mathrm{EC}}

%% Dagger notation for single-run empirical tables
\newcommand{\singrun}{$^\dagger$}

%% -----------------------------------------------------------------------

\title{%
  \textbf{Moving-Knife Bisection: A Directional}\\
  \textbf{Reock-Scored Splitter for BISECT}\\[0.8em]
  \normalsize\textit{T.13 --- Algorithm Design / Geometric Structure}%
}

\author{%
  Gio Della-Libera\\
  \textit{Apportionment Research Group}\\
  \small\texttt{giodl@microsoft.com}
}

\date{May 2026}

\begin{document}

\maketitle

\begin{abstract}
\noindent
\mka{} is BISECT's directional sweep splitter for Reock-scored bisection.
At each bisection-tree node it projects tract centroids into Albers EPSG:5070,
tests $n_{\mathrm{orient}}$ orientations in $[0,\pi)$, sorts tracts by
projection onto each sweep axis, cuts at the first prefix whose population
reaches one half of the node population, scores the two halves by the minimum
of their centroid-MEC Reock approximations, and keeps the best orientation.
The selected assignment is then passed through the same bounded local boundary
rebalance family used by the other geometric splitters.

This paper corrects the earlier presentation of Moving-Knife. The implemented
algorithm is a recursive two-way bisection primitive, not a proof of global
Reock optimality for a final $k$-district plan. Its current Reock score is a
centroid approximation: the denominator is the minimum enclosing circle of
projected tract centroids, and the numerator falls back to the number of
centroids in the current implementation. The \texttt{polsby-popper} metric path
currently falls back to the same Reock scorer because perimeter data are not
yet available in the split routine. The method is nevertheless useful: it gives
BISECT a legible geometric baseline whose candidate choices can be audited as
``show the sweep, show the population prefix, show the worst-half score.''

The publication gap is empirical and archival. NC/FL/WA quality tables,
partisan-neutrality claims, per-node optimal-angle manifests, polygon-boundary
Reock, and true Polsby-Popper scoring all require archived run packages before
they should carry certification weight. Implementation:
\texttt{--structure moving-knife}.
\end{abstract}

\tableofcontents
\newpage

\input{sections/01-introduction}
\input{sections/02-background}
\input{sections/03-algorithm}
\input{sections/04-comparison}
\input{sections/05-fair-division}
\input{sections/06-conclusion}

\bibliographystyle{plainnat}
\bibliography{references}

\end{document}
