\label{sec:implementation}

The \textsc{bisect-apportion} Adams implementation uses the priority:
\[
  P_i^{\text{Ad}}(s) = \frac{p_i}{s}
\]
for $s \geq 1$ (with the first seat guaranteed by the constitutional
floor initialisation).
The current implementation evaluates this priority as \texttt{f64} inside the
shared \texttt{apportionment\_divisor()} priority queue:

\begin{verbatim}
RoundingRule::Adams => p / n,
\end{verbatim}

\begin{verbatim}
let seats = apportionment_divisor(&populations, 435, RoundingRule::Adams);
\end{verbatim}

There is not yet a standalone \texttt{bisect apportion --method adams} CLI.

\subsection{European Usage of Adams}

Although Adams has not been used for U.S.\ congressional apportionment,
ceiling rounding (also called the ``method of smallest divisors'')
has been used in some European subnational apportionment contexts.
The Austrian Nationalrat uses a three-stage apportionment process
that includes a ceiling step at the regional level \citep{lijphart1994electoral}.
Italy used ceiling rounding in some historical proportional
representation allocations.
These international uses confirm that Adams is a practically viable
method, not merely a theoretical construct.

\subsection{Test Suite}

\begin{itemize}
  \item Equal-population and 3:1 ratio synthetic examples cover the Adams path.
  \item A small-state-bias fixture compares Adams against Huntington-Hill.
  \item Shared divisor-method tests assert total-seat preservation and the
    one-seat constitutional floor.
  \item The package fixture
    \texttt{docs/examples/j-apportionment-evidence-packages/divisor-method-smoke}
    hash-binds an Adams allocation and replays it through
    \texttt{bisect-apportion::divisor\_evidence}.
\end{itemize}
