\section{DIA Protections Against All Five Vectors}
\label{sec:dia}

\subsection{The Pre-Registration Protocol}

The \dia{} pre-registration protocol is the primary countermeasure against
all five gaming vectors.
Pre-registration works by requiring all design choices (parameters, algorithm
version, geographic definitions) to be publicly filed and SHA-256 hashed
before the Census Bureau releases P.L.\ 94-171 redistricting data.

The pre-registration timeline is:
\begin{enumerate}
\item \textbf{T-12 months}: \dia{} administrator files the parameter
  configuration and algorithm version hash with the Secretary of State.
\item \textbf{T-6 months}: Public comment period on the filed configuration.
  Any party can challenge parameters as inconsistent with the statutory
  specification.
\item \textbf{T-0}: Census redistricting data released; the algorithm
  runs immediately from the pre-registered configuration.
\item \textbf{T+30 days}: Final plan and manifest filed for public
  verification.
\end{enumerate}

Under this timeline, the adversary cannot observe the partisan implications
of different parameter choices before the parameters are locked.
V1 is therefore eliminated before the first byte of Census data is processed.

\subsection{Cryptographic Chain}

The \dia{} manifest contains:
\begin{enumerate}
\item SHA-256 hash of the algorithm binary ($\delta_5 = 1.00$)
\item SHA-256 hash of the parameter configuration ($\delta_1 = 1.00$)
\item SHA-256 hash of each downloaded Census P.L.\ 94-171 file ($\delta_2 \approx 1.00$)
\item SHA-256 hash of the adjacency graph ($\delta_3 = 0.95$)
\item SHA-256 hash of the final plan assignment ($\delta_4 = 0.80$, bounded
  by the ability to detect unreported runs)
\end{enumerate}

\texttt{bisect label-verify} replays the full computation chain and checks
each hash in sequence.
Any discrepancy at any level is reported as a verification failure.

\subsection{The Impossibility Theorem}

\begin{theorem}[Gaming Impossibility under DIA Protocol]
\label{thm:gaming-impossibility}
Let $\mathcal{A}$ be an adversary who knows the full \bisect{} algorithm
specification, has read access to all Census data, and has write access
to the deployment environment.
Under the \dia{} pre-registration protocol, $\mathcal{A}$ cannot produce a
plan with expected partisan shift $> 0.5$ Democratic seats from the
pre-registered baseline without generating a manifest hash mismatch
detectable by \texttt{bisect label-verify}.
\end{theorem}

\begin{proof}[Proof sketch]
To produce a shift $> 0.5$ seats without a parameter change, $\mathcal{A}$
must modify (a) the input data beyond the $\pm 1\%$ per-tract noise floor,
(b) the adjacency definition beyond the statutory rule, or (c) the algorithm
binary.
Each modification produces a detectable hash change.
The claim that the shift is $> 0.5$ under the undetectable $\pm 1\%$ noise
bound follows from the empirical V2 bound derived in Section~\ref{sec:v2}
and the saturation argument of Section~\ref{sec:combined}.
$\square$
\end{proof}

\subsection{Mandatory Disclosure for V4}

The \dia{} requires that all redistricting computations be logged and
that the manifest identify the submitted plan's run index.
If more than one run was performed, all run hashes must be disclosed.
Cherry-picking the most partisan-favorable run without disclosing
the distribution of outcomes is a protocol violation.

Enforcement: any party can subpoena computational logs and compare
against the disclosed run hashes.
The \texttt{bisect label-verify} tool can replay any run identified
by its seed hash.
