\section{Legal Framing}
\label{sec:legal}

\subsection{The Dominance Certificate}

When the \nsga{} run with \texttt{--enacted-plan} returns
\texttt{is\_enacted\_dominated: true}, the NDJSON metadata record contains:

\begin{verbatim}
{
  "type": "enacted",
  "ec": 3891, "d_seats": 0.8, "vra_deficit": 0.12,
  "is_dominated": true,
  "dominating_plans": 7
}
\end{verbatim}

This record supports the following legal assertion, which we call the
\emph{Pareto dominance certificate}:

\begin{quote}
\textbf{Certificate of Multi-Objective Dominance.}
\textit{A multi-objective optimisation algorithm (NSGA-II) with population size
100 and 200 generations, base seed 42, was applied to North Carolina
congressional redistricting ($k=14$, 2020 Census) across three objectives:
edge-cut compactness ($\EC$), partisan proportionality ($\Dseats$), and
VRA minority representation ($\VRA$).
The algorithm identified 7 redistricting plans that are simultaneously more
compact (lower $\EC$), at least as proportional (lower or equal $\Dseats$),
and more VRA-compliant (lower $\VRA$) than the enacted 2020 congressional
map.
The algorithm found that the legislature could have enacted a plan that was
provably better by all three stated criteria simultaneously.
The complete output file and algorithm parameters are retained in
\texttt{nc\_pareto\_2020.ndjson}; the file SHA-256 is recorded in the
metadata line and can be independently verified.}
\end{quote}

This certificate has no heuristic or ensemble equivalent.
A practitioner using ensemble analysis can show that the enacted plan is an
outlier on \emph{one} criterion; the dominance certificate shows it is
sub-optimal on \emph{all three criteria simultaneously}.

\subsection{Strength of the Pareto Dominance Argument}

\paragraph{Comparison with single-objective tests.}
A single-objective compactness test (e.g., from ILP redistricting,
B.24~\citep{dellaLibera2026ilp}) certifies that a more compact plan exists.
This is a valid legal argument for compactness, but it does not address whether
the more compact plan satisfies partisan proportionality or VRA requirements.
Opposing counsel can respond: ``The more compact plan reduces minority
representation.''

The dominance certificate anticipates this response: all three criteria are
optimised simultaneously.
There is no trade-off to exploit: the dominating plans are better on
compactness, proportionality, and VRA at once.
Opposing counsel cannot argue that compactness came at the cost of VRA
compliance, because the certificate demonstrates $\VRA$ also improves.

\paragraph{Comparison with ensemble-based arguments.}
Ensemble expert reports (Mattingly and Vaughn, DeFord et al.\ in redistricting
litigation) use statistical arguments: the enacted plan is at the $P$th
percentile of a distribution of valid plans.
These arguments are strong but statistical: they assert that the enacted plan
is \emph{unusual}, not that it is \emph{dominated}.

The dominance certificate is a combinatorial argument, not a statistical one.
It does not claim the enacted plan is unusual; it claims a specific set of
plans are strictly better on all criteria.
Whether this rises to the level of unconstitutional gerrymandering is a
legal question --- but the mathematical result is definitive within the
specified parameters.

\paragraph{Federal vs.\ state court context.}
In federal court, \textit{Rucho v.\ Common Cause}~\citep{rucho2019} held that
federal partisan gerrymandering claims are non-justiciable political questions.
The dominance certificate does not automatically convert a federal gerrymandering
claim into a justiciable one: federal courts declined to adopt any mathematical
standard.

In state court, the picture is different.
\textit{Common Cause v.\ Lewis}~\citeyear{common_cause_lewis} established that
the North Carolina Constitution prohibits partisan gerrymandering and that
quantitative evidence is relevant.
In this context, the dominance certificate provides strong evidence:
the enacted plan is worse than a discoverable alternative on the
legislature's own stated criteria of compactness, proportionality, and
VRA compliance.

\subsection{Validity vs.\ Completeness}

The dominance certificate makes a \emph{validity claim}, not a
\emph{completeness claim}.
This distinction is critical for accurate expert testimony.

\paragraph{Validity (what the certificate asserts).}
The 4-step verification protocol confirms:
\begin{enumerate}
\item The file SHA-256 matches (\texttt{file\_sha256} in metadata agrees
  with a fresh hash of all preceding NDJSON lines).
\item The plan assignments in the output are valid redistricting plans:
  all districts are contiguous and within population balance tolerance.
\item The objective values in the output are correct: \EC{}, $\Dseats$,
  and $\VRA$ are recomputed from the plan assignments and census data
  and agree with the recorded values.
\item The returned plans are mutually non-dominating: no plan in the
  frontier is dominated by any other plan in the frontier.
\end{enumerate}
These checks can be performed in minutes by any expert with access to the
census data.
They do not require the Rust implementation or access to the platform's
source code.

\paragraph{Completeness (what the certificate does not assert).}
The certificate does not assert that the returned plans are the \emph{true}
Pareto front of the redistricting feasible space.
A manipulated implementation could return any set of mutually non-dominating
plans that happen to dominate the enacted plan.
The returned plans are verified to be valid and non-dominating, but the
algorithm's convergence to the true Pareto front is not verified.

A \emph{completeness check} requires re-running the full algorithm with the
documented seeds on the same census data.
This is possible (the seed specification in Section~\ref{sec:algorithm}
provides everything needed for a verifier to reproduce the run) but
computationally expensive: 15--60 minutes for a medium state.
For litigation use, the completeness check can be performed by opposing
counsel's expert with the documented parameters; the seed specification
makes this straightforward.

\paragraph{Practical implication.}
An expert witness presenting the dominance certificate should qualify it as:
``NSGA-II with the specified parameters identified 7 plans that are
provably better than the enacted plan on all three stated criteria.
These 7 plans have been independently verified to be valid plans with
the stated objective values.
The completeness of the Pareto front (whether additional such plans exist)
was not verified, but opposing counsel may reproduce the run using the
documented seed parameters.''

\subsection{The SMC-Pareto Alternative as Legal Evidence}

The SMC-Pareto projection (\texttt{--method smc-project}) uses the
SMC particle cloud as the plan population.
Its legal value is somewhat different from the \nsga{} dominance certificate:

\begin{itemize}
\item \textbf{Distribution calibration}: SMC particles are drawn from a
  near-uniform distribution over valid plans, calibrated by importance
  weights.
  A dominating SMC plan is not just a searched-for extreme --- it is a
  plan that appears in the natural distribution of valid plans.
  This may make the SMC-Pareto dominance argument more intuitive to
  a non-technical audience: ``even randomly drawn plans tend to be
  better than the enacted map.''

\item \textbf{Coverage limitation}: SMC may miss the extreme compactness
  or VRA extremes of the Pareto front.
  The \nsga{} dominance certificate is therefore stronger: it actively
  searches for Pareto-optimal plans and is more likely to find a dominating
  plan if one exists.

\item \textbf{Combined use}: if both \nsga{} and SMC-Pareto find dominating
  plans, this is stronger evidence than either alone.
  The \nsga{} result shows dominating plans exist in the optimised frontier;
  the SMC result shows dominating plans appear in the natural distribution.
\end{itemize}

\subsection{Scope and Limitations}

\paragraph{Population balance tolerance.}
The certificate is conditional on $\varepsilon_{\mathrm{bal}} = 0.005$.
A certificate at $\pm 0.5\%$ does not certify dominance at $\pm 1.0\%$;
relaxing the balance constraint expands the feasible space and may change
which plans appear on the Pareto front.
The tolerance must be disclosed alongside the certificate.

\paragraph{Vote data vintage.}
The $\Dseats$ objective uses 2020 presidential precinct returns.
Results using 2016 or 2022 returns may differ; the choice of vote data is
a methodological decision that practitioners should disclose and justify.

\paragraph{Three-objective scope.}
The certificate covers the three stated objectives: \EC{}, $\Dseats$, and
$\VRA$.
Practitioner-added criteria --- preservation of political subdivisions,
communities of interest, or specific minority coalition structures ---
are not in the objective set.
A plan dominated on the three stated objectives may be non-dominated when
additional objectives are included.
Practitioners who argue from the dominance certificate should disclose
which objectives were included and why.

\paragraph{Legal caveat.}
The dominance certificate addresses a mathematical question (multi-objective
optimisation over the stated criteria) and does not constitute legal advice.
Redistricting compliance requires assessment of criteria beyond those
optimised here, including applicable state constitutional standards and
specific Voting Rights Act section analyses.
Practitioners should consult legal counsel before presenting the dominance
certificate in litigation or regulatory proceedings.
