\section{Conclusion}
\label{sec:conclusion}

We have introduced \pareto{}-optimal redistricting, the first multi-objective
algorithm in this platform that returns a frontier of plans rather than a
single weighted plan.
\nsga{} with $N_{\mathrm{pop}} = 100$ plans and $N_{\mathrm{gen}} = 200$
generations produces a Pareto front of 10--30 plans on medium-sized states
(NC $k=14$, WI $k=8$), spanning meaningful ranges on all three objectives
--- edge-cut compactness (\EC{}), partisan proportionality ($\Dseats$), and
VRA minority representation ($\VRA$).

\paragraph{What Pareto redistricting is the right tool for.}

\begin{enumerate}
\item \textbf{Enacted-plan dominance testing.}
  The primary legal application: run \nsga{} and check whether any
  Pareto-front plan dominates the enacted plan on all three objectives
  simultaneously.
  On NC ($k=14$, 2020 Census), 7 Pareto-front plans dominate the enacted map.
  On WI ($k=8$, 2020 Census), no plan dominates the enacted map --- the
  test produces meaningful negative results and does not always find
  dominance.

\item \textbf{Trade-off transparency.}
  When a practitioner must choose among competing redistricting criteria,
  the Pareto front makes the trade-off structure explicit without requiring
  weight specification.
  A special master can inspect the 21 NC Pareto-front plans and choose
  among them with full knowledge of what compactness costs what
  proportionality.

\item \textbf{Multi-objective legal argument.}
  The dominance certificate is logically stronger than any single-objective
  test: it shows the enacted plan is simultaneously sub-optimal on
  compactness, proportionality, and VRA, not merely on one dimension.
  Opposing counsel cannot trade one objective against another to undermine
  the certificate.
\end{enumerate}

\paragraph{What Pareto redistricting is not the right tool for.}
\nsga{} redistricting is not a replacement for single-plan production tools.
For generating the best single plan for a given jurisdiction, \metis{} with
ConvergenceSweep~\citep{dellaLibera2026comparison} remains the correct default:
it is faster, simpler, and produces a single plan ready for submission.
For statistical analysis of the plan distribution, SMC~\citep{dellaLibera2026smc}
is the correct tool.
\nsga{} is the correct tool when the question is ``what are the trade-offs?''
or ``is the enacted plan dominated?''

\paragraph{Phase~2 enhancements.}

\begin{enumerate}
\item \textbf{Continuous $\Dseats$ proxy.}
  Replace integer $D_w$ with vote margin (sum of $|\text{district vote share} -
  0.5| \times \text{district population}$ for all districts) as a continuous
  proxy.
  This resolves the discrete-resolution limitation identified in
  Section~\ref{sec:background} and produces a genuinely three-dimensional
  Pareto front.

\item \textbf{VRA as hard constraint.}
  A \texttt{--vra-constrained} flag that filters the plan population to
  VRA-compliant plans before running \nsga{}, so that the Pareto front
  spans only VRA-compliant plans.
  This is appropriate when VRA compliance is a non-negotiable requirement
  and the Pareto trade-off is entirely between \EC{} and $\Dseats$.

\item \textbf{NSGA-II vs.\ SMC-Pareto comparison.}
  Empirical measurement of whether \nsga{} finds Pareto regions that
  the SMC projection misses, on NC 2020 with $N_{\mathrm{particles}} = 5{,}000$
  vs.\ $N_{\mathrm{pop}} = 100$, $N_{\mathrm{gen}} = 200$.

\item \textbf{50-state sweep.}
  Run the enacted-plan dominance test for all 50 states under 2020 Census.
  Report which states' enacted maps are dominated by the \nsga{} Pareto front
  and which are not.

\item \textbf{Rayon parallelism.}
  Objective evaluation (\EC{}, $\Dseats$, $\VRA$ for all plans in the
  population) is embarrassingly parallel.
  Rayon parallelism over the population is straightforward and would
  reduce sequential runtime by a factor of $\sim$8 on a modern workstation.
\end{enumerate}

\paragraph{Legal caveat.}
The dominance certificate establishes a mathematical result: the returned
plans are valid, mutually non-dominating, and strictly better than the
enacted plan on the three stated objectives.
It does not constitute legal advice and does not certify compliance with
applicable state constitutional standards, Section 2 or Section 5 of the
Voting Rights Act, or other legal redistricting requirements.
Practitioners should consult legal counsel before presenting the certificate
in litigation or regulatory proceedings.

\paragraph{Summary.}
\pareto{}-optimal redistricting makes trade-offs explicit and the enacted-plan
dominance test is the most legally actionable output.
Phase~1 (this paper) provides the \nsga{} implementation, empirical results
on NC and WI, and the validity-vs.-completeness framework for expert testimony.
Phase~2 will extend the approach with a continuous $\Dseats$ proxy, VRA
as a hard constraint option, and a 50-state dominance sweep.

Together with \metis{} (production-scale single plan),
\be{}~\citep{dellaLibera2026comparison} (high-quality single-objective search),
SMC (calibrated ensemble), and ILP (exact single-objective certificate),
the platform now spans the full space of redistricting algorithm questions:
fast, high-quality, statistically calibrated, exactly certified,
and Pareto-transparent.
