Every major redistricting reform of the past two decades has placed ``communities of
interest'' on the list of criteria that map-drawers must satisfy or at least
consider~\citep{ncsl2021criteria}.
California's Proposition 11 (2008), which created the Citizens Redistricting Commission,
requires the Commission to ``favor keeping together communities of interest''---a
phrase that appears four times in the enabling statute.
Colorado's independent commission must give ``communities of interest, including but not
limited to ethnic, cultural, economic, trade area, geographic, and demographic factors''
explicit weight under Art.\ V, \S44 of the state constitution.
Arizona's Independent Redistricting Commission lists preservation of communities of
interest among its six mandatory criteria, alongside equal population, compactness,
and VRA compliance.
Washington's Redistricting Commission is directed by RCW 44.05.090 to give consideration
to ``communities of interest including racial and language minority groups.''

Yet courts that review maps drawn under these standards face an immediate problem:
none of the statutes define what a community of interest is, how to measure whether
one has been preserved, or how to adjudicate competing claims when communities of
interest conflict with compactness or with county integrity.
Judges must rely on the testimony of expert witnesses, who in turn rely on qualitative
descriptions, anecdotal neighborhood histories, and---when challenged---on informal
assertions that two neighborhoods share a common identity.
No mathematical definition exists in case law.
The result is that ``preserve communities of interest'' is legally recognized but
operationally unenforceable in quantitative terms.

This paper proposes to close that gap.
We define a community character vector $\mathbf{c}(t)$ for each census tract $t$,
whose dimensions encode measurable signals of tract identity: the occupational composition
of resident workers, the land-use mix, the housing tenure and vintage, the commuting
orientation, the topographic setting, the set of administrative zones the tract belongs
to, and the transit accessibility of its residents.
Two adjacent tracts $u$ and $v$ that share similar character vectors are, by this
definition, members of the same community; the cosine similarity of their character
vectors provides a continuous measure of community cohesion.
We then define an edge weight modifier
\begin{equation}
  w(u,v) = w_{\mathrm{base}} \times \mathrm{sim}(\mathbf{c}_u, \mathbf{c}_v),
  \label{eq:general-weight}
\end{equation}
which, when passed to the METIS graph partitioner underlying \textsc{bisect}, causes
the partitioner to be geometrically reluctant to cut through edges connecting
similar-character tracts.
A cut that would split a community of interest becomes a costly cut; a cut that
separates tracts of genuinely different character is cheap.
The framework is partisan-neutral: no electoral data, no racial composition, and no
incumbency information enters any component of the character vector or the weight
formula.

\subsection{Contribution}

This paper makes the following contributions:
\begin{enumerate}
  \item \textbf{Mathematical framework}: We provide the first precise mathematical
    definition of ``community of interest'' as cosine similarity of tract-level
    character vectors, formalized as an edge weight modifier for METIS graph
    partitioning (\S\ref{sec:framework}).
  \item \textbf{Legal grounding}: We trace the communities of interest doctrine from
    \textit{Reynolds v.\ Sims} (1964) through \textit{Shaw v.\ Reno} (1993)
    and into modern state-level commission criteria, demonstrating that the framework
    satisfies the neutrality requirements imposed by Shaw (\S\ref{sec:legal}).
  \item \textbf{Signal taxonomy}: We provide a structured catalogue of seven
    community character signals (M.1--M.7), each with its data source, similarity
    metric, availability, and legal precedent (\S\ref{sec:spectrum}).
  \item \textbf{Integration specification}: We specify how M-track weights compose
    with existing \textsc{bisect} edge weight modes via alpha-blending
    (\S\ref{sec:bisect-integration}).
\end{enumerate}

\subsection{Scope and Limitations}

This paper is a framework paper.
It does not report redistricting runs or district-level outcomes; those are the province
of M.1--M.7, which validate each signal's empirical effect on the plans produced by
\textsc{bisect}.
This paper establishes the mathematical and legal scaffolding that all M-track papers
share.
The framework is also not a claim about partisan effect: because the character signals
are partisan-neutral by construction, the framework does not predict or guarantee any
particular partisan distribution of districts.
Whether communities of interest so defined produce more or less competitive districts,
more or less proportional seat allocations, or better or worse compactness scores
are empirical questions addressed in M.1--M.7 and M.8.

\subsection{Paper Organization}

Section~\ref{sec:framework} defines the mathematical framework.
Section~\ref{sec:legal} grounds it in redistricting doctrine.
Section~\ref{sec:spectrum} previews the seven signal types.
Section~\ref{sec:bisect-integration} explains how M-track weights integrate with
\textsc{bisect}'s three-layer compositor.
Section~\ref{sec:conclusion} concludes with a discussion of the framework's
role in expert witness testimony.
