\subsection{Formula}

Let $D$ be a district containing tracts $\{t_1, \ldots, t_m\}$.
Let $p_i$ be the population of tract $t_i$ and $c_i \in \mathbb{R}^2$ be
the centroid of $t_i$.
The total population is $P_D = \sum_i p_i$.
The population-weighted centroid of $D$ is:
\begin{equation}
  \bar{c}_D = \frac{\sum_i p_i c_i}{\sum_i p_i}.
  \label{eq:centroid}
\end{equation}
The Population-Weighted Compactness is:
\begin{equation}
  \text{PWC}(D) = \frac{\sum_i p_i \cdot \|c_i - \bar{c}_D\|^2}{\sum_i p_i}.
  \label{eq:pwc}
\end{equation}
This is the population-weighted variance of tract centroid distances from
the district centroid, equivalently the second moment of inertia of the
population distribution about the centroid.

\subsection{Properties}

\begin{proposition}
  PWC$(D) \geq 0$ with equality iff all tracts have the same centroid.
\end{proposition}

\begin{proof}
  PWC is a weighted sum of squared distances $\|c_i - \bar{c}_D\|^2 \geq 0$,
  so PWC $\geq 0$.
  PWC $= 0$ iff $\|c_i - \bar{c}_D\|^2 = 0$ for all $i$ with $p_i > 0$,
  i.e., iff all populated tracts have centroid $\bar{c}_D$.
  For a single-tract district, PWC $= 0$ exactly. \qed
\end{proof}

\begin{proposition}
  PWC is translation-invariant and scale-dependent.
  If all coordinates are scaled by $\lambda$, PWC scales by $\lambda^2$.
\end{proposition}

This scale-dependence means PWC values are not directly comparable across
projections.
All PWC computations use EPSG:5070 (Albers Equal Area Conic, metres),
so values are in square metres.
For practical comparison, we normalise by state area:
$\text{PWC}_{\text{norm}} = \text{PWC} / (\text{state area})$.

\subsection{Test Invariants}

L0 invariants:
\begin{itemize}
  \item PWC $= 0$ for a single-tract district.
  \item PWC $\geq 0$ always.
  \item PWC is invariant to district relabelling (renaming districts
    does not change any individual district's PWC).
  \item Scaling all coordinates by $\lambda$ scales PWC by $\lambda^2$.
\end{itemize}
