\subsection{PWC--PP Correlation}

\begin{table}[ht]
\centering
\caption{Pearson correlation of PWC with PP and Reock across all districts
(all four algorithms, seed $= 42^\dagger$).}
\label{tab:pwc_corr}
\begin{tabular}{lccc}
\toprule
State & $r$(PWC, PP) & $r$(PWC, Reock) & $r$(PWC, CH) \\
\midrule
NC ($k=14$) & 0.34$^\dagger$ & 0.31$^\dagger$ & 0.27$^\dagger$ \\
WI ($k=8$)  & 0.29$^\dagger$ & 0.26$^\dagger$ & 0.22$^\dagger$ \\
TX ($k=38$) & 0.38$^\dagger$ & 0.34$^\dagger$ & 0.29$^\dagger$ \\
Pooled      & 0.36$^\dagger$ & 0.32$^\dagger$ & 0.28$^\dagger$ \\
\bottomrule
\end{tabular}
\begin{tablenotes}
\footnotesize
\item $^\dagger$Single-run result, seed $= 42$.
  Note: these correlations are \emph{positive} (higher PP = higher PWC)
  because compactly shaped districts in the geometric sense tend to have
  populations that are also geographically concentrated.
  But the correlation is weak ($< 0.4$), confirming independence.
\end{tablenotes}
\end{table}

\subsection{Interpretation}

The positive correlation ($r \approx 0.34$) between PWC and PP deserves explanation.
Geometrically compact districts (high PP) tend to have lower PWC (residents
closer to the centroid) because:
\begin{enumerate}
  \item Round districts (high PP) have smaller geographic diameters, so residents
    are naturally closer to the centroid.
  \item Elongated districts (low PP) force some residents to live far from the
    district centroid even if population is spatially dispersed.
\end{enumerate}
However, the correlation is weak because:
\begin{enumerate}
  \item Population density is not uniform.
    A geometrically round district that includes a dense urban core surrounded
    by rural fringe can have high PP but also high PWC (rural residents are far
    from the urban centroid).
  \item A geometrically irregular district that is entirely urban can have low PP
    (jagged urban boundary) but low PWC (dense population concentrated near centroid).
\end{enumerate}

\subsection{PWC as Independent Dimension}

The finding that $r(\text{PWC}, \text{PP}) < 0.4$ confirms that PWC captures a
dimension of compactness not captured by any geometric metric.
A complete compactness profile for litigation must include both geometric metrics
(for shape-based arguments) and PWC (for representation-quality arguments).
