\section{Conclusion}
\label{sec:conclusion}

Power analysis is a standard component of any rigorous statistical study
but has been absent from redistricting expert testimony.
This paper fills the gap by deriving the power function for redistricting
tests from first principles, translating it into state-specific minimum
ensemble sizes, and evaluating current practice.

The main finding is that ensemble size adequacy depends on the state.
For small delegations ($k \leq 17$), the power challenge is primarily
convergence: once the chain has mixed (5,000--10,000 steps), ESS of
300--1,000 provides high power for any practically significant
gerrymandering effect.
For large delegations ($k > 17$), the power challenge is mixing:
California requires 20,000--50,000 steps to achieve $\rhat < 1.1$,
and the effective ESS at 10,000 steps is based on a chain that is not
fully converged.

The two-tier recommendation---20,000 steps for $k \leq 17$ and 50,000
steps for $k > 17$---provides a clear, state-independent standard that
is computationally feasible (under 1 second per state with \bisect{}'s
ensemble module).

The practical implication for redistricting litigation is that experts
should report power alongside percentile claims.
``The enacted plan is at the 0.4th percentile of a converged ensemble
with $\ess = 1{,}390$ and 99\% power to detect the observed effect''
is a substantially stronger claim than ``the plan is at the 0.4th percentile
of 10,000 plans.''
The former tells the court how reliable the estimate is; the latter does not.
