\section{Method}

The staged method uses deterministic farthest-point seeds, assigns units subject
to population capacity, and records whether the resulting assignment is
\texttt{valid}, \texttt{needs-repair}, or \texttt{infeasible-capacity}. Small
repair routines are treated as auditable post-processing rather than invisible
cleanup.

The atlas view makes the key decision inspectable: a geographically nearest
cluster is only a candidate, not an entitlement. If assigning a unit would
overfill that cluster, the candidate is rejected with visible capacity
arithmetic and the next eligible cluster is tried. The paper's artifact claim
therefore depends on preserving seed order, rejected capacity candidates,
fallback assignments, and final status.

\subsection{Reproducible Procedure}

The constructor operates on the graph and unit attributes supplied by the
BISECT state pipeline. The current slice is graph-based and capacity-oriented:
community-character weights and richer attribute clustering are future or
optional inputs rather than claims of this baseline.

\begin{enumerate}
  \item Select deterministic farthest-point seeds from the available units,
        breaking ties by stable unit identifier.
  \item Assign remaining units to candidate clusters in deterministic order
        while respecting target population capacity and declared slack.
  \item Mark the assignment \texttt{valid} when population deviation and
        connectedness checks pass.
  \item Mark the assignment \texttt{needs-repair} when the assignment is close
        enough to repair but does not yet pass all checks.
  \item Run bounded repair only for \texttt{needs-repair} outputs. In the
        current slice this is an exhaustive small-fixture repair path.
  \item Mark the result \texttt{infeasible-capacity} when the requested capacity
        profile cannot be satisfied. Invalid input is returned as an error
        before a plan artifact is emitted.
\end{enumerate}

\begin{table}[h]
\centering
\small
\resizebox{\linewidth}{!}{%
\begin{tabular}{lll}
\toprule
Status & Meaning & Artifact behavior \\
\midrule
\texttt{valid} & Assignment satisfies declared profile & Write final sidecars with \texttt{repair-status=not-needed} or \texttt{repaired} \\
\texttt{needs-repair} & Assignment failed connectedness or balance checks but may be repairable & Keep repair requirement visible; do not collapse into final export \\
\texttt{infeasible-capacity} & Capacity cannot be satisfied under the declared tolerance & Write witness summary rather than a final plan claim \\
\texttt{invalid input} & Input shape is malformed, such as \(k=0\) or mismatched arrays & Return an error before producing a plan artifact \\
\bottomrule
\end{tabular}
}
\caption{Capacity-clustering status taxonomy. Repair is visible lineage, not
hidden cleanup.}
\end{table}

\subsection{Worked Capacity Decision}

The assignment step is easiest to read as a visible rejection table. Suppose two
clusters target \(50\) population each with ten percent slack, so the local upper
capacity is \(55\). The next unit has population \(20\). The nearest seed is not
automatically allowed to take it; capacity arithmetic decides which candidate
remains eligible.

\begin{table}[h]
\centering
\small
\begin{tabular}{lllll}
\toprule
Candidate cluster & Distance rank & Current pop. & Pop. after unit & Decision \\
\midrule
A & 1 & 45 & 65 & reject: exceeds \(55\) \\
B & 2 & 32 & 52 & accept: within slack \\
\bottomrule
\end{tabular}
\caption{T.15 assigns by deterministic ranked candidates subject to capacity.
The nearest cluster can lose when accepting the unit would overfill it.}
\end{table}
