In \textit{Rucho v. Common Cause}, 588 U.S.\ 684 (2019), the Supreme Court held that partisan gerrymandering claims are nonjusticiable in federal court because ``federal courts are not meant to resolve such political questions.'' But the Court also emphasized what it did not hold: it did not hold that partisan gerrymandering is acceptable, or that no legal remedy exists. ``[N]othing in this opinion should be interpreted as expressing support for partisan gerrymandering,'' Chief Justice Roberts wrote. ``Of course,'' the opinion continued, ``the limits of the federal courts' authority does not leave Americans without a remedy for extreme partisan gerrymandering. The States may regulate congressional elections; they may also provide independent state courts with authority to review legislatively drawn plans for compliance with state constitutional requirements.'' Most pointedly: ``[F]ederal courts are not the only potential check on partisan gerrymandering. Indeed, the plurality noted the 'daunting' challenges federal courts have faced in attempting to adjudicate partisan gerrymandering claims. But the limitations of the judicial role in no way suggest that partisan gerrymandering is permissible.'' \citep{rucho2019}

The implication is direct: Congress may act. Article I, \S4 of the Constitution grants Congress authority to regulate the ``Times, Places and Manner'' of congressional elections, and Congress may override or supplement state regulations. Congress has previously exercised this power to require single-member districts (2 U.S.C.\ \S2c), to prohibit certain election-day practices, and to require equal population among congressional districts in application of \textit{Wesberry v. Sanders}, 376 U.S.\ 1 (1964).

The model federal statute analyzed in this paper---the Districting Integrity Act (DIA)---takes its structural cue from the most successful prior exercise of Congress's redistricting authority: the Apportionment Act of 1941, which mandated the Huntington-Hill equal-proportions method for apportioning House seats among states after each census. That statute resolved a decades-long dispute over apportionment by specifying a single deterministic algorithm whose output any person can verify. The DIA applies the same logic to the drawing of district lines within states.

Section~2 analyzes the constitutional authority for federal redistricting legislation. Section~3 describes the DIA's key architectural choices. Section~4 compares the DIA against four alternative designs. Section~5 develops the historical analogy to the Apportionment Act of 1941 in depth. Section~6 concludes with implications for federal legislative strategy.

The full text of the DIA model statute is maintained in \texttt{docs/legal/MODEL\_FEDERAL\_STATUTE.md} as part of the \texttt{bisect} research program's documentation.
