\section{Introduction and Post-\textit{Rucho} Landscape}
\label{sec:introduction}

On June 27, 2019, the Supreme Court closed the federal courthouse to partisan gerrymandering claims.
In \textit{Rucho v.\ Common Cause},\footnote{\textit{Rucho v.\ Common Cause}, 588 U.S.\ 684 (2019).}
Chief Justice Roberts, writing for a five-Justice majority, held that partisan gerrymandering
claims present political questions beyond the reach of federal courts. The majority acknowledged
that gerrymandering was ``unjust'' and ``incompatible with democratic principles,'' yet concluded
that no judicially manageable standard existed to distinguish permissible partisan line-drawing
from unconstitutional excess.\footnote{\textit{Id.}\ at 693, 721.}

The doctrinal effect was immediate. Federal circuit and district courts ceased adjudicating
partisan gerrymandering claims. The 2020 redistricting cycle---the first following
\textit{Rucho}---produced congressional maps that the federal judiciary largely left intact
regardless of their partisan consequences. State courts stepped into the void with
inconsistent results: Pennsylvania invalidated its congressional map under the state
constitution's ``free and equal elections'' clause;\footnote{\textit{League of Women Voters
v.\ Commonwealth}, 178 A.3d 737 (Pa.\ 2018).} North Carolina's Supreme Court invalidated
its map under the ``free elections'' clause;\footnote{\textit{Harper v.\ Hall}, 868 S.E.2d
499 (N.C.\ 2022).} Ohio's Supreme Court invalidated its map under the Ohio Constitution's
requirement that districts not be drawn to favor a political party;\footnote{\textit{League
of Women Voters of Ohio v.\ Ohio Redistricting Comm'n}, 167 Ohio St.3d 255 (2022).}
but courts in other states declined to reach the merits.

The practical consequence for practitioners is this: partisan gerrymandering litigation
now proceeds almost entirely through state constitutional claims, evaluated under state
law standards that vary in text, interpretation, and evidentiary expectations across
jurisdictions. Expert witnesses in these proceedings must select among six distinct
partisan fairness metrics, each with its own mathematical definition, behavioral properties,
legal history, and known limitations. The selection decision is not merely technical---it
is strategic. A metric endorsed in Pennsylvania state court may not have been cited in
North Carolina proceedings; a threshold adopted in Wisconsin academic commentary may
have been explicitly rejected in federal dicta. Metric-shopping criticism from opposing
counsel can undermine the credibility of any individual metric if the expert cannot
explain why that metric was chosen over the alternatives.

This paper provides a unified framework. Section~\ref{sec:definition} taxonomizes the
six metrics into descriptive and normative categories, explaining what each measures
and how the measures relate to each other. Section~\ref{sec:behavior} introduces
\textsc{bisect} algorithmic maps as a natural reference distribution for state court
proceedings---a class of maps drawn without partisan data that cluster near the
partisan-neutral centroid on all six metrics, providing a non-partisan baseline
against which enacted maps can be compared. Section~\ref{sec:empirical} presents
an empirical overview across four states: North Carolina ($k=14$), Wisconsin ($k=8$),
Texas ($k=38$), and Florida ($k=28$), reporting all six metrics for both algorithmic
and enacted maps. Section~\ref{sec:legal} analyzes the legal landscape, including
the relevant state constitutional provisions, the case law through 2026, and a
decision table for expert witnesses. Section~\ref{sec:conclusion} concludes.

\subsection{What Post-\textit{Rucho} Left Open}

\textit{Rucho}'s holding was deliberately narrow. The Court explicitly preserved
multiple non-federal pathways for partisan gerrymandering reform. Chief Justice
Roberts identified state courts as the primary institutional alternative, writing
that ``[p]rovisions in state statutes and state constitutions can provide standards
and guidance for state courts to apply'' to partisan gerrymandering claims.\footnote{
\textit{Rucho}, 588 U.S.\ at 721.} He also preserved the possibility of federal
legislation, noting that ``[o]ther reform mechanisms are available'' including
``independent redistricting commissions'' and federal statutory action by Congress.\footnote{
\textit{Id.}}

These open pathways are not rhetorical---they have been operationalized. State courts
in Pennsylvania, North Carolina, and Ohio have issued map-invalidating decisions
precisely on the state constitutional grounds the Chief Justice identified as viable.
The metrics in this paper provide the empirical inputs for those proceedings.
The \textsc{bisect} reference distribution provides the conceptual tool that courts
need to move from ``this map is extreme'' (a qualitative assertion vulnerable to
political characterization) to ``this map lies $X$ standard deviations from what
a neutral process produces'' (a quantitative claim that translates partisan fairness
into the statistical framework courts are accustomed to in other expert testimony contexts).

\subsection{The Six Metrics: A Preview}

The L-series papers (L.1 through L.6) provide dedicated treatments of each metric.
This paper provides the integrating framework:

\begin{itemize}
\item \textbf{Efficiency Gap} (L.1): $\text{EG} = (\text{Wasted}_D - \text{Wasted}_R)/\text{Total}$.
  Measures asymmetry of wasted votes. Introduced by \citet{stephanopoulosMcGhee2015}.
  8\% threshold proposed but not legally adopted.
\item \textbf{Mean-Median Difference} (L.2): $\text{MM} = \bar{v}_D - \tilde{v}_D$.
  Measures the gap between mean and median district vote share.
  Introduced by \citet{wang2016}. Less sensitive to election swings than EG.
\item \textbf{Partisan Bias} (L.3): $\text{Bias} = S(0.50) - 0.50$.
  Symmetry standard at the 50\% pivot. Part of the framework in
  \citet{grofmanKing2007}.
\item \textbf{Declination} (L.4): $\delta = 2(\theta_D - \theta_R)/\pi$.
  Geometric measure of packing and cracking. Introduced by \citet{warrington2018}.
  No court has adopted it as of 2026.
\item \textbf{Seats-Votes Curve and Responsiveness} (L.5): $S(v)$, $R = dS/dv|_{v=0.50}$.
  The sufficient statistic from which EG and Bias are recoverable. Best courtroom exhibit.
\item \textbf{Proportionality / Gallagher Index} (L.6):
  $\text{LSq} = \sqrt{\tfrac{1}{2}\sum_i(v_i-s_i)^2}$.
  Normative standard; no federal constitutional requirement post-\textit{Rucho}.
\end{itemize}
