\section{Implementation Contract}

The current implementation centers on a one-move boundary improvement baseline
plus structured statuses for no-improvement, rejected-move, and staged-method
paths. The summary must record parent plan id, attempted method, accepted move
or rejection/no-op reason, objective delta, validity checks, and output artifact
id when a child plan is emitted.

The atlas framing treats every boundary move as a candidate that must pass
separate gates. A move that improves the edge-cut objective can still be
rejected for population or connectedness, and a no-op run is meaningful only if
the evaluated neighborhood and rejection reasons are recorded. The descendant
package is therefore evidence of a local transition from a parent plan, not a
new construction from scratch.

The L0 improvement fixture makes the transition small enough to inspect. The
starting assignment is \([0,0,0,1,1,1]\) on a six-unit graph with equal
populations and two districts. Unit 2 touches the opposite district through
units 3, 4, and 5. Moving unit 2 from district 0 to district 1 preserves
population tolerance at 34 percent, keeps both districts connected, and reduces
edge cut from 3 to 1. The emitted child assignment is
\([0,0,1,1,1,1]\). A competing move that improves edge cut but breaks population
or contiguity would be recorded as evaluated and rejected, not silently treated
as an improvement.

\begin{table}[htbp]
\centering
\small
\begin{tabular}{p{0.20\linewidth}p{0.36\linewidth}p{0.30\linewidth}}
\toprule
Outcome & Required summary & Claim supported \\
\midrule
Improved & parent, child, move, delta, validity checks & local improvement \\
Rejected move & parent, candidate move, rejection reason & validity-preserving search \\
No-op & parent, search status, reason & reproducible non-improvement \\
Staged method & requested method and unsupported scope & roadmap boundary \\
\bottomrule
\end{tabular}
\caption{Improvement-family statuses.}
\end{table}

\begin{table}[htbp]
\centering
\small
\begin{tabular}{p{0.24\linewidth}p{0.30\linewidth}p{0.32\linewidth}}
\toprule
Summary field & Current value family & Interpretation \\
\midrule
Summary schema & local-search summary v1 & Stable JSON envelope for transition evidence \\
Method & one-move, staged tabu, staged LNS & Search surface requested \\
Status & improved or no-improvement & Whether a child assignment changed \\
Move counts & evaluated and accepted counts & Why no-op runs are inspectable \\
Edge cut & initial and final values & Objective delta evidence \\
Population deviation & initial and final values & Validity gate evidence \\
Parameter hash & SHA-256 digest & Tamper check for summary-defining inputs \\
\bottomrule
\end{tabular}
\caption{Concrete local-search summary fields emitted by \texttt{bisect-local-search}.}
\end{table}

The paper therefore does not treat every command invocation as a successful
improvement. A no-op transcript is evidence that the search was run under a
declared method and profile. A rejected-move transcript is evidence that the
implementation preserved validity by refusing a candidate. Only an
\emph{Improved} status supports the claim that a descendant plan was emitted
with a measured local objective delta.
