Thomas Jefferson, as Secretary of State and later as President, was the
first American statesman to analyse the apportionment problem mathematically.
His 1792 memo to President Washington argued that the correct method of
apportionment was to divide each state's population by a common divisor
and take the floor --- exactly what we now call the Jefferson method
\citep{schmeckebier1941congressional}.
Jefferson's method was embedded in the Apportionment Act of 1792 and
used for the next six apportionments (1790, 1800, 1810, 1820, 1830,
and 1840, with the 1840 apportionment being the last).

The political controversy surrounding Jefferson's method was apparent
early.
After the 1830 Census, New York's apportionment under Jefferson's method
gave it 40 seats when its exact quota was only 38.59 ---
a clear violation of the quota rule.
Daniel Webster led the congressional opposition, arguing that Jefferson's
method was a ``palpable violation of the Constitution'' \citep{balinski1982fair}.
The 1842 Apportionment Act replaced Jefferson with Webster's method.

Simultaneously, and entirely independently, the Belgian mathematician
Victor d'Hondt published in 1882 a method for allocating seats among
political parties in proportional representation elections
\citep{dhondt1882systeme}.
D'Hondt's formula --- assign the next seat to the party with the highest
priority $V_j/s_j$, where $V_j$ is the party's vote total ---
is mathematically identical to Jefferson's priority formula $P_i(s_i) = p_i/s_i$.
The two methods were independently derived for different applications
but converge to the same algorithm.

Today, d'Hondt is the world's most widely used seat allocation formula.
Belgium, Spain, Portugal, the Netherlands, Brazil, Argentina, Turkey,
and many other countries use d'Hondt for their parliamentary elections.
Jefferson's method is thus simultaneously the first U.S.\ apportionment
method (abandoned for domestic use) and the dominant global PR formula
(adopted by much of the democratic world).

This paper's contributions:
\begin{enumerate}
  \item \textbf{Equivalence proof}: We prove that Jefferson and d'Hondt
    are mathematically identical (Section~\ref{sec:dhondt}).
  \item \textbf{Bias characterisation}: We prove that Jefferson
    maximises aggregate representation for large states and show
    the quantitative effect on 2020 Census data (Section~\ref{sec:bias}).
  \item \textbf{Historical documentation}: We trace Jefferson's use
    in six U.S.\ apportionments and the political controversy
    that led to its abandonment (Section~\ref{sec:history}).
  \item \textbf{Implementation}: We document the \textsc{bisect-apportion}
    crate's Jefferson/d'Hondt implementation for both U.S.\ apportionment
    and PR seat allocation (Section~\ref{sec:implementation}).
\end{enumerate}
