\label{sec:empirical}

\subsection{2020 Census Jefferson Apportionment}

Table~\ref{tab:jefferson-2020} shows the Jefferson seat allocation
for all states that differ from Huntington-Hill.

\begin{table}[h]
\centering
\caption{States differing between Jefferson and Huntington-Hill,
  2020 Census (selected; see J.0 Table~1 for complete comparison).}
\label{tab:jefferson-2020}
\begin{tabular}{lrrrr}
\toprule
State & Population & Exact quota & Jefferson & Huntington-Hill \\
\midrule
California & 39,538,223 & 51.95 & 53 & 52 \\
Minnesota  & 5,706,494  & 7.50  &  7 &  8 \\
Rhode Island & 1,097,379 & 1.44 &  1 &  2 \\
Montana    & 1,084,225  & 1.42  &  1 &  2 \\
\bottomrule
\end{tabular}
\begin{tablenotes}
\small
\item The Jefferson divisor for 2020 is adjusted (via binary search)
  to produce exactly 435 total seats; it is lower than the HH divisor,
  allowing large states to accumulate extra floor seats.
  California gains 1 seat under Jefferson relative to HH (53 vs.\ 52),
  consistent with Jefferson's large-state-favoring bias.
  Minnesota, Rhode Island, and Montana each lose 1 seat relative to HH.
\end{tablenotes}
\end{table}

The most significant divergence is Minnesota, which has exact quota
$\approx 7.50$: exactly at the arithmetic mean threshold.
Under Jefferson (floor), Minnesota gets 7 seats; under HH (geometric mean),
the threshold is $\sqrt{7 \times 8} = \sqrt{56} \approx 7.483 < 7.50$,
so Minnesota gets 8 seats under HH.
Under Webster (threshold 7.5), Minnesota is at the exact boundary ---
rounding half-up gives 8 seats under Webster.
California, by contrast, gains a seat under Jefferson (53 vs.\ HH's 52)
because the lower Jefferson divisor allows its large absolute floor quotient
to capture an additional seat.

\subsection{Historical Census Comparison}

Jefferson was used for six U.S.\ apportionments (1790--1840).
The method's large-state bias was documented contemporaneously.
After the 1832 reapportionment, New York received 40 seats
when its exact quota (under the national divisor) was 38.59 seats
--- an overrepresentation of more than 1 seat.
This quota violation (New York receiving more than its upper quota)
was politically controversial and was cited by Webster as evidence
that Jefferson's method was constitutionally problematic
\citep{balinski1982fair}.
The 1842 switch to Webster resolved this concern by ensuring that
no state would receive more than $\lceil q_i \rceil$ seats.

\subsection{D'Hondt vs.\ Hamilton: International Implications}

The Jefferson-d'Hondt equivalence creates an interesting parallel
to the Hamilton method:
just as U.S.\ apportionment history moved away from Hamilton
(quota method, paradox-susceptible) toward divisor methods,
some European PR systems have moved away from d'Hondt toward
Sainte-Lagu\"e (Webster) on grounds that d'Hondt's large-party
bias is undesirable.
Norway, Sweden, and Denmark use Sainte-Lagu\"e (with a first-seat
divisor of 1.4 rather than 1 to discourage small parties).
This international movement reflects the same normative judgment
that motivated Webster's 1832 critique: floor rounding is unfair
to small entities.
