\section{Negative Fixtures}

The hash layer earns its keep when adjacent checks disagree. Two negative
fixtures show different failures:

\begin{center}
\small
\begin{tabularx}{\textwidth}{lXX}
\toprule
Fixture & Failure & Interpretation \\
\midrule
\texttt{tampered-source} & \texttt{source\_hash\_match} fails after raw source bytes are edited. & Arithmetic can still pass, but the evidence file no longer matches the declared hash. \\
\texttt{missing-source-hash} & \texttt{source\_hash\_match} fails because the source index is empty. & The package omits the raw source-hash evidence needed for replay. \\
\bottomrule
\end{tabularx}
\end{center}

In both cases, local contest sums and jurisdiction totals can pass. That is the
point: hash failures are not arithmetic failures. A good verifier transcript
keeps them separate so a public reader can ask the right follow-up question.

\begin{center}
\small
\begin{tabularx}{\textwidth}{lXX}
\toprule
Fixture & Expected result & Carrying checks \\
\midrule
\texttt{summary-basic} & pass & \texttt{contest\_selection\_sum}, \texttt{jurisdiction\_contest\_total}, \texttt{source\_hash\_match}. \\
\texttt{mail-batch-added} & pass & \texttt{accepted\_ballots}, \texttt{batch\_summary\_total}, \texttt{source\_hash\_match}. \\
\texttt{tampered-source} & fail & \texttt{source\_hash\_match} reports a declared/computed hash mismatch. \\
\texttt{missing-source-hash} & fail & \texttt{source\_hash\_match} reports an empty source index. \\
\bottomrule
\end{tabularx}
\end{center}
