\section{Boundaries}

Hashing is not certification. A passing RCOUNT hash check says that declared
source bytes still match their recorded hashes. It does not say that an election
official certified the result, that the source export was complete, or that a
court would accept the record.

Hashing is also not encryption. If a source file contains private data, hashing
the file does not make the data safe to publish. RCOUNT packages should publish
only receipt-safe public evidence, and private support files should be handled
under the jurisdiction's custody rules.

Finally, tamper evidence is not malware resistance or end-to-end cryptographic
voting. RCOUNT can detect that a published source file changed after hashing. It
cannot prove that a tabulator, ballot-marking device, or export pipeline was
correct before that source file was produced.

\begin{center}
\small
\begin{tabularx}{\textwidth}{lXX}
\toprule
Concern & RCOUNT hash layer can show & It does not show \\
\midrule
Source byte change & Declared raw bytes no longer match. & Who changed them or why. \\
Source omission & Source index is empty or file is missing. & Whether another office record exists elsewhere. \\
Package mutation & Normalized content hash changed. & Whether the old or new content is legally correct. \\
Parser substitution & Different normalized records can be detected by package hash. & That the parser captured every source nuance. \\
Private data & Nothing; hashes are not redaction. & That sensitive files are safe to publish. \\
Malware or tabulator error & Nothing directly. & Device or export-pipeline correctness. \\
Certification & Evidence consistency. & Legal certification or court acceptance. \\
\bottomrule
\end{tabularx}
\end{center}

For public-record workflows, a certification artifact can itself be a hashed
source. That still separates evidence integrity from the legal act of
certification. Detached signatures, source-custody attestations, and superseding
or amended source packages are future work.
