\documentclass[12pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage{algorithm}
\usepackage{algorithmic}

% Define theorem environments
\newtheorem{proposition}{Proposition}

\title{Cross-Census Temporal Stability:\\
Recursive Bisection vs N-Way Partitioning Over Decades}
\author{Giovanni Della-Libera}
\date{\today}

\begin{document}

\maketitle

\begin{abstract}
Congressional redistricting occurs every decade following the census, raising questions about how algorithmic redistricting methods respond to demographic change across census cycles. This paper provides the first empirical analysis of cross-census boundary variation for the \textsc{bisect} recursive bisection algorithm, comparing actual 2010 and 2020 pipeline outputs for five southern states with significant minority populations (Alabama, Georgia, Louisiana, Mississippi, South Carolina).

We measure county-level disruption (fraction of counties whose district-split count changed from 2010 to 2020 runs) and top-level bisection stability (whether counties remain in the same primary geographic half). Mean county disruption is 66.1\% (range: 58.5--69.2\%), substantially higher than prior projection-based estimates. Top-level stability ranges from 35.4\% (Mississippi) to 86.2\% (Georgia), reflecting state-specific demographic dynamics: fast-growing metros and declining rural regions drive higher cross-census variation regardless of algorithm.

The key finding is that observed disruption reflects \emph{demographic responsiveness}, not algorithmic instability. The bisect algorithm applies identical parameters (geographic edge weights, $\pm$0.5\% balance tolerance) across census cycles; any boundary changes are attributable to updated population data, not strategic manipulation. This property---algorithmic consistency paired with demographic responsiveness---provides a stronger legal defence than geographic continuity: courts can verify that both the 2010 and 2020 maps were produced by the same neutral process responding to census data.

Comparison with n-way partitioning is deferred to Phase 2 pending availability of n-way 2010 run outputs.
\end{abstract}

\section{Introduction}
\input{sections/01_introduction}

\section{Background}
\input{sections/02_background}

\section{Methodology}
\input{sections/03_methodology}
\input{sections/03_section_3.5_hierarchical_validation}
\input{sections/03_section_3.6_theoretical_foundation}

\section{Results}
\input{sections/04_results}
\input{sections/04_section_4.4_vra_compliance}

\section{Discussion}
\input{sections/05_discussion}

\section{Limitations}
\input{sections/06_limitations}

\section{Future Work}
\input{sections/07_future_work}

\section{Conclusion}
\input{sections/08_conclusion}

\bibliographystyle{plain}
\bibliography{references}

\end{document}
