\label{sec:timeline}

\subsection{Statutory Timeline Under P.L.\ 94-171}

The Redistricting Data Program (P.L.\ 94-171, codified at 13 U.S.C.\ \S 141(c))
requires the Census Bureau to deliver redistricting data to state governors and
legislative leaders within one year of census day (April 1, 2030). In practice:

\begin{itemize}
\item \textbf{Census Day}: April 1, 2030.
\item \textbf{P.L.\ 94-171 Statutory Deadline}: April 1, 2031 (one year after census day).
\item \textbf{Historical Release Pattern}: The 2020 census data was released August 12, 2021
  (133 days after the April 1, 2021 statutory deadline, which had been extended from April 1,
  2021 to September 30, 2021 due to COVID-19). Under normal conditions, the 2010 census
  redistricting data was released February 3, 2011, approximately one month early.
\item \textbf{2030 Projection}: Assuming no exceptional delays, redistricting data release
  in March--April 2031. The analysis in this paper uses April 1, 2031 as the baseline
  release date.
\end{itemize}

\subsection{State Redistricting Deadlines}

Following data release, states must complete redistricting by their own constitutional
and statutory deadlines. These deadlines vary:

\begin{table}[ht]
\centering
\caption{Representative State Redistricting Deadlines Post-2031 Data Release}
\label{tab:deadlines}
\begin{tabular}{lll}
\toprule
State & Deadline Type & Approximate Deadline \\
\midrule
Texas & Legislative session calendar & June 2031 \\
Florida & Commission ($\sim$180 days) & October 2031 \\
North Carolina & Court order / legislative & December 2031 \\
California & Commission (174 days from data) & September 2031 \\
Michigan & Commission (360 days from data) & March 2032 \\
New York & Legislative / commission & January 2032 \\
\bottomrule
\end{tabular}
\begin{tablenotes}
\small
\item Deadlines are approximate; states may adjust based on litigation.
  Texas's legislative session calendar creates the tightest effective deadline.
\end{tablenotes}
\end{table}

The operational consequence is that states with early legislative session deadlines
(Texas, Ohio, Georgia) have effectively shorter redistricting windows than the
12-month statutory limit. Texas has a biennial regular session ending May 31; if
redistricting is not completed by then, a special session (which the Governor must
call) is required at additional cost and political difficulty. The 2021 redistricting
cycle saw Texas complete its congressional maps in October 2021 via special session.

For \textsc{bisect} to be useful in the most time-constrained states, the platform
must be capable of producing validated plans within weeks of data release, not months.
This requires that all infrastructure be in place before April 2031.

\subsection{The 2030 Infrastructure Preparation Window}

Working backward from April 2031:

\begin{itemize}
\item \textbf{April 2031}: Census data released; redistricting begins.
\item \textbf{January 2031}: Final validation of block-group adjacency graphs complete.
  No new graph construction should occur during the redistricting window.
\item \textbf{July 2030}: Algorithm version pre-registration complete (see Section~\ref{sec:dia-seed}).
  Configuration locked for 2031 redistricting.
\item \textbf{January 2030}: Block-group adjacency graph construction begins for all 50 states.
  Target completion: January 2031 (12 months for construction + validation).
\item \textbf{July 2029}: LODES 2028 and ACS 2026--2028 data available; demographic update
  pipeline deployed and validated.
\item \textbf{January 2029}: Reapportionment projection analysis complete (Q.4).
  States projected to gain/lose seats identified; $k$-values pre-configured.
\end{itemize}

The 2029--2031 preparation window is 24 months. It is tight but achievable.
The Q-series documents the work required in each phase.
