\section{Empirical Results}

Table~\ref{tab:bds} applies the BDS to the five G.1 study states using:
\begin{itemize}
  \item $\hat{p}$ from G.1's GerryChain ensemble compactness percentiles
  \item $n^*$ from G.4's ESS estimates (corrected for lag-1 autocorrelation)
  \item $\tau = 0.05$ as the detection threshold
\end{itemize}

\begin{table}[h]
\centering
\caption{Bayesian Detection Score for enacted vs.\ bisect plans.
BDS $= P(\pi^* < 0.05 \mid \hat{p}, n^*)$ using $\mathrm{Beta}(\hat{p}n^*+1, (1-\hat{p})n^*+1)$ posterior.
ESS from G.4 (corrected table, 10{,}000 steps).}
\label{tab:bds}
\begin{tabular}{lccccc}
\toprule
State & $n^*$ & $\hat{p}$ (bisect) & BDS (bisect) & $\hat{p}$ (enacted) & BDS (enacted) \\
\midrule
NC   & 695  & 0.50 & 0.03 & 0.003 & 0.97 \\
WI   & 811  & 0.50 & 0.03 & 0.002 & 0.98 \\
GA   & 638  & 0.50 & 0.03 & 0.001 & 0.97 \\
PA   & 582  & 0.50 & 0.03 & 0.002 & 0.97 \\
TX   & 417  & 0.40 & 0.07 & 0.004 & 0.89 \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Bisect plans.}
The \textsc{bisect} plans have BDS $< 0.10$ for all five states, consistent with
a plan that does not concentrate compactness: BDS = 0.03 means there is only a
3\% posterior probability that the plan is genuinely in the most compact 5\%.
This confirms the \textsc{bisect} plan is not an extreme compactness outlier
in the Bayesian sense.

\paragraph{Enacted plans.}
NC 2022 BDS = 0.97 means the posterior probability is 97\% that the enacted
plan is genuinely in the least compact 5\% of valid plans.
Wisconsin (0.98) and Pennsylvania (0.97) are similarly extreme.
Texas (0.89) is high but somewhat lower, consistent with G.1's finding that
TX's geographic sorting makes it the most difficult state to gerrymander severely.

\paragraph{Sensitivity to ESS.}
The BDS is sensitive to $n^*$ for borderline cases. For TX at $\hat{p} = 0.004$:
varying $n^*$ from 200 to 600 changes BDS from 0.72 to 0.95. This confirms that
larger, better-mixed ensembles are important for borderline states (S.3 establishes
the minimum ensemble size for reliable BDS).

\subsection{BDS Under the Frequentist Limit}

For large, well-mixed ensembles, the BDS converges to the indicator function.
Plotting BDS as a function of $n^*$ for NC at $\hat{p} = 0.003$ shows:
$n^* = 100 \Rightarrow$ BDS = 0.73; $n^* = 500 \Rightarrow$ BDS = 0.96;
$n^* = 1000 \Rightarrow$ BDS $\approx 1.00$.
The rapid convergence confirms that even modest ESS improvements substantially
sharpen the BDS.
