\documentclass[12pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{amsmath, amssymb, amsthm}
\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}
\newtheorem{proposition}{Proposition}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{remark}{Remark}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage{booktabs}
\usepackage{xcolor}
\usepackage{multirow}
\usepackage{longtable}
\usepackage{array}
\usepackage{microtype}
\usepackage{setspace}
\usepackage{titlesec}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{lmodern}
\usepackage[T1]{fontenc}

%% Style
\titleformat{\section}{\normalfont\large\bfseries}{\thesection.}{0.5em}{}
\titleformat{\subsection}{\normalfont\normalsize\bfseries}{\thesubsection.}{0.5em}{}
\titleformat{\subsubsection}{\normalfont\normalsize\itshape}{\thesubsubsection.}{0.5em}{}

\setstretch{1.4}

%% Macros
\newcommand{\ar}{\textsc{ApportionRegions}}
\newcommand{\hh}{\textsc{Huntington-Hill}}
\newcommand{\dia}{\textsc{Districting Integrity Act}}
\newcommand{\recom}{\textsc{ReCom}}
\newcommand{\case}[1]{\textit{#1}}

\title{%
  \textbf{Redistricting Stability Across Reapportionment Cycles:}\\
  \textbf{How Seat Changes Affect the ApportionRegions Tree}\\[0.8em]
  \normalsize\textit{T.9 --- Algorithm Theory / Temporal Stability}%
}

\author{%
  Gio Della-Libera\\
  \textit{Apportionment Research Group}\\
  \small\texttt{giodl@microsoft.com}
}

\date{May 2026}

\begin{document}

\maketitle

\begin{abstract}
\noindent
What happens in 2030 if states gain or lose congressional seats?
The \ar{} algorithm constructs a bisection tree whose structure is
determined by the split prescription for the district count $k$: composites
split by their largest prime factor, while large primes use a binary
floor/ceiling fallback.
A change in $k$ can therefore produce a dramatic change in tree
topology---not because the algorithm has changed, but because the
arithmetic has.
This paper characterises an illustrative 2030 reapportionment scenario
(TX: $38 \to 41$, CA: $52 \to 50$, FL: $28 \to 29$, NY: $26 \to 25$,
and others), traces how each seat change alters the split tree, and
defines the package evidence needed to measure tract-level boundary
movement.
The paper's current empirical tables should be read as scenario/audit
targets unless accompanied by archived run manifests.
In that scenario, the largest reported disruption is Texas at
$d_{\rm Ham}=0.23$, algebraically equivalent to roughly 9--10 one-step
\recom{} units under the approximation $1/k$.
That comparison is a scale heuristic, not an empirical mixing result.
The reapportionment adjustment itself is deterministic once the input
data, district count, seed, graph, and split prescription are fixed.
The \ar{} algorithm recomputes the tree fresh from the new
apportionment each decade, which is the intended statutory
behavior under the \dia{}.
Texas is the most interesting case: $38 = 2 \times 19$ (large-factor,
largest-prime-first tree) changes to $41$ (large-prime binary fallback),
producing
the largest structural change in the nation.
California's move from $52 = 4 \times 13$ to $50 = 2 \times 5^2$
produces a qualitatively different but similarly disruptive
factorisation change.
We conclude that decadal recomputation is the right algorithmic staging,
while stronger claims about disruption magnitude or political neutrality
require archived comparison packages.
\end{abstract}

\tableofcontents
\newpage

\input{sections/01-introduction}
\input{sections/02-projections}
\input{sections/03-factorization}
\input{sections/04-boundary-stability}
\input{sections/05-gerrychain-baseline}
\input{sections/06-statutory}
\input{sections/07-conclusion}

\bibliographystyle{plainnat}
\bibliography{references}

\end{document}
