\section{Results: The 0.3 D-Seat Maximum}
\label{sec:results}

\subsection{Baseline Comparison}

We compare three configurations:
(1)~the DIA default ($\theta_{\rm default}$),
(2)~the most Republican-favorable configuration ($\theta_{\rm Rep}$), and
(3)~the most Democratic-favorable configuration ($\theta_{\rm Dem}$).

\begin{table}[h]
\centering
\caption{National 50-state partisan outcomes under three parameter
         configurations. $\Dnat$: expected Democratic seats, 435-seat delegation.
         $\overline{\PP}$: mean Polsby-Popper compactness. County splits:
         total split-county instances nationally.
         $\Delta D$ relative to default.}
\label{tab:main-results}
\begin{tabular}{lccccc}
\toprule
Configuration & $\theta$ & $\Dnat$ & $\Delta D$ & $\overline{\PP}$ & County splits \\
\midrule
Most Rep-favorable & $\theta_{\rm Rep}$ & 210.7 & $-0.3$ & 0.296 & 198 \\
DIA default & $\theta_{\rm default}$ & 211.0 & 0.0 & 0.309 & 271 \\
Most Dem-favorable & $\theta_{\rm Dem}$ & 211.0 & $0.0$ & 0.320 & 315 \\
\midrule
\multicolumn{2}{l}{Range ($\theta_{\rm Rep}$ to $\theta_{\rm Dem}$)} &
  0.3 & & 0.024 & 117 \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Key Finding}

The most Republican-favorable parameters produce 210.7 expected Democratic
seats; the most Democratic-favorable parameters produce 211.0 seats.
The difference is exactly 0.3 Democratic seats nationally---confirming
the U.2 result in the adversarial framing.

Note that the most Democratic-favorable configuration coincidentally
matches the default in partisan outcome ($\Dnat = 211.0$).
This is because lowering $\acounty$ and $\ufactor$ slightly increases
compactness in Democratic-leaning urban areas, but this effect is
exactly offset by the lower $T$ setting's slightly less optimised
leaf-level cuts.

\subsection{State-Level Breakdown}

The 0.3-seat national difference is distributed across states as follows.
In the worst case, a single state (Wisconsin, $k = 8$) contributes
$\Delta D_{\rm WI} = 0.3$ seats at the joint parameter extremes.
No other state contributes more than 0.1 seats.
This concentration in Wisconsin reflects the state's geographic
peculiarity: the metropolitan Milwaukee footprint is large relative
to the state area, and the bisection cut through it is sensitive
to county weighting.

\begin{table}[h]
\centering
\caption{State-level partisan effect of moving from $\theta_{\rm Rep}$
         to $\theta_{\rm Dem}$. Only states with $|\Delta D| > 0$ shown.}
\label{tab:state-results}
\begin{tabular}{llcc}
\toprule
State & $k$ & $\Delta D$ (Rep $\to$ Dem) & Direction \\
\midrule
Wisconsin & 8 & $+0.3$ & Dem-favored at lower $\acounty$ \\
Michigan & 13 & $+0.1$ & Dem-favored at lower $\ufactor$ \\
Minnesota & 8 & $-0.1$ & Rep-favored at lower $\acounty$ \\
\midrule
All other states & -- & 0.0 & No change \\
\midrule
National total & 435 & $+0.3$ & \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Comparison to B.7 Seed Noise}

B.7~\citep{b7paper} establishes the seed-variance noise floor for partisan
outcomes: the standard deviation across random seed choices is $\sigma \approx 0.8$
seats nationally.
The 0.3-seat parameter gaming effect is:
\[
  \frac{\Delta \Dnat^{\rm parameter}}{\sigma} = \frac{0.3}{0.8} = 0.375
  \text{ standard deviations}.
\]
Parameter gaming is below the one-sigma threshold.
A neutral observer cannot distinguish between a map produced by the most
Republican-favorable parameters and one produced by the default parameters
using only the partisan outcome.

\subsection{Comparison to Algorithm-Structure Effect}

B.0~\citep{b0paper} reports that switching from unweighted bisection to
GeoSection produces a partisan gap of approximately 18 seats nationally.
The parameter-gaming maximum of 0.3 seats is $60\times$ smaller.

This comparison is the central policy argument:
the real design lever is algorithm structure (which the DIA pre-specifies
in statute), not parameter tuning.
Locking algorithm structure via statute while allowing administrative
parameter flexibility is safe because parameter flexibility produces
at most 0.3 seats of variation---below the noise floor.
