\subsection{Minimum Bounding Rectangle (MBR)}

The Minimum Bounding Rectangle (MBR) of a polygon $D$ is the smallest-area
rectangle (at any orientation) that contains $D$.
Formally, for each orientation $\theta \in [0, \pi/2)$, let $B_\theta(D)$ be
the axis-aligned bounding box of $D$ rotated by $\theta$.
The MBR is:
\[
  \text{MBR}(D) = \arg\min_{\theta \in [0,\pi/2)} A(B_\theta(D)).
\]

\begin{proposition}
  The MBR of a convex polygon has one side flush with an edge of the polygon.
\end{proposition}

This is the rotating calipers theorem: the optimal bounding rectangle is
always flush with one convex hull edge.

\subsection{Length-Width Ratio}

Let $\ell$ (length) and $w$ (width) with $\ell \geq w$ be the dimensions of MBR$(D)$.
Then:
\begin{equation}
  \text{LW}(D) = \frac{\ell}{w} \geq 1.
  \label{eq:lw}
\end{equation}
LW $= 1$ for a square; LW $= 3$ for a $1 \times 3$ rectangle.

\subsection{AABB vs.\ MBR}

The Axis-Aligned Bounding Box (AABB) reports:
\[
  \text{AABB}(D) = [\min x, \max x] \times [\min y, \max y],
\]
with LW$_{\text{AABB}}(D) = (\max x - \min x) / (\max y - \min y)$
(or the reciprocal if height exceeds width).

\begin{proposition}
  $\text{LW}_{\text{MBR}}(D) \leq \text{LW}_{\text{AABB}}(D)$
  is not guaranteed; AABB can both over- and under-estimate true LW.
\end{proposition}

The key property that makes AABB unreliable:
\[
  A(\text{AABB}(D)) \geq A(\text{MBR}(D)),
\]
but the dimensions of AABB are not monotone transforms of the MBR dimensions.

\subsection{Test Invariants}

L0 invariants:
\begin{itemize}
  \item LW $\geq 1$ for any polygon.
  \item LW of a square $= 1.0$ (exact).
  \item LW of a $1 \times 3$ rectangle $= 3.0$ (exact, for MBR).
  \item LW is rotation-invariant when computed via MBR.
  \item MBR area $\leq$ AABB area for any polygon.
\end{itemize}
