We quantify the AABB vs.\ MBR discrepancy for NC, WI, and TX districts
to document the magnitude of AABB's orientation bias.

\subsection{Discrepancy Measure}

We define the AABB overestimate ratio:
\[
  \text{Disc}(D) = \frac{\text{LW}_{\text{AABB}}(D)}{\text{LW}_{\text{MBR}}(D)} - 1,
\]
so Disc $= 0$ means AABB and MBR agree; Disc $= 0.41$ means AABB overestimates
LW by 41\%.

\subsection{Empirical Discrepancy}

\begin{table}[ht]
\centering
\caption{AABB vs.\ MBR LW discrepancy statistics (ratio-optimal, seed $= 42^\dagger$).}
\label{tab:discrepancy}
\begin{tabular}{lccc}
\toprule
State & Mean Disc & Max Disc & Median Disc \\
\midrule
NC ($k=14$) & 0.12$^\dagger$ & 0.38$^\dagger$ & 0.09$^\dagger$ \\
WI ($k=8$)  & 0.09$^\dagger$ & 0.31$^\dagger$ & 0.07$^\dagger$ \\
TX ($k=38$) & 0.14$^\dagger$ & 0.41$^\dagger$ & 0.11$^\dagger$ \\
\bottomrule
\end{tabular}
\begin{tablenotes}
\footnotesize
\item $^\dagger$Single-run result, seed $= 42$.
  Positive Disc means AABB overestimates LW (AABB LW $>$ MBR LW).
\end{tablenotes}
\end{table}

The maximum discrepancy of $0.41$ (TX) confirms Claim~3: AABB can overestimate
LW by up to 41\% for diagonally oriented districts.
In practice, TX has more diagonally oriented districts than NC or WI because
several TX congressional districts run diagonally across the state boundary.

\subsection{When AABB Underestimates LW}

In rare cases, AABB can underestimate LW (Disc $< 0$).
This occurs when a district's main axis is nearly cardinal (north-south or
east-west) but has minor appendages that widen the AABB in the perpendicular
direction, making the AABB appear more equidimensional than the MBR.
The frequency of underestimation is $< 5\%$ of districts across our dataset,
with maximum underestimate $< 10\%$---smaller in magnitude than the overestimate cases.

\subsection{Recommendation}

AABB-based LW is \emph{not legally defensible} because it is orientation-dependent.
A court or opposing expert can challenge any AABB-based LW value by rotating
the district geometry and obtaining a different result.
Only MBR-based LW (rotating calipers) is rotation-invariant and thus legally credible.
