\section{Three-Level Provenance Architecture}
\label{sec:provenance}

\subsection{Architecture Overview}

The \bisect{} audit chain comprises three levels, each hashing the output
of the previous level and the inputs at that level.

\begin{definition}[bisect Manifest]
A \bisect{} manifest $\mathcal{M}$ is a structured JSON document containing:
\begin{align*}
\mathcal{M} = \{&
  \texttt{algorithm\_hash}: h_{\rm alg},\\
  &\texttt{config\_hash}: h_{\rm cfg},\\
  &\texttt{download\_hashes}: \{h_{\rm pl94171}, h_{\rm tiger}, h_{\rm acs}\},\\
  &\texttt{adjacency\_hash}: h_{\rm adj},\\
  &\texttt{plan\_hash}: h_{\rm plan},\\
  &\texttt{manifest\_hash}: h_{\rm manifest}
\}
\end{align*}
where $h_{\rm manifest} = \mathrm{SHA256}(\mathcal{M} \setminus \{h_{\rm manifest}\})$
is a self-referential integrity hash.
\end{definition}

\subsection{Level 1: Data Provenance}

Level 1 establishes that the algorithm used the Census Bureau's published
data and not a modified version.
The three Level 1 hashes are:
\begin{itemize}
\item $h_{\rm pl94171} = \mathrm{SHA256}(\text{P.L.\ 94-171 file})$:
  verified against Census Bureau's published checksum at
  \url{https://www2.census.gov/}.
\item $h_{\rm tiger} = \mathrm{SHA256}(\text{TIGER/Line shapefile ZIP})$:
  verified against Census Bureau's TIGER archive.
\item $h_{\rm acs} = \mathrm{SHA256}(\text{ACS 5-year table CSV})$:
  verified against Census Bureau's ACS data portal.
\end{itemize}

Any party can independently verify these hashes by downloading the same
files from the Census Bureau and computing SHA-256.

\subsection{Level 2: Computation Provenance}

Level 2 establishes that the adjacency graph was correctly constructed
from the Level 1 data.
\[
  h_{\rm adj} = \mathrm{SHA256}(G(h_{\rm tiger}))
\]
where $G(h_{\rm tiger})$ denotes the adjacency graph constructed from
the TIGER shapefile identified by $h_{\rm tiger}$.

Level 2 is verified by: (a) independently downloading the TIGER shapefile
and verifying $h_{\rm tiger}$, then (b) running \bisect{}'s adjacency
construction and checking that it produces $h_{\rm adj}$.

\subsection{Level 3: Output Provenance}

Level 3 establishes that the output plan was produced by running the
specified algorithm binary with the specified parameters on the specified
data.
\[
  h_{\rm plan} = \mathrm{SHA256}(\text{bisect}(h_{\rm alg}, h_{\rm cfg},
  h_{\rm pl94171}, h_{\rm adj}))
\]

Level 3 is verified by \texttt{bisect label-verify}, which replays
the full computation from the manifest and checks that the output hash
matches $h_{\rm plan}$.

\subsection{Chain Integrity}

The manifest hash $h_{\rm manifest}$ binds all six hashes together.
Any modification to any level---algorithm binary, configuration, download,
adjacency, plan, or the manifest itself---changes $h_{\rm manifest}$.
A modified manifest is detectable by anyone with the original manifest hash,
which is filed with the Secretary of State under the \dia{} protocol.

\begin{theorem}[Manifest Integrity]
Given the SHA-256 collision resistance bound, an adversary cannot produce
a modified manifest $\mathcal{M}'$ such that:
(a) $\mathcal{M}' \neq \mathcal{M}$, and
(b) $h_{\rm manifest}(\mathcal{M}') = h_{\rm manifest}(\mathcal{M})$,
without performing $\geq 2^{80}$ hash evaluations.
\end{theorem}
