\section{Audit Semantics}

The verifier contract is profile-scoped. It can confirm that declared
constraints, lineage, canonical ordering, hashes, and package contents match
the profile. It cannot certify fairness, neutrality, or legal sufficiency
outside the declared audit profile and available context.

Solver reports, heuristic transcripts, and selected-frontier manifests are
upstream evidence. The certificate consumes their final exported artifacts and
checks the package under a declared profile. This keeps exact proofs,
heuristic-improvement summaries, and Pareto-selection records from being
collapsed into one ambiguous claim.

A successful verification should therefore be read as: this plan, with this
context, under this profile, with these hashes and lineage records, passed the
declared checks. It should not be read as: this plan is fair, legally adequate,
or preferred among all available alternatives.

The public verifier workflow is intentionally fixed point shaped: algorithm
output becomes a package, the package certificate is checked, and the checked
package can then be referenced by reports. The neutral surface is
the RPLAN certificate verifier; the BISECT manifest verifier consumes the same
package manifest shape as a bridge.

\begin{quote}
\small
\begin{verbatim}
rplan-cli verify-certificate \
  --certificate audit-certificate.json \
  --plan plan.rplan \
  --context context.rctx

bisect-cli verify --manifest manifest.json
\end{verbatim}
\end{quote}

The first command checks the certificate against the plan and context. The
second command checks the BISECT package manifest bridge. In the public
examples these commands are intentionally run on small packages so that a
reader can inspect every file and every hash involved in the decision.

\begin{table}[htbp]
\centering
\small
\begin{tabular}{p{0.25\linewidth}p{0.35\linewidth}p{0.27\linewidth}}
\toprule
Fixture & Expected verifier signal & Meaning \\
\midrule
Missing context & certificate requires context hash & contextual claims need
the RCTX, not only the RPLAN \\
Plan tamper & certificate plan hash mismatch & assignment content changed
after certification \\
Context tamper or stale context & certificate context hash mismatch & graph,
population, order, or source content changed \\
Canonical unit-order mismatch & certificate context hash mismatch & positional
assignments no longer match the certified unit order \\
Profile mismatch & legal profile chamber mismatch or RPLAN legal profile
mismatch & the certificate cannot be reused for a different declared profile \\
Reserved lineage field & algorithm lineage extra uses reserved certificate
field & method metadata cannot override top-level certificate semantics \\
\bottomrule
\end{tabular}
\caption{Verifier-facing negative fixtures for the U.20 contract.}
\end{table}
