Daniel Webster proposed his major fractions method in 1832 in response
to the perceived inequities of Jefferson's floor-rounding method,
which he argued gave disproportionate advantage to large states.
Webster's insight was that the fair rounding point for a fractional
quota $q \in (s, s+1)$ is the midpoint $s + 0.5$: states with
fractional parts above 0.5 should round up, and those below should
round down.
This is precisely the standard arithmetic rounding rule taught in
every elementary school --- and its simplicity has made it a persistent
alternative to Huntington-Hill in the academic literature.

Webster's method was used for the U.S.\ apportionment of 1842,
the first apportionment conducted under a statute that required
single-member congressional districts \citep{apportionment1842}.
It was reinstated in the Apportionment Act of 1911 and used for
the 1910, 1920 (though the 1920 apportionment was never enacted),
and planned 1930 apportionments.
The failure of Congress to pass a reapportionment act after the 1920
Census --- partly because Webster's method favoured urban, immigrant-heavy
states in a politically fraught era --- contributed to the eventual
move to Huntington-Hill in 1941 \citep{balinski1982fair}.

In proportional representation systems for legislative elections,
Webster is universally known as the Sainte-Lagu\"e method, named for
the French mathematician who independently proposed it in 1910
\citep{sainte-lague1910}.
It is used for seat allocation among parties in Norway, Sweden, Denmark,
and several other countries.
The formal equivalence between Webster and Sainte-Lagu\"e establishes
that the same mathematical principle --- arithmetic mean rounding ---
appears in both U.S.\ apportionment history and modern electoral systems.

This paper's contributions:
\begin{enumerate}
  \item \textbf{Mathematical characterisation}: We prove that Webster
    minimises the sum of squared deviations from exact quotas,
    distinguishing it from Huntington-Hill (which minimises the
    relative population pair test).
  \item \textbf{Empirical comparison}: We compute Webster for 2020
    Census data, show that it agrees with Huntington-Hill for all 50
    states, and identify the states closest to the rounding boundary.
  \item \textbf{Historical documentation}: We document Webster's
    role in three U.S.\ apportionment statutes (1842, 1911, and the
    planned 1930 apportionment that was never enacted).
  \item \textbf{Sainte-Lagu\"e connection}: We establish the formal
    equivalence between Webster and Sainte-Lagu\"e, connecting
    U.S.\ apportionment history to the comparative electoral systems
    literature.
\end{enumerate}
