\section{Implemented Matrix}

\begin{center}
\small
\begin{tabularx}{\textwidth}{lXX}
\toprule
RCOUNT check & Evidence used & What it mechanically proves \\
\midrule
\texttt{contest\_selection\_sum} & Contest summaries & Selection totals plus residual counts equal counted ballots. \\
\texttt{jurisdiction\_contest\_total} & Reporting-unit and jurisdiction summaries & Local totals roll up to the jurisdiction total. \\
\texttt{batch\_summary\_total} & Batch manifests and batch-tagged summaries & Counted, accepted, and rejected ballot accounting is internally consistent. \\
\texttt{lineage\_conservation} & Prior/current reporting units and lineage records & Split/merge records reference real units and have valid cardinality. \\
\texttt{source\_hash\_match} & Source index and package-relative files & Published source bytes match the declared hashes. \\
\texttt{proof\_privacy\_gate} & Inclusion proof records & Voter-facing proofs do not disclose candidate selections or linkable voter metadata. \\
\texttt{cvr\_summary\_total} & CVR rows and public summaries & CVR aggregates reproduce public reporting-unit summaries. \\
\texttt{rla\_sampler\_replay} & CVR population, seed, manifest hash, and algorithm id & Published RLA sample draws replay from public material. \\
\texttt{rla\_margin\_metadata} & RLA margin record and canvassed summary & Reported audit margin fields match public totals. \\
\texttt{rla\_stopping\_rule} & Sample observations, discrepancies, margin, and risk estimate & Declared pass/escalate status follows from the named RCOUNT stopping contract. \\
\texttt{rla\_jurisdiction\_adapter} & Jurisdiction method metadata & State-style record-shape requirements are present for synthetic Colorado and California fixtures. \\
\texttt{manual\_audit\_reconciliation} & Manual audit hand totals and canvassed machine totals & Hand-count totals are within declared tolerance, and declared status agrees. \\
\bottomrule
\end{tabularx}
\end{center}

This matrix is intentionally operational. It names equation ids visible in
\texttt{rcount verify} transcripts, not abstract desiderata. A public reader
can ask which row failed, which evidence file fed that row, and whether the
failure is arithmetic, hashing, privacy, audit, or adapter metadata.
