\documentclass[12pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{lmodern}
\usepackage[T1]{fontenc}
\usepackage{amsmath, amssymb, amsthm}
\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}
\newtheorem{proposition}{Proposition}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}[proposition]
\newtheorem{remark}{Remark}
\DeclareMathOperator*{\argmin}{arg\,min}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage{booktabs}
\usepackage{xcolor}
\usepackage{microtype}
\usepackage{multirow}
\usepackage{array}
\usepackage{algorithm}
\usepackage{algpseudocode}

\title{VRASection: Geographic Alignment Score for\\
       Minority Opportunity District Bisection}

\author{Gio Della-Libera}

\date{Draft --- \today}

\begin{document}

\maketitle

\begin{abstract}
Congressional redistricting algorithms that attempt to satisfy the Voting
Rights Act (\S 2) by adding minority population as an explicit METIS
constraint (``MetisVra'') encounter a practical scale problem in the
companion B.3 experiments: minority-fraction values (0.0--1.0) and census
population counts (10,000--200,000) differ by orders of magnitude, producing
infeasible or degenerate partitions in several tested configurations
\citep{b3paper}.

We introduce \textbf{VRASection}, which abandons explicit minority constraints
and instead exploits the geographic structure of \emph{Gingles Prong 1}:
legally cognizable minority communities are, by definition, sufficiently
large and geographically compact.
This geographic concentration is already encoded in the census-tract adjacency
graph; VRASection makes it measurable through a \emph{geographic alignment score}
$A(\text{split}) = |\text{MVAP}_\text{frac}(L) - \text{MVAP}_\text{frac}(R)|$
layered on top of GeoSection's isoperimetric ratio scan \citep{b8paper}.
In the current BISECT implementation, the score is computed from a demographic
proxy: tract minority fraction multiplied by tract population, loaded from
\texttt{data/\{year\}/demographics}.  True VAP/CVAP inputs are supported in
the audit data model, but the shipped ratio-selection path should be read as a
minority-population spatial proxy rather than a legal finding.

The modified selection objective is:
\[
  \text{score}(\text{ratio}) = \mathrm{EC}_{\text{norm}} - w_\text{vra}
  \cdot A \cdot \max(\mathrm{EC}_{\text{norm}},\, 1)
\]
where $w_\text{vra} = 0.40$ by default.
At this weight, compactness is weighted 50\,\% more heavily than alignment.
That coefficient is useful evidence for a geography-first design record, but
it is not a mechanical safe harbor under \emph{Shaw v.\ Reno} or
\emph{Miller v.\ Johnson}; racial predominance remains a fact-specific legal
inquiry.  The alignment score is best understood as a bounded use of a
race-associated demographic attribute to detect spatial concentration, not as
an automatic Gingles or constitutional certification.

Applied to Alabama (7 districts, 27\,\% Black VAP), VRASection selects
a $2:5$ first-level bisection, separating the Black Belt region from the
northern and coastal areas, compared to GeoSection's $3:4$ split
(near-equal geographic division).
The $2:5$ split more effectively concentrates minority population on one
side for downstream minority-opportunity-district formation.
Full final-plan comparison packages across 6 historically significant VRA
states remain pending; Alabama's first-level $2:5$ result provides the lead
case study.

VRASection requires no user-specified minority-fraction targets.
It passes the MetisVra failure modes by design: there is no second METIS
constraint and no scale mismatch.
We analyse the algorithm's legal risk boundaries, sensitivity to
$w_\text{vra}$, and relationship to the \emph{Callais v.\ Landry} evidence
workflow.
\end{abstract}

\input{sections/01-introduction}
\input{sections/02-related-work}
\input{sections/03-algorithm}
\input{sections/04-evaluation}
\input{sections/05-discussion}
\input{sections/06-conclusion}

\bibliographystyle{plainnat}
\bibliography{references}

\end{document}
