\section{Implementation Contract}

The branch-and-price surface has three contracts: a pricing contract that
generates connected balanced columns or explicit failure statuses, a master
contract that reports objective and bound state, and a package contract that
turns solved fixture outputs into final-plan artifacts.

The key reproducibility requirement is that a reader can tell which loop stage
produced the artifact. A pricing artifact is not a solved partition; a master
report is not necessarily a branch-and-price proof; and an exact fixture
package is scoped to the fixture model, objective, and column universe recorded
in the run context.

In the atlas view, the candidates are whole district columns rather than
individual unit moves. A low-cost column can still be rejected by the master if
it overlaps selected columns or leaves another unit uncovered. The selected
solution must therefore name column ids and coverage status, not only the final
assignment colors.

On the current path4 fixture, the graph has units \(0-1-2-3\), equal
populations, \(k=2\), and a one-percent tolerance. Pricing enumerates connected
balanced subsets. The eligible district columns are \(\{0,1\}\), \(\{1,2\}\),
and \(\{2,3\}\). The middle column is individually valid, but it cannot appear
with either endpoint column without double-covering unit 1 or unit 2. The
master therefore selects the endpoint columns, records selected column ids
\([0,2]\), emits assignment \([0,0,1,1]\), and reports objective 1. This is the
small example that separates column validity from exact cover validity.

\begin{table}[htbp]
\centering
\small
\begin{tabular}{p{0.23\linewidth}p{0.34\linewidth}p{0.29\linewidth}}
\toprule
Contract & Must report & Evidence level \\
\midrule
Pricing & column validity, balance, reduced-cost status & L0 fixture \\
Master & selected columns, objective, bound status & L0/L1 exact package \\
Package & final assignments, manifest, certificate & L1 audit path \\
\bottomrule
\end{tabular}
\caption{Branch-and-price staged implementation contracts.}
\end{table}

\begin{table}[htbp]
\centering
\small
\begin{tabular}{p{0.22\linewidth}p{0.29\linewidth}p{0.35\linewidth}}
\toprule
Report field & Current value family & Interpretation \\
\midrule
Report schema & branch-price report v1 & Stable JSON envelope for U.17 evidence \\
Status & formulation-only, exact-fixture-optimal, infeasible & Allowed claim vocabulary \\
Pricing counts & rounds and generated columns & Size of generated column pool \\
Bounds & lower bound, upper bound, gap & Master evidence when solved \\
Formulation size & constraints, columns, incidence nonzeros & Rebuildability and scale context \\
Solution & selected column ids, assignment, objective & Bridge from columns to final RPLAN \\
Parameter hash & SHA-256 digest & Tamper check for report-defining parameters \\
\bottomrule
\end{tabular}
\caption{Concrete branch-price report fields emitted by \texttt{bisect-column}.}
\end{table}

\begin{table}[htbp]
\centering
\small
\begin{tabular}{p{0.24\linewidth}p{0.34\linewidth}p{0.28\linewidth}}
\toprule
Status & Meaning & Allowed claim \\
\midrule
Pricing-valid & connected balanced columns generated & pricing contract works on fixture \\
Pricing-failed & no valid column or explicit fallback & failure is reproducible \\
Master-bounded & objective and bound reported & scoped bound under column pool \\
Fixture-solved & exact fixture package emitted & exact fixture result \\
Formulation-only & model/report built without solve & staged implementation only \\
\bottomrule
\end{tabular}
\caption{Status vocabulary for branch-and-price outputs.}
\end{table}

The status table intentionally uses a richer explanatory vocabulary than the
current enum. The implementation enum has \texttt{formulation-only},
\texttt{exact-fixture-optimal}, and \texttt{infeasible}. The paper-level terms
``pricing-valid,'' ``master-bounded,'' and ``fixture-solved'' are reader-facing
states that explain which part of the lifecycle produced the evidence.
