\section{Implementation in \texttt{BISECT research}}
\label{sec:implementation}

\subsection{Command-Line Interface}

The diagnostic math described in this paper is implemented in the
\texttt{bisect-analysis} crate. The stable command-line surface for metric
traces is \texttt{BISECT research validate-ensemble}. It consumes one JSONL file
per chain, where each line contains a step number and a metric map.

\begin{verbatim}
  BISECT research validate-ensemble \
      --ensemble-dir outputs/ensembles/nc_2020/chains \
      --output validate-ensemble.json \
      --rhat-threshold 1.1 \
      --ess-min 500 \
      --strict
\end{verbatim}

The command:
\begin{enumerate}
\item Reads every \texttt{*.jsonl} file in \texttt{--ensemble-dir} as one chain.
\item Validates that R-hat has enough equal-length chains.
\item Computes $\rhat$ and $\ess$ for each metric present in every chain.
\item Optionally records enacted-plan percentile ranks when supplied.
\item Reports pass/fail against the explicit thresholds from
      Section~\ref{sec:statutory}.
\end{enumerate}

\subsection{Output Format}

The output is a JSON file with the following structure:

\begin{verbatim}
{
  "state": "NC",
  "year": 2020,
  "k": 14,
  "chains": 10,
  "steps": 10000,
  "diagnostics": {
    "rhat_pp": 1.02,
    "rhat_seats": 1.03,
    "ess_pp": 695,
    "ess_seats": 1538,
    "hamming_lag1": 0.87
  },
  "thresholds": {
    "rhat_max": 1.1,
    "ess_min": 500,
    "hamming_max": 0.95
  },
  "verdict": "PASS"
}
\end{verbatim}

A \texttt{PASS} verdict indicates that the supplied metrics meet the explicit
thresholds used for this validation run. It is evidence about convergence of
the recorded ensemble metrics, not a standalone legal conclusion about the
enacted plan.

\subsection{Implementation Details}

\paragraph{ReCom proposal.}
The \recom{} step selects two adjacent districts uniformly at random,
merges their tract sets, generates a random spanning tree of the merged
subgraph, and bisects the spanning tree at its centroid edge subject to
population balance.
The implementation uses the BISECT spanning-tree and ensemble components when
chains are generated inside the workspace.

\paragraph{R-hat computation.}
Each chain's PP time series is stored in a ring buffer.
$\rhat$ is computed from the last half of each chain (disarding the
burn-in), following the recommendation of \citet{vehtari2021}.

\paragraph{ESS computation.}
$\ess$ uses the initial monotone sequence estimator of \citet{geyer1992}
to truncate the autocorrelation sum before it becomes noisy.

\paragraph{Hamming autocorrelation.}
The Hungarian algorithm for label canonicalization is $O(k^3)$ per plan
pair. For large $k$ we use the greedy approximation (sort districts by
centroid proximity), which is accurate to within 1\% for nearly-converged
chains.

\subsection{Relation to Diagnostics Code}

The diagnostics code is exposed through the BISECT analysis crate:

\begin{verbatim}
  bisect_analysis::ensemble_diagnostics::{
      rhat,          // Gelman-Rubin R-hat
      ess,           // Effective sample size
      hamming_lag1,  // Lag-1 Hamming autocorrelation
      diagnose,      // Run all three and return verdict
  }
\end{verbatim}

The module has 21 unit tests covering boundary cases (all chains identical,
all chains diverged, single-chain degenerate input) and is part of the
standard \texttt{cargo test} suite.
