\section{Selection Matrix}

The matrix maps task type to algorithm family, CLI surface, evidence
requirement, and audit artifact. Construction methods provide baselines, search
methods explore alternatives, exact methods can provide bounded certificates,
multi-objective methods expose trade-offs, and audit methods verify declared
artifact contracts.

\subsection*{Compositor Framing}

The selection matrix is also a three-layer compositor checklist.
For any recommended workflow, the operator should name the structure layer,
weight layer, and search layer before interpreting the result.
Changing one layer creates a new run recipe; it should not be described as an
isolated algorithm comparison unless the other two layers are held fixed.
The candidate object is therefore the resolved run recipe itself: structure
choice, weighting choice, search mode, backend settings, and config hash.
The decision rule is the row chosen for the intended claim, and the consequence
is the artifact bundle that must be retained for audit.

\begin{table}[htbp]
\centering
\scriptsize
\resizebox{\linewidth}{!}{%
\begin{tabular}{p{0.15\linewidth}p{0.17\linewidth}p{0.20\linewidth}p{0.20\linewidth}p{0.18\linewidth}}
\toprule
Need & Preferred family & Required evidence & Audit handoff & Do not claim \\
\midrule
Fast baseline & Construction & L0/L1 fixture and smoke tests & RPLAN with constructor metadata & Optimality or distribution \\
Improve candidate & Local/LNS search & Validity-preserving move trace & RPLAN/RCTX run summary & Global bound \\
Bounded claim & Exact optimization & Bound, witness, or solve report & Certificate and manifest & Legal sufficiency \\
Distributional context & Ensemble/SMC & Protocol and sampled-plan record & Exported plans and context & Equal likelihood without target \\
Trade-off choice & Pareto selection & Frontier and selection rule & Selected-plan package & Best on every objective \\
Artifact verification & Audit verifier & Declared profile pass/fail & Certificate bundle & Substantive fairness \\
\bottomrule
\end{tabular}
}
\caption{Method-family selection by claim need, audit handoff, and claim boundary.}
\end{table}

The matrix is conservative by design. If the claim requires a bound, a search
heuristic can provide useful candidates but should not be described as exact.
If the claim requires distributional context, a single optimized plan is an
example, not a sample. If the claim is legal or policy-facing, the workflow
must carry enough evidence to support interpretation without letting the audit
certificate stand in for that interpretation.

\subsection*{Preflight Questions}

Before selecting a workflow, the operator should answer four questions:
What claim is the output meant to support? What evidence level is required for
that claim? Which CLI surface or crate currently produces the relevant plan or
certificate artifact? Which RPLAN/RCTX sidecars must be retained so another
reader can inspect the run?

\subsection*{Worked Reading}

Suppose an operator needs a fast baseline for a new state. The matrix routes the
work to a construction family and asks for fixture coverage, smoke-test output,
and constructor metadata in the RPLAN package. The resulting claim is modest:
``this run produced a package matching the declared construction recipe.'' If
the same operator wants to say the result is typical, the row changes to
Ensemble/SMC and the package must include a sampling protocol and sampled-plan
record. If the operator wants to say no better solution exists under a stated
objective, the row changes again to Exact optimization and requires a solve
report or bound.

The rows are intentionally not mutually exclusive over a project lifecycle.
A careful study may begin with construction, improve with local search, compare
against ensembles, and finally package an audit certificate. U.12 only requires
that each exported claim name the row whose evidence supports it.

\subsection*{Evidence Fields}

A publication-ready selection record should include the resolved structure,
weights, search mode, seed rule or seed budget, objective, selected artifact,
RPLAN/RCTX identifiers, sidecar list, certificate or solve-report path when
available, and the claim boundary stated in plain language.
This prevents a valid audit certificate from being read as an optimality proof,
a local-search improvement from being read as a global bound, or a percentile
seed run from being read as an ensemble over all valid plans.
