\section{Method}

The staged method constructs assignments under population capacity limits. It
records whether the implemented model produced a valid result, a result needing
repair, or an infeasible-capacity result. When capacity is infeasible, the
output carries an explicit witness rather than emitting a final plan.

The atlas framing treats eligible arcs as the candidate objects. A unit can
send supply only along arcs allowed by the declared graph/profile, and a
selected arc remains valid only while the target bin stays inside its lower and
upper capacity bounds. The important failure is therefore concrete: the paper
should name the bound that fails and whether the artifact is a different
feasible extraction, a model-relative infeasibility witness, or an invalid
post-extraction block.

\subsection{Reproducible Procedure}

The constructor uses the graph and unit attributes supplied by the BISECT state
pipeline. The baseline is feasibility- and audit-oriented. It is not yet the
branch-and-cut or branch-and-price machinery planned for U.16 and U.17.

\begin{enumerate}
  \item Select deterministic farthest-point seeds on the adjacency graph.
  \item Compute lower, ideal, and upper district-population bounds from the
        declared tolerance.
  \item Grow balanced frontiers from the seeds. A frontier arc is eligible only
        when assigning that unit keeps the target district at or below the
        upper capacity bound.
  \item Fill remaining unassigned units by nearest seed only after frontier
        growth stalls, then validate population and connectedness.
  \item Mark the result \texttt{valid} when assignment and validation checks
        pass.
  \item Mark the result \texttt{needs-repair} when extraction exists but fails
        declared checks, and run the staged small-fixture repair when enabled.
  \item Mark the result \texttt{infeasible-capacity} when no balanced capacity
        assignment exists for the declared fixture. Invalid input returns an
        error before a flow artifact is emitted.
\end{enumerate}

\begin{table}[h]
\centering
\small
\resizebox{\linewidth}{!}{%
\begin{tabular}{lll}
\toprule
Status & Meaning & Artifact behavior \\
\midrule
\texttt{valid} & Model produces a validated assignment & Write final sidecars \\
\texttt{needs-repair} & Extracted assignment fails declared checks & Try staged repair or block export \\
\texttt{infeasible-capacity} & Model/profile cannot satisfy capacity & Write model-relative witness \\
\texttt{invalid input} & Input shape is malformed & Return an error before artifact emission \\
\bottomrule
\end{tabular}
}
\caption{Flow-construction status taxonomy. Infeasibility is relative to the
declared model and profile.}
\end{table}

\subsection{Worked Frontier Decision}

The current baseline does not solve a general min-cost-flow instance. Its
flow-like decision is a deterministic frontier assignment under capacity. If
the upper bound is \(55\), a candidate arc into a district with population
\(48\) is rejected for a unit of population \(10\), while a lighter unit remains
eligible.

\begin{table}[h]
\centering
\small
\resizebox{\linewidth}{!}{%
\begin{tabular}{lllll}
\toprule
District frontier & Candidate unit & Current pop. & Pop. after arc & Decision \\
\midrule
A & \(u_7\) pop. 10 & 48 & 58 & reject: exceeds \(55\) \\
B & \(u_8\) pop. 6 & 47 & 53 & accept: within bound \\
\bottomrule
\end{tabular}
}
\caption{T.17 exposes capacity-gated frontier arcs. Rejection is evidence about
the declared model, not evidence that all possible districting models fail.}
\end{table}
