Congressional redistricting determines not only who represents each community but shapes the
democratic health of the legislature itself. Yet most algorithmic redistricting research evaluates
maps on geometric properties---compactness, population equality, minority representation---without
connecting these properties to the downstream democratic outcomes that redistricting is ultimately
meant to serve. This paper establishes a framework linking map geometry to four downstream outcome
dimensions: electoral competitiveness (the frequency and margin of competitive elections), voter
turnout (participation rates conditional on district competitiveness), legislative polarization
(the ideological distance between parties' median members), and constituent-representative
geographic distance (mean travel time from constituent to representative office).

We survey prior literature across all four dimensions, with particular attention to the
quasi-experimental identification strategies that exploit redistricting cycles to establish
causal rather than merely correlational relationships. Key priors: \citet{caugheyTausanovitch2017}
show that state legislative maps affect policy outcomes through ideological selection;
\citet{rodden2019} demonstrates that geographic sorting produces a structural Democratic
efficiency disadvantage that gerrymandering amplifies; \citet{jacobson2015} documents
the secular decline of competitive seats from the 1990s through the 2020 cycle.

The central prediction of the O-series, developed formally in O.1--O.4, is that compact
algorithmic plans produced by \textsc{bisect} predict better outcomes on all four dimensions
independent of partisan composition. The mechanism: gerrymandering requires non-compact geometry
to achieve its effects (packing and cracking create district ``necks'' that increase travel
distance, reduce intra-district political heterogeneity, reduce competitiveness, and select for
ideologically extreme candidates). Compact plans disrupt these mechanisms even when they are
not drawn to produce any particular partisan outcome.

This framework paper synthesizes the O.1--O.4 companion papers and provides the identification
strategy, theoretical foundations, and measurement protocols used throughout the O-series.
