\label{sec:legal}

\subsection{Experimental Design}

Quantitative results for M.8 composite community weights are deferred to
Phase~2, pending completion of all component signal data pipelines.
This section describes the experimental protocol governing Phase~2 execution.

\paragraph{Component data availability.}
The M.8 composite activates signals as their data becomes available.
For the Phase~2 experiments, the following data availability is assumed:

\begin{itemize}
  \item M.6 zone co-membership: available immediately (TIGER/Line, no download
    required beyond standard \textsc{bisect} data fetch).
  \item M.1 economic character: available immediately (LODES WAC already
    downloaded for NC, WI, TX).
  \item M.2 land use: available after NLCD GeoTIFF download and zonal statistics
    computation (Phase~2 M.2 prerequisite).
  \item M.3 housing character: available after ACS tract-level download
    (Phase~2 M.3 prerequisite).
  \item M.4 commuting shed: available after LODES OD download (Phase~2 M.4
    prerequisite).
  \item M.5 topographic: available after SRTM download (Phase~2 M.5 prerequisite).
  \item M.7 transit: available only for states with GTFS data; may be unavailable
    for initial NC and WI runs.
\end{itemize}

Phase~2 will run the composite using the minimum three-signal set (M.6 + M.1 + M.2)
as soon as NLCD data is available, and progressively add signals as their data
pipelines complete.

\paragraph{States.}
Three states are selected for Phase~2 composite evaluation:

\begin{itemize}
  \item \textbf{North Carolina ($k=14$).}
    NC provides the richest test environment for the composite.
    NC has coastal water boundaries (M.2 hard cut), concentrated employment
    clusters (M.1 economic character), mixed rural-urban land use (M.2 land use),
    diverse school district patterns (M.6 zone), distinct commuting sheds
    (M.4 Research Triangle vs.\ Charlotte), and modest topographic variety
    (Piedmont vs.\ Mountains vs.\ Coastal Plain).
    All six non-transit signals are expected to be meaningful for NC.
  \item \textbf{Wisconsin ($k=8$).}
    WI provides a partial test.
    M.4 commuting shed is expected to show a Milwaukee suburban clustering effect.
    M.2 land use will be tested for the urban-agricultural boundary.
    M.1 economic character showed a null result in B.27, which is correctly
    expected behavior.
    WI tests whether the composite aggregates meaningfully when one component
    contributes no signal.
  \item \textbf{Texas ($k=38$).}
    TX provides the large-state test.
    TX has Gulf Coast water boundaries (M.2), petroleum and logistics
    industrial concentrations (M.1), diverse land cover from East Texas
    forest to West Texas desert (M.2), and distinct metro commuting sheds
    (Houston, Dallas-Fort Worth, San Antonio, Austin are non-overlapping
    functional economic areas, M.4).
    TX also has the largest census tract count (6,877 tracts), providing
    a computational scalability test.
\end{itemize}

\subsection{Primary Metrics}

Four primary metrics will be reported for Phase~2.

\paragraph{Within-district economic variance.}
The M.8 composite claims a $\geq$25\% reduction in mean within-district
economic variance compared to standard-bisect geographic weights on NC-14:
\begin{equation}
  \frac{\overline{\sigma}_{\mathrm{JPR}}^{\mathrm{geographic}} -
        \overline{\sigma}_{\mathrm{JPR}}^{\mathrm{composite}}}
       {\overline{\sigma}_{\mathrm{JPR}}^{\mathrm{geographic}}}
  \geq 0.25
  \label{eq:variance-reduction}
\end{equation}
This prediction extends the M.1 finding: adding zone co-membership (M.6),
land use (M.2), and housing character (M.3) signals on top of M.1 should
produce additional economic coherence within districts.\textsuperscript{$\dagger$}

\paragraph{Zone boundary crossings.}
The M.8 composite claims zero school district boundary crossings in the
NC-14 plan, given the 0.30 weight assigned to M.6 zone co-membership.
This is a strong claim: even a 30\% weight contribution from a signal that
assigns zero similarity to cross-school-district adjacencies should be
sufficient to prevent the algorithm from cutting within school districts.
Phase~2 will verify this by counting the number of NC school district
boundaries crossed by district lines in the composite-weighted plan.\textsuperscript{$\dagger$}

\paragraph{Polsby-Popper compactness.}
The M.8 composite predicts Polsby-Popper compactness equal to or greater
than geographic weights for NC-14, WI-8, and TX-38.
The mechanism is the same as in M.1 and M.2: cutting at community seams
(economic transitions, land use boundaries, administrative zone edges)
tends to produce cleaner, rounder districts than cutting at arbitrary
population-balance boundaries.\textsuperscript{$\dagger$}

\paragraph{Partisan neutrality check.}
The M.8 composite claims a correlation below 0.05 (Pearson $r$) between
composite edge weights and tract-level Democratic vote share:
\begin{equation}
  \bigl|\rho\bigl(w_{\mathrm{comp}}(u,v),\, V_D(u) + V_D(v)\bigr)\bigr| < 0.05
  \label{eq:partisan-check}
\end{equation}
where $V_D(t)$ is the Democratic vote fraction in tract $t$ from the 2020
presidential election.
This correlation measures whether the composite systematically assigns
higher or lower weights to edges in Democratic or Republican-leaning
neighborhoods.
A correlation below 0.05 confirms that no partisan signal is embedded
in the composite.

Note that this check uses election data only for \emph{validation}, not
as an input.
The validation computes the correlation after the weights are finalized;
election data is never used to compute the weights themselves.
The validation data is used to confirm that the claim ``no partisan signal
entered the composite'' is empirically true, not merely asserted.\textsuperscript{$\dagger$}

\subsection{Comparison Baseline}

All Phase~2 metrics compare the composite-community-weighted plan against
the geographic-weights baseline on the same state, year, seed, and algorithm
configuration.
No multi-seed comparison is performed in Phase~2 (designated Phase~3).

\subsection{Phase 2 Timeline}

All quantitative results in this section are marked ${}^{\dagger}$Phase~2.
Prerequisites for Phase~2 execution:
\begin{enumerate}
  \item M.2 land use data pipeline complete (NLCD GeoTIFF downloaded and
    zonal statistics computed for NC, WI, TX).
  \item M.3 housing character data pipeline complete (ACS 5-year estimates
    downloaded for NC, WI, TX).
  \item M.4 commuting shed data pipeline complete (LODES OD downloaded
    for NC, WI, TX).
  \item M.5 topographic data pipeline complete (SRTM downloaded for NC, WI, TX).
  \item M.6 zone co-membership weights verified for NC, WI, TX (already
    available pending M.6 paper completion).
  \item Composite formula implemented in Rust with all five L0 test invariants
    passing.
\end{enumerate}
The composite run for NC-14 using the minimum three-signal set (M.6 + M.1 + M.2)
can proceed as soon as step~1 is complete.
