\section{Fixtures}

The positive fixture is \texttt{rla-replay}. It extends the V.6 CVR summary
fixture with one audit round for \texttt{syn-2024-mayor}. The package contains
140 CVR contest rows, a 5 percent risk limit encoded as \texttt{50000} ppm, a
20-digit public seed, and 12 replayed sample draws.

\begin{center}
\small
\begin{tabularx}{\textwidth}{lX}
\toprule
Path & Role \\
\midrule
\path{audits/rla.ndjson} & RLA replay records and sample draws. \\
\path{normalized/cvr.ndjson} & Contest CVR rows used to derive the audit population. \\
\path{normalized/summaries.ndjson} & Public totals reconciled against CVR rows. \\
\path{sources/source-index.json} & Source hash evidence for the synthetic export. \\
\path{transcripts/verify-transcript.json} & CLI verifier transcript. \\
\bottomrule
\end{tabularx}
\end{center}

The negative fixture is \texttt{bad-rla-replay}. It changes the first published
sample draw to \texttt{cvr:P-999:999}. The manifest hash and source hash still
verify, but sampler replay recomputes draw 0 as \texttt{cvr:P-001:022} and
fails the package.

The stopping fixture is \texttt{rla-stopping}. It uses the same replayed sample
but adds one observation per draw. Each observation repeats the sampled CVR id
and records the observed contest interpretation. With zero discrepancies and
\texttt{max\_discrepancies = 0}, the declared status \texttt{pass} verifies.

The stopping negative fixture is \texttt{bad-rla-stopping}. It leaves sampler
replay intact, but changes one observed interpretation. The computed status is
\texttt{escalate}, so the declared \texttt{pass} fails under
\texttt{rla\_stopping\_rule}.

The discrepancy fixture is \texttt{rla-discrepancy}. It contains the same
observed mark mismatch as \texttt{bad-rla-stopping}, but declares the mismatch
as a selection discrepancy and declares \texttt{escalate}. The package passes
because the observation, discrepancy taxonomy, and status agree.

The taxonomy negative fixture is \texttt{bad-rla-discrepancy}. It declares the
same observed mark mismatch as a residual discrepancy. The verifier recomputes
a selection discrepancy and fails the taxonomy check.

The margin fixture is \texttt{rla-margin}. It adds reported winner votes, loser
votes, reported margin, and diluted-margin denominator to the audit record.
Those fields are recomputed from the canvassed jurisdiction summary before the
audit passes.

The margin negative fixture is \texttt{bad-rla-margin}. It changes the
reported margin while leaving the public summary unchanged. The package fails
under the margin metadata check.

The statistical fixture is \texttt{rla-statistical}. It keeps the replayed
sample, observations, and margin metadata, then names
\texttt{comparison-margin-threshold-v1} as the stopping method. The audit
record declares the recomputed risk estimate in parts per million. The verifier
recomputes that value from sample size, reported margin, and discrepancy count,
then accepts the declared \texttt{pass} only when both the discrepancy threshold
and risk limit are satisfied.

The statistical negative fixture is \texttt{bad-rla-statistical}. It leaves the
sample, observations, margin metadata, and declared status unchanged, but edits
the declared risk estimate by one part per million. The package therefore fails
under \texttt{rla\_stopping\_rule}; the transcript exposes the declared and
computed risk estimates side by side.

The jurisdiction fixture is \texttt{colorado-rla}. It adds a Colorado-style
method id to the statistical fixture. The added adapter check verifies that the
public seed is exactly 20 decimal digits and that the record carries the
expected sampler, margin, stopping method, risk estimate, and declared status
fields.

The jurisdiction negative fixture is \texttt{bad-colorado-rla}. It keeps the
sample replayable by recomputing draws and observations from an edited seed,
but the edited seed contains a non-digit character. The generic sampler and
stopping layers can still be made internally consistent; the jurisdiction
adapter rejects the package under \texttt{rla\_jurisdiction\_adapter}.

The public-tool fixture is \texttt{california-rla}. It adds a California-style
method id, ballot manifest format id, audit software id, and public source URL
to the statistical fixture. The adapter verifies that this public tooling
metadata is present and shaped like a reproducible public audit record.

The public-tool negative fixture is \texttt{bad-california-rla}. It keeps the
audit otherwise internally consistent, but changes the source field to a
non-URL identifier. The package fails under
\texttt{rla\_jurisdiction\_adapter}.
