\section{Boundaries}

This slice verifies sample selection and a deliberately small family of
stopping contracts. The newest method,
\texttt{comparison-margin-threshold-v1}, is a deterministic comparison-margin
risk-estimate check: it binds sample size, reported margin, discrepancy count,
declared risk estimate, and risk limit into one replayable equation. It is not
a Kaplan-Markov, ballot-polling, or full statutory ballot-level comparison
risk measure. It also does not decide whether a jurisdiction's legal RLA method
permits replacement sampling, modulo reduction, or a particular manifest
construction.

Those boundaries are deliberate. V.7 now has a replayable sampler surface, a
minimal observation-to-status surface, explicit discrepancy taxonomy, bound
margin metadata, and a first named statistical-risk surface that later work can
reuse. The next layer should add Kaplan-Markov or jurisdiction-specific
stopping math, escalation workflows, and fuller state-specific sampler
adapters. The Colorado- and California-style adapters landed here are only
record-shape, seed-format, and public-tool metadata contracts for synthetic
fixtures; they are bridges toward statutory review, not legal certification
claims.

The privacy boundary also matters. A public RLA sample can reveal ballot-level
records selected for audit. RCOUNT can make the sample reproducible, but
jurisdictions still need publication rules for small batches, rare ballot
styles, and any join between CVRs and voter-facing inclusion tokens.
